---
title: "dcrab_sdmTMB"
format: html
editor: visual
---

## Intro

Putting together a simple spatial model for DCRB to look for useful/consistent spatial patterns. The core dataset is the joined and 5k-gridded, monthly summarized data for 2010-2023

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = F)

# import libraries
library(tidyverse)
library(sf)
library(cowplot)
library(rnaturalearth)
library(viridis)
library(knitr)
library(here)
library(sdmTMB)

options(dplyr.summarise.inform = FALSE)

# change default ggplot theme for plotting
plot_theme <- theme_minimal()+theme(panel.border = element_rect(color='black',fill=NA),
                                    panel.background=element_rect(fill="#abd0d6"),panel.grid=element_line(color='gray70',linewidth = 0.2))
theme_set(plot_theme)
```

## Import Data

### Gridded Fishing Effort

This is summarized Dungeness crab fishing effort from a combination of VMS and fish tickets, projected/summarized on a 5km, coastwide grid, monthly.

```{r}
dat <- read_csv(here('dcrb_5km_monthly_nonconfidential_df.csv'))
```

## Spatial Grid

Next, load the spatial layers we need for mapping.

```{r}
# 5km grid, entire coast
grid5k <- read_sf(here('spatial_data','grids','fivekm_grid_polys_shore_lamb.shp'))

# background map, from rnaturalearth
coast <- ne_states(country='United States of America',returnclass = 'sf') |>
  filter(name %in% c('California','Oregon','Washington','Nevada')) |>
  st_transform(crs = st_crs(grid5k))
```

```{r}
# filter the overall grid using grid cells with DCRB effort in any month/year
cells_we_need <- unique(dat$GRID5KM_ID)
grid5k_filt <- grid5k |> filter(GRID5KM_ID %in% cells_we_need)
```

```{r}
ll <- grid5k_filt |> select(GRID5KM_ID,Centroid_y,Centroid_x)
dat <- dat |> left_join(ll,by=join_by(GRID5KM_ID))
```

### Expand?

With zeroes?

```{r}
dat_expanded <- dat |> 
  complete(GRID5KM_ID,nesting(month_numeric,year_numeric),
           fill=list(n_vms_records=0))
  
```

```{r}
x <- count(dat,year_numeric,month_factor)
y <- x |> group_by(month_factor) |> summarize(mean_n=mean(n)) |> 
  mutate(mth=factor(month_factor,levels=c("November","December","January","February","March","April","May","June","July","August","September","October")))
y |> ggplot(aes(mth,mean_n))+geom_point()+geom_hline(yintercept=30,linetype=2)

mths_swap <- tibble(month_numeric=1:12,crab_month=c(3:12,1:2))

dat_filt <- dat |> 
  left_join(mths_swap,by=join_by(month_numeric)) |> 
  filter(!(month_numeric%in%c(8,9,10)))
```

```{r}
dat_filt |> ggplot(aes(n_vms_records))+geom_density()
dat_filt |> ggplot(aes(log(n_vms_records)))+geom_density()

```

## Models

### Mesh

```{r}
mesh <- make_mesh(dat_filt |> mutate(X=Centroid_x/1000,Y=Centroid_y/1000),c("X","Y"),cutoff=50)
plot(mesh)
```

```{r}
m1 <- sdmTMB(
  n_vms_records~s(crab_month,k=3),
  data=dat_filt,
  mesh=mesh,
  family=gaussian(link='log'),
  spatial = T,
  anisotropy = T,
  time="year_numeric",
  spatiotemporal = 'iid')
```
