---
title: "01_compare_cdfw_block_vms_data"
author: "Jameal Samhouri"
date: "Written beginning 2025-06-05. Last Run `r Sys.Date()`"
output: 
  html_document:
      toc: true
      toc_float: true
geometry: margin=1in
---

# Setup

```{r setup, include=FALSE}

# import libraries
library(tidyverse)
library(here)
library(fredr)


```

```{r create-output-directories}
# make hindcast output directory, if doesn't yet exist
output_dir <- here('Confidential', 'hindcast_output')
if (!dir.exists(output_dir)) dir.create(output_dir)

# create hindcast output subdirectory name based on system date
output_subdir_name <- paste0("hindcast_output_", Sys.Date())

# make hindcast output subdirectory, if doesn't yet exist
output_subdir <- here('Confidential', 'hindcast_output', output_subdir_name)
if (!dir.exists(output_subdir)) dir.create(output_subdir)

# make tables, figures, and maps subdirectories
for (temp_name in c('tables', 'figures')) {
  temp_subdir <- here('Confidential', 'hindcast_output', output_subdir_name, temp_name)
  if (!dir.exists(temp_subdir)) dir.create(temp_subdir)
}
rm(temp_name, temp_subdir)
```

# Introduction

This document compares the spatial distribution of CA Dungeness crab fishing activity based on self-reported CDFW block data from fish tickets and geolocations obtained from VMS data.

Analysis with fish tix only:

1) Summarize the number/% of tix and number/% of vessels by time period for which CDFW offshore blocks are reported, and examine for structural bias (vessel size, time of year, region, etc)

Analysis with matched VMS and fish tix:

1) Using the VMS record-by-record df, calculate distance of VMS point from centroid or edge of CDFW block reported in fish ticket; 
- Make plot of frequency distribution showing distance of VMS point from CDFW block, with 50/75/90% labeled 

2) Summarize VMS record-by-record df by trip_ID, and calculate % of VMS points within 0/1/2-degree neighboring blocks across time periods. Try this package: https://link.springer.com/article/10.1007/s10980-024-01930-z 
- 0-degree neighbor = inside/outside of CDFW block
- Make plot of % of VMS points (y-axis) and degree neighbor (x-axis), each point estimate representing a time period (eg, yr)
- Facet by region, vessel size, etc
- Consider summarizing by vessel first

3) Spatial visualizations (maps)
- Summarize VMS record-by-record df by trip_ID and CDFW block
    - # tix per CDFW block vs # pings/day per CDFW block
    - # unique vessels per CDFW block based on tix vs # unique vessels/day per CDFW block based on VMS
- Follow this vignette to make a correlation map: https://statnmap.com/2018-01-27-spatial-correlation-between-rasters/ 





# Load and join data frames for processing

Fish tix (fishtix_vlengths_withFTID_yyyy.rds)
Matched VMS and fish tix (matched_filtered_withFTID_length_yyyy.rds) 
10 arc minute grid. https://github.com/noaa-nwfsc/VMS-pipeline/tree/main/spatial_data

NWFSC_DSACDFW_DcrabPermitDetail2013-2023_08072023.xlsx
TrapTierNumberKey.csv

The code below borrows heavily from Brooke Hawkins' notebook "10_blh_dungeness-crab-time-series.Rmd".

## Load data

1. Load the cleaned fish ticket data for specified years, which will be used to plot revenue and landings across time.

Pre-requisite: Run the pipeline steps 1-6 (including interpolation) for calendar years 2011-2023 for DCRB.

```{r load-data}
# choose years of data to load
load_years <- 2011:2023

# load fish ticket data joined to vessel length data, but before it was joined to VMS data
ticket_df <- purrr::map(load_years, function(ly) {
  read_rds(here('Confidential', 'processed_data', 'processed_2025-03-19', 'fish_tickets', paste0('fishtix_vlengths_withFTID_', ly, '.rds')))
}) %>% bind_rows()

# the ticket df above is missing some columns, so also load the raw fish ticket data, downloaded by Brooke Hawkins
ticket_df2 <- purrr::map(load_years, function(ly) {
  read_csv(here('Confidential', 'raw_data', 'fish_tickets', paste0('fish_tickets_', ly, '.csv')), #) %>%
    #dplyr::select(FISH_TICKET_ID, CDFW_AREA_BLOCK)
  #}) %>% bind_rows()
           col_types = list(
             HILLE_PERMIT = "c",
             DEALER_NUM = "c",
             FTID = "c",
             NUM_OF_FISH = "n",
             EFP_CODE = "c",
             EFP_NAME = "c",
             GF_PERMIT_NUMBER = "c",
             TRIP_SEQ = "i",
             GEAR_CODE = "d"
             )
  )
  }) %>% bind_rows() # works for 2013, but not for all years. 06-13-2025

# 06-11-2025 changed from read_csv to read.csv to avoid warning messages associated with the HILLE_PERMIT column in at least the 2013 data file. that didnt work. so on 06-12-2025, decided to just read in columns i need. that didnt work either and brooke addressed this issue more comprehensively here: https://github.com/noaa-nwfsc/VMS-pipeline/blob/main/code/notebooks/brooke/16_blh_check-new-fish-tickets.md


```

2. Load inflation adjustment factors from FRED.

Adjust revenue for inflation using Personal Consumption Expenditures: Chain-Type Price Index, [PCEPI](https://fred.stlouisfed.org/series/PCEPI) from FRED. PCEPI index is chosen per Dan Holland's suggestion. The reference year is 2017 by default. This is a monthly time series that is used by multiplying the revenue by 100 / PCEPI. It is used at a monthly time scale (rather than taking an annual average).

Pre-requisite: Create a FRED account and [API key](https://fred.stlouisfed.org/docs/api/api_key.html). This product uses the FRED® API but is not endorsed or certified by the Federal Reserve Bank of St. Louis. 

```{r inflation-adjustment}
# insert your FRED API key
fredr_set_key('ea19d3d447eb4dc773db00c5a2e61f22')

# download the monthly inflation adjustments from FRED from start of min load year to end of max load year
fred_pcepi <- fredr(
    series_id = "PCEPI", 
    observation_start = as.Date(paste0(min(load_years), "-01-01")),
    observation_end = as.Date(paste0(max(load_years), "-12-31"))
  ) %>%
  rename(year_month = date, pcepi = value) %>% 
  select(-c(realtime_start, realtime_end, series_id))

# write inflation adjustment factors
write.csv(file = here("Confidential", "hindcast_output", output_subdir_name, "tables", "inflation_pcepi.csv"), x = fred_pcepi, row.names = FALSE)

# plot inflation adjustment factors
fred_pcepi %>% ggplot(aes(x = year_month, y = pcepi)) + geom_line() + scale_x_date(date_breaks = "1 year", date_labels = "%b %Y")
ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "inflation_pcepi.png"), create.dir = TRUE)
```


## Transform data

Transform the joined, cleaned, interpolated VMS and fish ticket data:

1. Filter for dungeness crab related records.
2. Add temporal columns.
3. Adjust revenue for inflation with data from FRED.
4. add CDFW_AREA_BLOCK and permit tier info (not yet included 06-10-2025)

From Brooke: find the field in the raw fish tickets and join there instead

Longer version - Interesting, yeah that's silly I wasn't thinking about that when you asked for feedback on the joins, but that's probably correct. Many columns are dropped in step 1 when the fish tickets are cleaned, and that drop propagates through the rest of the pipeline. The fastest workaround is to join the raw fish tickets with the fish ticket ID and get it from there.
I'd faced the same issue with the sablefish VMS pipeline and had incorporated the FOS_GROUNDFISH_SECTOR_CODE (which identifies CS/LE/OA) to be kept in all the steps in the pipeline, but it was more work and not that helpful, I ended up going back to the raw fish tickets anyway. It would be faster to not re-run the pipeline and just do an additional join in your analysis.


Some commonly used acronyms for variable naming in the code include:

* `dcrb` dungeness crab
* `rev` revenue
* `lbs` landings
* `VMS` vessel monitoring system
* `afi` adjusted for inflation

```{r define-filters}
# define filters
target_rev <- "DCRB"         # revenue target
target_lbs <- "DCRB"         # landings target
min_depth <- 0               # minimum depth in meters
max_depth <- -150            # maximum depth in meters
min_speed <- 0               # minimum speed in m/s
max_speed <- 4.11556         # maximum speed in m/s (4.11556 m/s = 8 knots)
crab_year_start <- 11        # month defines start of crab year
winter_months <- c("November", "December", "January", "February", "March") # determine Winter or Spring-Summer season
```

```{r transform-ticket-data}
# apply same filters as VMS dataframe
dcrb_ticket_df <- ticket_df %>%
  # add temporal columns
  mutate(
    year_numeric = year(date),
    month_numeric = month(date),
    week_numeric = week(date),
    day_numeric = yday(date),
    month_factor = month(date, label = TRUE, abbr = FALSE),
    year_month_character = paste0(year(date),"_", substr(ymd(date), 6, 7)),
    crab_year_character = ifelse(month_numeric >= crab_year_start, 
                                 paste0(year_numeric, "_", 1+year_numeric),
                                 paste0(year_numeric-1, "_", year_numeric)),
    season_character = as.character(ifelse(month_factor %in% winter_months, "Winter", "Spring-Summer")),
    year_month_date = ym(year_month_character)
  ) %>%
  # apply filters
  filter(TARGET_rev == target_rev | TARGET_lbs == target_lbs) %>%
  # join inflation adjustment factor and adjust revenue
  left_join(fred_pcepi, by = join_by(year_month_date == year_month)) %>%
  mutate(DCRB_revenue_afi = DCRB_revenue * 100 / pcepi) %>%
  # select columns
  dplyr::select(
    # identifiers
    Rec_ID,           # fish ticket ID
    drvid,            # vessel ID
    pacfin_port_code, # port ID
    port_group_code,  # port group ID
    agency_code,      # agency code
    # temporal fields
    date,
    year_month_date,
    crab_year_character,
    year_month_character,
    month_factor,
    year_numeric,
    month_numeric,
    week_numeric,
    day_numeric,
    # vessel length
    FINAL_LENGTH,
    # inflation adjustment index
    pcepi,
    # dungeness crab fields
    DCRB_lbs,
    DCRB_revenue,
    DCRB_revenue_afi
  ) %>%
  # de-duplicate records
  distinct() %>%
  # add in CDFW block info 06-13-2025. have not tried this yet!
  left_join(ticket_df2, by = join_by(Rec_ID == FISH_TICKET_ID))

# take a peek at columns in the resulting dataframe
colnames(dcrb_ticket_df)
```

```{r}
# count total records, trips and years
n_records <- nrow(dcrb_ticket_df)
n_trips   <- n_distinct(dcrb_ticket_df$Rec_ID)
n_years   <- n_distinct(dcrb_ticket_df$year_numeric)
```

The Dungeness crab fish ticket dataframe has `r n_records` records (fish tickets), `r n_trips` distinct trips (fish tickets), across `r n_years` years.

There are more records than trips due to the vessel registration processing code including the lookback window data in step 2 of the pipeline. I looked at some examples of duplicates, and they occur for tickets within the lookback window - for example, a ticket landed on Dec. 28, 2014 is processed in the 2014 and 2015 runs of the pipeline for vessel registration. I suspect this could be fixed in the pipeline itself, but I'm leaving it be for now.

This is fixed here in the plotting code by removing duplicates when one length is NA, and I didn't see any duplication within the VMS data.

```{r}
# identify which records need to be de-duplicated, by finding RecIDs with 2 records, 1 of which has NA length
qa_df <- dcrb_ticket_df %>% 
  group_by(Rec_ID) %>%
  summarize(n = n(),
            na_length = sum(is.na(FINAL_LENGTH))) %>%
  mutate(remove_na = (na_length == 1) & (n == 2)) %>%
  arrange(desc(n))

# take a peek at columns in the resulting dataframe
colnames(qa_df)

# how many records need de-duplication (have 2 records, 1 of which has NA length)?
qa_df %>% group_by(n, na_length, remove_na) %>% summarise(n_records = n()) %>% arrange(desc(n_records)) # 7499 records

# join back to dcrb_ticket_df
qa_dcrb_ticket_df <- dcrb_ticket_df %>%
  left_join(qa_df, by = join_by(Rec_ID)) %>%
  filter(!(remove_na & is.na(FINAL_LENGTH)))

# count total records, trips and years
n_qa_records <- nrow(qa_dcrb_ticket_df)
n_qa_trips   <- n_distinct(qa_dcrb_ticket_df$Rec_ID)
n_qa_years   <- n_distinct(qa_dcrb_ticket_df$year_numeric)
```

The QA'ed Dungeness crab fish ticket dataframe has `r n_qa_records` records (fish tickets), `r n_qa_trips` distinct trips (fish tickets), across `r n_qa_years` years. Now there are no duplicate fish tickets.

# Make plots

VMS point frequency distributions to be split into before and after that ping rate increase, or by year

