---
title: "01_compare_cdfw_block_vms_data"
author: "Jameal Samhouri"
date: "Written beginning 2025-06-05. Last Run `r Sys.Date()`"
output: 
  html_document:
      toc: true
      toc_float: true
geometry: margin=1in
---

# Setup

```{r setup, include=FALSE}

# import libraries
library(tidyverse)
library(here)
library(fredr)
library(magrittr)
library(sf)
library(knitr)
library(lubridate)
library(glue)
library(cowplot)
library(rnaturalearth)
library(viridis)
library(knitr)

options(dplyr.summarise.inform = FALSE)

# change default ggplot theme for plotting
plot_theme <- theme_minimal()+theme(panel.border = element_rect(color='black',fill=NA),
                                    panel.background=element_rect(fill="#abd0d6"),panel.grid=element_line(color='gray70',linewidth = 0.2))
theme_set(plot_theme)

```

```{r create-output-directories}
# make hindcast output directory, if doesn't yet exist
output_dir <- here('Confidential', 'hindcast_output')
if (!dir.exists(output_dir)) dir.create(output_dir)

# create hindcast output subdirectory name based on system date
output_subdir_name <- paste0("hindcast_output_", Sys.Date())

# make hindcast output subdirectory, if doesn't yet exist
output_subdir <- here('Confidential', 'hindcast_output', output_subdir_name)
if (!dir.exists(output_subdir)) dir.create(output_subdir)

# make tables, figures, and maps subdirectories
for (temp_name in c('tables', 'figures')) {
  temp_subdir <- here('Confidential', 'hindcast_output', output_subdir_name, temp_name)
  if (!dir.exists(temp_subdir)) dir.create(temp_subdir)
}
rm(temp_name, temp_subdir)
```

# Introduction

This document qualitatively compares the spatial distribution of CA Dungeness crab fishing activity based on self-reported CDFW block data from fish tickets and geolocations obtained from VMS data. 

It uses coarse aggregations of DCRB fishing activity from both data sources. Specifically, it aggregates fishing records on a 10-arc minute grid and in CDFW small, medium, and large offshore blocks.

The specific analyses conducted here rely upon a summarized version of the VMS record-by-record data frame. The summary is based on trip (equivalent to fish ticket). These analyses include:

1) Frequency histograms of number of trips with 1-n unique blocks indicated by VMS pings for each trip, binned by trips with lower and higher numbers of VMS records

Remember that VMS point frequency distributions need to be split into before and after that ping rate increase, or by year.

2) Annual time series of % of trips with >=1 VMS record in the reported CDFW block.
- small (10-arc minute) blocks
- medium blocks
- large offshore blocks

Remember that VMS point frequency distributions need to be split into before and after that ping rate increase, or by year.

<!-- Blake suggested we try this package: https://link.springer.com/article/10.1007/s10980-024-01930-z  -->

3) Spatial visualizations (maps)

  (a) number of tix per CDFW block vs number of pings per CDFW block
  (b) quantile of # tix per CDFW block vs quantile of # pings per CDFW block vs difference in those qunatiles per CDFW block
    
  for 60-min ping interval years and 15-min ping interval years
  
  <!-- , then subset by whale and non-whale seasons and/or by port group (a la Free et al. 2023).  -->
    
  also make annual maps of 3ab above.
  
  (c) number of tix per RAMP zone vs number of pings per RAMP zone
  (d) quantile of # tix per RAMP zone vs quantile of # pings per RAMP zone vs difference in those qunatiles per RAMP zone
    
 <!-- future: consider # unique vessels per CDFW block based on tix vs # unique vessels per CDFW block based on VMS -->

<!-- - Consider following this vignette to make a correlation map: https://statnmap.com/2018-01-27-spatial-correlation-between-rasters/  -->


Remember that VMS point frequency distributions need to be split into before and after that ping rate increase, or by year.


# Load and join data frames for processing

### Fish tix matched to VMS data

1. Load grid

```{r load_10arcminute_grid}

# 10-arcminute grid shapefile
grd <- read_sf(here('spatial_data', 'grids', 'ten_arcminute_grid_shore_lamb.shp'))
names(grd)

# background map, from rnaturalearth
coast <- ne_states(country='United States of America',returnclass = 'sf') %>%
  filter(name %in% c('California','Oregon','Washington','Nevada')) %>%
  st_transform(crs = st_crs(grd))

grd_dissolved <- grd %>%
  dplyr::select(BLOCK10_ID,State,WM_NGDC_m) %>% 
  group_by(BLOCK10_ID,State) %>% 
  summarise(NGDC_m_mean=mean(WM_NGDC_m)) %>% 
  ungroup()

```

2. Load Matched VMS and fish tix (matched_filtered_withFTID_length_yyyy.rds) 

<!-- Pre-requisite: Run the pipeline steps 1-5 for calendar years 2011-2023. -->

<!-- The code below borrows heavily from Brooke Hawkins' notebook: 03_blh_hindcasts-of-california-dungeness-crab-fishing-activity.Rmd -->



<!-- ```{r load vms_tix} -->

<!-- # choose years of data to load -->
<!-- load_years <- 2011:2023 -->

<!-- # load fish ticket data joined to vessel length data, but before it was joined to VMS data -->
<!-- vms_df <- purrr::map(load_years, function(ly) { -->
<!--   read_rds(here('Confidential', 'processed_data', 'processed_2025-03-19', 'filtered', paste0('matched_filtered_withFTID_length_', ly, '.rds'))) -->
<!-- }) %>% bind_rows() -->

<!-- glimpse(vms_df) -->


<!-- ``` -->

<!-- 3. Join VMS and fish ticket data to grid -->

<!-- identifies the CDFW block in which each VMS record falls -->

<!-- This chunk is inspired by overlay_points_on_grid.Rmd from the grid_overlay_edits branch -->

<!-- ```{r match_points} -->

<!-- pt <- proc.time() -->

<!-- # convert vms to spatial object (longitude/latitude) and join to grid -->
<!-- vms_sf <- vms_df %>% -->
<!--   st_as_sf(coords=c('LON','LAT'),crs=4326) %>%  -->
<!--   # then, convert to planar projection to match the grid -->
<!--   st_transform(st_crs(grd)) %>% -->
<!--   st_join(grd)  -->

<!-- vms_grid10_df <- vms_sf %>% -->
<!--   st_set_geometry(NULL) -->

<!-- proc.time()-pt -->

<!-- # 55s for 2022 data -->
<!-- # 764 s for 2011-2023 data -->

<!-- rm(vms_df) -->
<!-- rm(vms_sf) -->

<!-- ``` -->


<!-- 4. Load inflation adjustment factors from FRED. Use annual instead of monthly -->

<!-- Adjust revenue for inflation using Fred GDP data, adapted from R code by Erin Steiner. This product uses the FRED® API but is not endorsed or certified by the Federal Reserve Bank of St. Louis. -->

<!-- Pre-requisite: Create a FRED account and API key. See https://fred.stlouisfed.org/docs/api/api_key.html. -->

<!-- ```{r inflation-adjustment} -->
<!-- # insert your FRED API key -->
<!-- fredr_set_key('ea19d3d447eb4dc773db00c5a2e61f22')  -->

<!-- # download the quarterly inflation adjustments from Fred from 1985 to present -->
<!-- fred_gdpdefl <- fredr( -->
<!--     series_id = "GDPDEF",  -->
<!--     observation_start = as.Date(paste0(min(load_years)-1, "-01-01")) -->
<!--   ) -->

<!-- # generate the mean annual deflators based on the quarterly values -->
<!-- gdp_defl <- mutate(fred_gdpdefl, year = year(date)) %>% -->
<!--   group_by(year) %>% -->
<!--   summarize(defl = mean(value), .groups = 'drop') %>% -->
<!--   mutate(inflation_adjustment_factor = defl / defl[year == max(year)]) %>% -->
<!--   select(year, inflation_adjustment_factor) -->

<!-- # write inflation adjustment factors -->
<!-- write.csv(file = here("Confidential", "hindcast_output", "tables", "annual_inflation_adjustment.csv"), x = gdp_defl) -->

<!-- # plot inflation adjustment factors -->
<!-- gdp_defl %>% ggplot(aes(x = as.factor(year), y = inflation_adjustment_factor)) + geom_point() -->
<!-- ggsave(here("Confidential", "hindcast_output", "figures", "annual_inflation_adjustment.png")) -->

<!-- # clean up -->
<!-- rm(fred_gdpdefl) -->
<!-- ``` -->

<!-- #### Transform data -->

<!-- ```{r define-filters-vms_df} -->
<!-- # define filters -->
<!-- target_rev <- "DCRB"         # revenue target -->
<!-- target_lbs <- "DCRB"         # landings target -->
<!-- state_agency_code <- "C"     # landings brought to California ports -->
<!-- min_depth <- 0               # minimum depth in meters -->
<!-- max_depth <- -150            # maximum depth in meters -->
<!-- min_speed <- 0               # minimum speed in m/s -->
<!-- max_speed <- 4.11556         # maximum speed in m/s (4.11556 m/s = 8 knots) -->
<!-- crab_year_start <- 11        # month defines start of crab year -->
<!-- winter_months <- c("November", "December", "January", "February", "March") # determine Winter or Spring-Summer season -->
<!-- ``` -->

<!-- ```{r transform_vms_data} -->

<!-- # transform VMS and fish ticket data -->
<!-- dcrb_df <- vms_grid10_df %>% -->
<!--   # use distinct to avoid duplicates caused by vessel length matching step (do not keep rows that are the same except for "FINAL_LENGTH") -->
<!--   distinct( -->
<!--     VMS_RECNO, Rec_ID, FTID, pacfin_port_code, port_group_code, dealerID, drvid, agency_code, date, TARGET_lbs, all_species_lbs, DCRB_lbs, DCRB_lbs_no_revenue, TARGET_rev, all_species_rev, DCRB_revenue, species_code_all, X_COORD, Y_COORD,  westcoastdate_notime, avg_speed_recalc, BLOCK10_ID, -->
<!--      .keep_all = TRUE -->
<!--   ) %>% -->
<!--   # add temporal columns -->
<!--   mutate( -->
<!--     year = lubridate::year(westcoastdate_notime), -->
<!--     month = lubridate::month(westcoastdate_notime, label=TRUE, abbr=FALSE), -->
<!--     month_numeric = month(westcoastdate_notime), -->
<!--     year_month = paste0(lubridate::year(westcoastdate_notime),"_", substr(lubridate::ymd(westcoastdate_notime), 6, 7)), -->
<!--     crab_year = ifelse(month_numeric >= crab_year_start, paste0(year, "_", 1+year), paste0(year-1, "_", year)), -->
<!--     season = as.character(ifelse(month %in% winter_months, "Winter", "Spring-Summer")), -->
<!--     week_of_year = week(westcoastdate_notime), -->
<!--     day_of_year = yday(westcoastdate_notime), -->
<!--     # year_month_character = paste0(year(date),"_", substr(ymd(date), 6, 7)), -->
<!--     year_month_date = floor_date(ymd(westcoastdate_notime), 'month') -->
<!--   ) %>% -->
<!--   # apply filters -->
<!--   filter(TARGET_rev == target_rev | TARGET_lbs == target_lbs) %>% -->
<!--   filter(agency_code == state_agency_code) %>% -->
<!--   filter(NGDC_M <= min_depth & NGDC_M >= max_depth) %>% -->
<!--   filter(avg_speed_recalc <= max_speed & avg_speed_recalc >= min_speed) %>%  -->
<!--   # join inflation adjustment factor and adjust revenue -->
<!--   left_join(gdp_defl, by = join_by(year == year)) %>% -->
<!--   mutate(DCRB_revenue_afi = DCRB_revenue * 100 / inflation_adjustment_factor) %>% #at this point, 14,418,569 rows have been reduced to 1,077,651 rows -->
<!--   # select columns -->
<!--   # drop "gear_code_all", "end", "prev_ticket_dttm", "lookback_dttm", "begin", "trip_dur", "UTCDATETIME", "VESSEL_NAME", "AVG_SPEED", "AVG_COURSE", "DECLARATION_CODE", "westcoastdate",  "has_vms", "multispecies", "segment_dur",      "in_port", "Shoreline", "row_index", "col_index"  -->
<!--   dplyr::select( -->
<!--     # identifiers -->
<!--     Rec_ID,           # fish ticket ID -->
<!--     FTID, -->
<!--     VMS_RECNO,        # VMS ping ID -->
<!--     drvid,            # vessel ID -->
<!--     agency_code, -->
<!--     pacfin_port_code, # port ID -->
<!--     port_group_code,  # port group ID -->
<!--     BLOCK10_ID,       # grid cell ID -->
<!--     BLOCK10_n, -->
<!--     Offshrblok, -->
<!--     Adj_cell_1,  -->
<!--     Adj_cell_2,  -->
<!--     Adj_cell_3,  -->
<!--     Adj_Cell_4,  -->
<!--     Adj_Cell_5,  -->
<!--     Adj_Cell_6,  -->
<!--     Adj_Cell_7,  -->
<!--     Adj_Cell_8, -->
<!--     Area_m2, -->
<!--     WM_NGDC_m, -->
<!--     # temporal fields -->
<!--     date, -->
<!--     westcoastdate, -->
<!--     westcoastdate_notime, -->
<!--     crab_year,  -->
<!--     season, -->
<!--     year, -->
<!--     year_month, -->
<!--     month,  -->
<!--     month_numeric, -->
<!--     week_of_year, -->
<!--     day_of_year, -->
<!--     # vessel length -->
<!--     FINAL_LENGTH, -->
<!--     # dungeness crab fields -->
<!--     DCRB_lbs, -->
<!--     DCRB_revenue_afi -->
<!--   ) %>% -->
<!--    left_join( -->
<!--     dcrb_ticket_df %>% dplyr::select(date, crab_year=crab_year_character, FTID, CDFW_AREA_BLOCK) # changed from year=year_numeric 08-21-2025 to avoid problem that sum(is.na(dcrb_df$CDFW_AREA_BLOCK)) = 25343. also tried crab_year=crab_year_character, which reduced the sum of NAs to 170 (but relates to only 3 fish tix). however, there remained a many to many join problem so i added date and drvid to the join. that cleared up the the many to many join issue but increased the NA problem to 3444. so then i dropped drvid to see what would happen, and that the sum of NAs to 227 but only 5 unique fish tix. call it good here -->
<!--     ) -->

<!-- glimpse(dcrb_df) -->

<!-- dim(distinct(dcrb_df)) -->

<!-- #length(which(is.na(dcrb_df$CDFW_AREA_BLOCK))) #  -->


<!-- ``` -->



<!-- ```{r write dcrb_vms_grd} -->

<!-- # write rds for non-spatial version -->
<!-- write_rds(dcrb_df, here( -->
<!--   'Confidential',  -->
<!--   'processed_data',  -->
<!--   'vms_fish_tickets', -->
<!--   paste0( -->
<!--     'dcrb_vms_w_grd_10_arcmin_', -->
<!--     paste0(load_years[1],'-',load_years[length(load_years)]), -->
<!--     '_',Sys.Date(),'.rds' -->
<!--     ) -->
<!-- ) -->
<!-- ) -->



<!-- ``` -->

## shortcut load

```{r load_dcrb_vms}

load_years <- 2011:2023

dcrb_df <- read_rds(here(
  'Confidential', 
  'processed_data', 
  'vms_fish_tickets',
  paste0(
    'dcrb_vms_w_grd_10_arcmin_',
    paste0(load_years[1],'-',load_years[length(load_years)]),
    '_2025-08-21.rds'
    )
  )
)

# if dcrb_df was developed before Blake added RAMP zone attributes to grd on 09-19-2025, add those in here
dcrb_df <- grd %>%
  st_set_geometry(NULL) %>%
  dplyr::select(BLOCK10_ID, EEZ, State, Salish_Sea, BLOCK10_n, Offshrblok, RAMP_zone, RAMP_over, WM_NGDC_m) %>%
  right_join(dcrb_df)


```

# Analyze matched VMS and fish tix

Drop the 5 trips where we were unable to include a CDFW_AREA_BLOCK 

```{r drop_noblksinfishtix}
#View(dcrb_df[which(is.na(dcrb_df$CDFW_AREA_BLOCK)),]) 
# unique(dcrb_df$FTID[which(is.na(dcrb_df$CDFW_AREA_BLOCK))])
# [1] "O415366" "D579106" "143713E" "229032E" "328806E"

dcrb_df <- dcrb_df[-which(is.na(dcrb_df$CDFW_AREA_BLOCK)),]

```

Add before/after ping interval change to df

```{r add_pinginterval}

dcrb_df <- dcrb_df %>%
  mutate(
    vms_ping_interval_period = ifelse(westcoastdate_notime < '2020-09-01',"60-minute ping intervals","15-minute ping intervals")
  )

```

## 1) Frequency histograms of number of trips with 1-n unique blocks indicated by VMS pings for each trip, binned by trips with lower and higher numbers of VMS records

Remember that VMS point frequency distributions need to be split into before and after that ping rate increase, or by year.

Make summarized data frames

```{r blks_per_trip_df}

BLOCK10_ID_per_trip <- dcrb_df %>% # dcrb_df[10000:12000,]
  group_by(Rec_ID, vms_ping_interval_period) %>%
  summarise(
    total_BLOCK10_ID = length(unique(BLOCK10_ID)),
    total_vms_rec = length(unique(VMS_RECNO))
  )

glimpse(BLOCK10_ID_per_trip)

single_v_multiple_BLOCK10_ID_per_trip <- BLOCK10_ID_per_trip %>%
         mutate(
           single_v_multiple_BLOCK10_ID = ifelse(total_BLOCK10_ID > 1, 'multiple', 'single')
           ) %>%
  group_by(vms_ping_interval_period, single_v_multiple_BLOCK10_ID) %>%
  summarise(
    tot_trips = n_distinct(Rec_ID)
  )

single_v_multiple_BLOCK10_ID_per_trip$single_v_multiple_BLOCK10_ID <- factor(single_v_multiple_BLOCK10_ID_per_trip$single_v_multiple_BLOCK10_ID, levels = c('single','multiple'))

glimpse(single_v_multiple_BLOCK10_ID_per_trip)

#Remember that VMS point frequency distributions need to be split into before and after that ping rate increase, or by year.


```

Make 4 plots:
1) absolute frequency histogram of blocks visited by ping interval period
2) proportional frequency histogram of blocks visited by ping interval period
3) single or multiple blocks visited v number of trips by ping interval period
4) blocks visited v pings per trip by ping interval period

```{r blks_per_trip_plots}

ggplot(data = BLOCK10_ID_per_trip) +
  facet_wrap(~vms_ping_interval_period) + 
  geom_histogram(aes(x = total_BLOCK10_ID), binwidth=2) +
  labs(x='Number of BLOCK10_IDs Visited',y="Number of Trips") +
  theme_classic()

ggsave(
  filename = here("Confidential", "hindcast_output", output_subdir_name, "figures", "absolute frequency histogram of blocks visited by ping interval period.png"), 
  width = 8, height = 5
  )

ggplot(data = BLOCK10_ID_per_trip) +
  facet_wrap(~vms_ping_interval_period) + 
  geom_histogram(aes(x = total_BLOCK10_ID, y = after_stat(density)), binwidth=2) +
  labs(x='Number of BLOCK10_IDs Visited',y="Proportion of Trips") +
  theme_classic()

ggsave(
  filename = here("Confidential", "hindcast_output", output_subdir_name, "figures", "proportional frequency histogram of blocks visited by ping interval period.png"), 
  width = 8, height = 5
  )

ggplot(data = single_v_multiple_BLOCK10_ID_per_trip) +
  facet_wrap(~vms_ping_interval_period) + 
  geom_col(aes(x = single_v_multiple_BLOCK10_ID, y = tot_trips)) +
  labs(x='BLOCK10_IDs Visited', y="Number of Trips") +
  theme_classic()

ggsave(
  filename = here("Confidential", "hindcast_output", output_subdir_name, "figures", "single or multiple blocks visited v number of trips by ping interval period.png"), 
  width = 8, height = 5
  )

ggplot(data = BLOCK10_ID_per_trip) +
  facet_wrap(~vms_ping_interval_period) + 
  geom_point(
    aes(x = total_vms_rec, y = total_BLOCK10_ID)
    ) +
  labs(y='BLOCK10_IDs Visited', x="Number of Pings Per Trip") +
  theme_classic()

ggsave(
  filename = here("Confidential", "hindcast_output", output_subdir_name, "figures", "blocks visited v pings per trip by ping interval period.png"), 
  width = 8, height = 5
  )


```


RESULTS: 

Both during the 60-minute ping interval period and the 15-minute ping interval period: 

(i) Most trips transit through multiple BLOCK10_IDs.
(ii) The most common number of BLOCK10_IDs visited is in the range of 2-10.
(iii) There is a weak positive correlation between the number of BLOCK10_IDs visited and the number of pings per trip.

## 2) Annual time series of % of trips with >=1 VMS record in the reported CDFW block.
- small (10-arc minute) blocks
- medium blocks
- large offshore blocks

Remember that VMS point frequency distributions need to be split into before and after that ping rate increase, or by year.

Goal is to determine correspondence between reported and tracked fishing locations.


Define the medium and offshore blocks. 
```{r define_blks}

offshore_blk_IDs <- 1032:1042
medium_blk_IDs <- c(138:139, 240:241, 280:281, 484:487, 545:546, 568, 649:650, 697, 776:777, 895:897)

```

Make data frames for small, medium, and large blocks reported.

```{r make_blk_size_dfs}

dcrb_df_sm <- dcrb_df %>%
  filter(!CDFW_AREA_BLOCK %in% offshore_blk_IDs) %>%
  filter(!CDFW_AREA_BLOCK %in% medium_blk_IDs)

dcrb_df_med <- dcrb_df %>%
  filter(CDFW_AREA_BLOCK %in% medium_blk_IDs)

dcrb_df_lg <- dcrb_df %>%
  filter(CDFW_AREA_BLOCK %in% offshore_blk_IDs)

```

```{r count_trips}
# count total trips for each blk size
n_smblk_trips <- length(unique(dcrb_df_sm$Rec_ID))
n_medblk_trips <- length(unique(dcrb_df_med$Rec_ID))
n_lgblk_trips <- length(unique(dcrb_df_lg$Rec_ID))

```

The Dungeness crab fish ticket dataframe has `r n_smblk_trips` trips (fish tickets) reporting small blocks, `r n_medblk_trips` trips (fish tickets) reporting medium blocks, and `r n_lgblk_trips` trips (fish tickets) reporting large blocks.


Quantify % of pings that match reported block, and % of tix where at least one ping matches reported block

For now, I am not accounting for individual vessel level effects, but instead looking across all trips and all vessels all at once

```{r match_ping_reported}

dcrb_df_sm_match_long <- dcrb_df_sm %>%
  mutate(
    match_ping_report = case_when(
      BLOCK10_ID == CDFW_AREA_BLOCK ~ 1,
      BLOCK10_ID != CDFW_AREA_BLOCK ~ 0
    ),
    block_type = 'small'
  )

dcrb_df_med_match_long <- dcrb_df_med %>%
  mutate(
    match_ping_report = case_when(
      BLOCK10_ID == CDFW_AREA_BLOCK ~ 1,
      BLOCK10_ID != CDFW_AREA_BLOCK ~ 0
    ),
    block_type = 'medium'
  )

dcrb_df_lg_match_long <- dcrb_df_lg %>%
  mutate(
    match_ping_report = case_when(
      Offshrblok == CDFW_AREA_BLOCK ~ 1,
      Offshrblok != CDFW_AREA_BLOCK ~ 0
    ),
    block_type = 'large'
  )

dcrb_df_match_long <- dcrb_df_sm_match_long %>%
  bind_rows(dcrb_df_med_match_long) %>%
  bind_rows(dcrb_df_lg_match_long)

dcrb_df_match_all_fishtix <- dcrb_df_match_long %>%
  # summarise by each fish ticket
  group_by( 
    block_type, crab_year, vms_ping_interval_period, Rec_ID, FTID, drvid, agency_code, pacfin_port_code, port_group_code, FINAL_LENGTH #, DCRB_lbs, DCRB_revenue_afi. commented out DCRB_revenue_afi for now because for tix with vms records that span multiple months we end up with >1 row for each unique ticket. fix later
  ) %>%
  summarise(
    total_match_ping_report = sum(match_ping_report, na.rm=TRUE),
    num_vms_records = n(),
    match_ping_report_YN = ifelse(total_match_ping_report >=1, 1, 0),
    prop_match_ping_report = total_match_ping_report / num_vms_records
  ) %>%
  ungroup()

dcrb_df_match_all_crabyrs <- dcrb_df_match_all_fishtix %>%
  # summarise across all fish tix
  group_by(
    block_type, crab_year, vms_ping_interval_period
  ) %>%
  summarise(
    prop_match_ping_report_YN = sum(match_ping_report_YN)/n(),
    mean_prop_match_ping_report = mean(prop_match_ping_report, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(
    year_numeric = substr(crab_year,6,9)
  ) %>%
  filter(year_numeric != 2024)



```

### Make % match plots

Plot % of tickets with at least one match between reported and tracked blocks - include both reported offshore blocks and reported block10s

```{r low_bar_plot_sm}

dcrb_df_match_all_crabyrs %>%
  mutate(block_type = fct_relevel(block_type, "small", "medium", "large")) %>%
  ggplot(
    aes(
        x = year_numeric, y = prop_match_ping_report_YN
        )
  ) +
  geom_point(aes(shape = vms_ping_interval_period, colour = block_type, group = block_type)) +
  geom_line(aes(colour = block_type, group = block_type)) +
  # facet_wrap(~vms_ping_interval_period) + 
  ylim(0,1) +
  labs(
    y = 'Proportion of trips with >=1 VMS record\nmatching reported CDFW 10-arc minute block',
    x = 'Crab year',
    colour = 'Block size',
    shape = ' VMS ping interval period'
       ) #+
 # theme_bw()

ggsave(filename = here("Confidential", "hindcast_output", output_subdir_name, "figures", "time_series_prop_dcrb_tix_match_ping_report_all_byblksize.png"), width = 8, height = 6)

```

RESULTS: 

Both during the 60-minute ping interval period and the 15-minute ping interval period: 

i) Almost 100% of trips reporting catch from large blocks show at least 1 VMS record in the reported large block.
ii) ~50$ of trips reporting catch from small blocks show at least 1 VMS record in the reported small block. This percentage seems to be rising from <50% in 2011-2012 to ~60% in 2022-2023.
iii) None of the trips reporting catch from medium blocks show at least 1 VMS record in the reported medium block. (Jameal thinks this is a real issue)

## 3) Spatial visualizations (maps)

Summarize data frame for mapping

  (a) number of tix per CDFW block vs number of pings per CDFW block
  (b) quantile of # tix per CDFW block vs quantile of # pings per CDFW blockvs difference in those qunatiles per CDFW block
    
  for 60-min ping interval years and 15-min ping interval years,
  
  <!-- then subset by whale and non-whale seasons and/or by port group (a la Free et al. 2023).  -->
  
  also make summaries for each fishing month (Nov-July) separately, across all years within each ping-interval period, for 3ab above.

Focus on fish tix with self-reported 10-arc minute (small) blocks and associated VMS data

  (c) number of tix per RAMP zone vs number of pings per RAMP zone
  (d) quantile of # tix per RAMP zone vs quantile of # pings per RAMP zone vs difference in those qunatiles per RAMP zone

Use Owen's code as a basis. https://github.com/noaa-nwfsc/VMS-pipeline/blob/7fac4c72968f6235ad1f73975128821cf3ddd933/code/notebooks/owen/summary%20mapping%20dcrb.qmd 

  

```{r sum_for_maps}

# filter to CDFW small block tix only
# group_by by crab_year and BLOCK10_ID
# sum num Rec_ID reporting each block, num VMS_RECNO in each block
# repeat faceted by pacfin_port_code, season if desired

#offshore_blk_IDs <- 1032:1042
medium_blk_IDs <- c(138:139, 240:241, 280:281, 484:487, 545:546, 568, 649:650, 697, 776:777, 895:897)

dcrb_df_block10 <- dcrb_df %>%
  # mutate(
  #   offshore_reported = case_when(
  #     CDFW_AREA_BLOCK %in% offshore_blk_IDs ~ 1,
  #     !CDFW_AREA_BLOCK %in% offshore_blk_IDs ~ 0#,
  #     #.default = 9999
  #   )
  # ) %>%
  # filter(offshore_reported == 0) %>%
  filter(BLOCK10_ID < 1000) %>%
  filter(!BLOCK10_ID %in% medium_blk_IDs)

# function to aggregate. make sure not to compare post 2020 high ping rate to pre 2020
summarize_dcrb_attribute_block10 <- function(att,temporal_breaks){
  
  # first, establish the attribute to be summarized, and select it from the dataset
  if(att=="trips") df <- dcrb_df_block10 %>%
      rename(attribute=Rec_ID)
  if(att=="pings") df <- dcrb_df_block10 %>%
      rename(attribute=VMS_RECNO)
  
  # then, group the data into the appropriate temporal breaks and summarize based on the chosen measure
  # if the option is all years, sum across all months of the crab year, separating pre Sep 2020 from post due to an increase in ping rate from 60-mn intervals to 15-min intervals
  if(temporal_breaks=="all years") {
    df2 <- df %>% 
      filter(month_numeric %in% c(1,2,3,4,5,6,7,11,12)) %>% 
      mutate(temporal_break=ifelse(westcoastdate_notime < '2020-09-01',"60-minute ping intervals","15-minute ping intervals")) %>% 
      group_by(temporal_break,crab_year,BLOCK10_ID, RAMP_zone, RAMP_over) %>% 
      summarise(attr_n=n_distinct(attribute,na.rm=T)) %>% 
      group_by(temporal_break,BLOCK10_ID, RAMP_zone, RAMP_over) %>% 
      # calculate mean and sd
      summarise(
        mean_allyrs=mean(attr_n,na.rm=T),
        sd_allyrs=sd(attr_n,na.rm=T),
      ) %>% 
      # calculate rank and quantile
      ungroup() %>% 
      mutate(ranked=cume_dist(mean_allyrs)) %>% 
      arrange(desc(ranked)) %>%
      mutate(cume=cumsum(mean_allyrs)) %>%
      mutate(quant=cume/max(cume)) %>%
      # binned into 5% chunks
      mutate(quantbinned=ntile(quant,10)*10) %>%
      mutate(quantbinned=ordered(quantbinned)) %>% 
      dplyr::select(-cume)
  }
  
    if(temporal_breaks=="month across years") {
    df2 <- df %>%
      filter(month_numeric %in% c(1,2,3,4,5,6,7,11,12)) %>%
      mutate(temporal_break=ifelse(westcoastdate_notime < '2020-09-01',paste(month.name[month_numeric], "60-minute ping intervals"), paste(month.name[month_numeric], "15-minute ping intervals"))
             ) %>%
      group_by(temporal_break, month_numeric, year, BLOCK10_ID, RAMP_zone, RAMP_over) %>%
      summarise(attr_n=n_distinct(attribute,na.rm=T)) %>%
      group_by(temporal_break, month_numeric, BLOCK10_ID, RAMP_zone, RAMP_over) %>%
      # calculate mean and sd
      # note that some sd_allyrs values may be NA because there are not multiple months within a temporal_break with values for attr_n
      summarise(
        mean_allyrs=mean(attr_n,na.rm=T),
        sd_allyrs=sd(attr_n,na.rm=T),
      ) %>% 
      # calculate rank and quantile
      ungroup() %>% 
      mutate(ranked=cume_dist(mean_allyrs)) %>% 
      arrange(desc(ranked)) %>%
      mutate(cume=cumsum(mean_allyrs)) %>%
      mutate(quant=cume/max(cume)) %>%
      # binned into 5% chunks
      mutate(quantbinned=ntile(quant,10)*10) %>% 
      mutate(quantbinned=ordered(quantbinned)) %>%
      dplyr::select(-cume)
  }
  
  #df_out <- df2
  
  # finally, join the spatial data
  df_out <- df2 %>%
    left_join(grd_dissolved,by="BLOCK10_ID") %>%
    #left_join(grd,by="BLOCK10_ID") %>% 
    st_as_sf()
  
  df_out
}

```


## Compare metrics

### Summarize data frames across all years and all months, and for each month across all years

Remember that quantile means “count up this cell and all those larger than it, and see how much of the total sum we have accumulated”. If, say, a grid cell has a value for quant of 0.05 for pings, this would mean that it is in the first 5% of all grid cells when we are counting up total pings from largest to smallest.

```{r make_vms_blk_map_df}

trips_allyrs <- summarize_dcrb_attribute_block10('trips', 'all years') %>% mutate(metric = 'trips') %>% 
  rename(quant_trips = quant, mean_allyrs_trips = mean_allyrs) %>% 
  dplyr::select(-sd_allyrs,-ranked,-quantbinned,-metric) %>% 
  mutate(
    RAMP_zone_whales = case_when(
      RAMP_zone == 'Zone 2 & Zone 7' ~ 'Zone 2',
      RAMP_zone == 'Zone 3 & Zone 7' ~ 'Zone 3',
      RAMP_zone == 'Zone 4 & Zone 7' ~ 'Zone 4',
      TRUE ~ RAMP_zone
      )
    )

pings_allyrs <- summarize_dcrb_attribute_block10('pings', 'all years') %>% mutate(metric = 'pings') %>% 
  rename(quant_pings = quant, mean_allyrs_pings = mean_allyrs) %>% 
  dplyr::select(-sd_allyrs,-ranked,-quantbinned,-metric) %>% 
  mutate(
    RAMP_zone_whales = case_when(
      RAMP_zone == 'Zone 2 & Zone 7' ~ 'Zone 2',
      RAMP_zone == 'Zone 3 & Zone 7' ~ 'Zone 3',
      RAMP_zone == 'Zone 4 & Zone 7' ~ 'Zone 4',
      TRUE ~ RAMP_zone
      )
    )

# combine all yrs df's into a single one, adding a column for the difference between quantiles

trips_pings_allyrs_full <- trips_allyrs %>%
  st_join(pings_allyrs,
          suffix = c("_trips", "_pings"),
          join = st_intersects
          ) 

common_col_names <- c('temporal_break', 'BLOCK10_ID', 'RAMP_zone', 'RAMP_over', 'State', 'NGDC_m_mean')

logical_checks <- map(
  .x = common_col_names,
  .f = ~ trips_pings_allyrs_full[[paste0(., "_trips")]] == trips_pings_allyrs_full[[paste0(., "_pings")]]
)

final_filter_condition <- reduce(logical_checks, `&`)

trips_pings_allyrs <- trips_pings_allyrs_full %>%
  filter(final_filter_condition) %>%
  mutate(
    diff_quants_t_minus_p = quant_trips - quant_pings
  ) %>% 
  mutate(
    RAMP_zone_whales = case_when(
      RAMP_zone_pings == 'Zone 2 & Zone 7' ~ 'Zone 2',
      RAMP_zone_pings == 'Zone 3 & Zone 7' ~ 'Zone 3',
      RAMP_zone_pings == 'Zone 4 & Zone 7' ~ 'Zone 4',
      TRUE ~ RAMP_zone_pings
      )
    )

# now focus on months across years

trips_monthsallyrs <- summarize_dcrb_attribute_block10('trips', 'month across years') %>% mutate(metric = 'trips') %>% 
  rename(quant_trips = quant, mean_allyrs_trips = mean_allyrs) %>% 
  dplyr::select(-sd_allyrs,-ranked,-quantbinned,-metric) %>% 
  mutate(
    RAMP_zone_whales = case_when(
      RAMP_zone == 'Zone 2 & Zone 7' ~ 'Zone 2',
      RAMP_zone == 'Zone 3 & Zone 7' ~ 'Zone 3',
      RAMP_zone == 'Zone 4 & Zone 7' ~ 'Zone 4',
      TRUE ~ RAMP_zone
      )
    )

pings_monthsallyrs <- summarize_dcrb_attribute_block10('pings', 'month across years') %>% mutate(metric = 'pings') %>% 
  rename(quant_pings = quant, mean_allyrs_pings = mean_allyrs) %>% 
  dplyr::select(-sd_allyrs,-ranked,-quantbinned,-metric) %>% 
  mutate(
    RAMP_zone_whales = case_when(
      RAMP_zone == 'Zone 2 & Zone 7' ~ 'Zone 2',
      RAMP_zone == 'Zone 3 & Zone 7' ~ 'Zone 3',
      RAMP_zone == 'Zone 4 & Zone 7' ~ 'Zone 4',
      TRUE ~ RAMP_zone
      )
    )



```


### Example comparisons between fish tix and VMS records

#### Correlations

```{r make_corr_plots}

# trips_allyrs <- summarize_dcrb_attribute_block10('trips', 'all years') %>% mutate(metric = 'trips') %>% rename(quant_trips = quant, mean_allyrs_trips = mean_allyrs)
# pings_allyrs <- summarize_dcrb_attribute_block10('pings', 'all years') %>% mutate(metric = 'pings') %>% rename(quant_pings = quant, mean_allyrs_pings = mean_allyrs)
# 
# trips_allyrs %>% st_join(pings_allyrs) %>%
#   ggplot()+
#   geom_point(aes(x=quant_trips, y=quant_pings))+
#   theme(panel.background=element_rect(fill='white'))
# 
# pings_allyrs %>% st_join(trips_allyrs) %>%
#   ggplot()+
#   geom_point(aes(x=mean_allyrs_trips, y=mean_allyrs_pings))+
#   theme(panel.background=element_rect(fill='white'))

trips_allyrs <- summarize_dcrb_attribute_block10('trips', 'all years') %>% mutate(metric = 'trips') %>% 
  rename(quant_trips = quant, mean_allyrs_trips = mean_allyrs) %>% 
  dplyr::select(-sd_allyrs,-ranked,-quantbinned,-metric) %>% 
  mutate(
    RAMP_zone_whales = case_when(
      RAMP_zone == 'Zone 2 & Zone 7' ~ 'Zone 2',
      RAMP_zone == 'Zone 3 & Zone 7' ~ 'Zone 3',
      RAMP_zone == 'Zone 4 & Zone 7' ~ 'Zone 4',
      TRUE ~ RAMP_zone
      )
    )

pings_allyrs <- summarize_dcrb_attribute_block10('pings', 'all years') %>% mutate(metric = 'pings') %>% 
  rename(quant_pings = quant, mean_allyrs_pings = mean_allyrs) %>% 
  dplyr::select(-sd_allyrs,-ranked,-quantbinned,-metric) %>% 
  mutate(
    RAMP_zone_whales = case_when(
      RAMP_zone == 'Zone 2 & Zone 7' ~ 'Zone 2',
      RAMP_zone == 'Zone 3 & Zone 7' ~ 'Zone 3',
      RAMP_zone == 'Zone 4 & Zone 7' ~ 'Zone 4',
      TRUE ~ RAMP_zone
      )
    )

trips_allyrs %>% 
  # non-spatial join
  st_set_geometry(NULL) %>% 
  left_join(pings_allyrs %>% st_set_geometry(NULL)) %>% 
  # st_join(pings_allyrs) %>%
  ggplot()+
  geom_point(aes(x=quant_trips, y=quant_pings))+
  theme(panel.background=element_rect(fill='white'))

ggsave(filename = here("Confidential", "hindcast_output", output_subdir_name, "figures", "correlation_quant_trips_quant_pings.png"), width = 8, height = 6)

trips_allyrs %>% 
  # non-spatial join
  st_set_geometry(NULL) %>% 
  left_join(pings_allyrs %>% st_set_geometry(NULL)) %>% 
  # st_join(pings_allyrs) %>%
  ggplot()+
  geom_point(aes(x=mean_allyrs_trips, y=mean_allyrs_pings))+
  theme(panel.background=element_rect(fill='white'))

ggsave(filename = here("Confidential", "hindcast_output", output_subdir_name, "figures", "correlation_mean_allyrs_trips_mean_allyrs_pings.png"), width = 8, height = 6)

# pings_allyrs %>% st_join(trips_allyrs) %>%
#   ggplot()+
#   geom_point(aes(x=mean_allyrs_trips, y=mean_allyrs_pings))+
#   theme(panel.background=element_rect(fill='white'))

# looking better

```

#### Distribution of quantile differences

```{r make_plot_quant_diffs}

ggplot(trips_pings_allyrs, aes(x = diff_quants_t_minus_p)) +
  geom_histogram(binwidth = 0.05, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(
    title = "Frequency Distribution of Quantile Differences",
    subtitle = "(Trips - Pings)",
    x = "Quantile Difference",
    y = "Frequency"
  ) +
  theme_minimal()

ggsave(filename = here("Confidential", "hindcast_output", output_subdir_name, "figures", "hist_quant_diffs_trips_pings_allyrs.png"), width = 8, height = 6)

ggplot(trips_pings_allyrs, aes(x = diff_quants_t_minus_p)) +
  geom_histogram(binwidth = 0.05, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(
    title = "Frequency Distribution of Quantile Differences",
    subtitle = "(Trips - Pings)",
    x = "Quantile Difference",
    y = "Frequency"
  ) +
  facet_wrap(~temporal_break_trips) +
  theme_minimal()

ggsave(filename = here("Confidential", "hindcast_output", output_subdir_name, "figures", "hist_quant_diffs_trips_pings_allyrs_bypinginterval.png"), width = 8, height = 6)

```

### Maps

#### Get set up to make maps.

We can filter the large 10-arc minute grid to only encompass the cells with some crab fishing effort. Do this with a filter-buffer-crop workflow (could also use, e.g., a distance from shore or depth cutoff):

```{r dcrb_cells}
# filter the overall grid using grid cells with DCRB effort in any month/year
cells_we_need <- unique(dcrb_df$BLOCK10_ID)
grid10_filt <- grd %>% filter(BLOCK10_ID %in% cells_we_need)

# buffer by 50km
grid10_filt_buff <- grid10_filt %>% summarise() %>% st_buffer(1e4)

# crop to make the final grid
grid10_final <- grd %>% st_filter(grid10_filt_buff)
bbox <- st_bbox(grid10_final)

# map this
ggplot()+
  geom_sf(data=grid10_final,aes(fill=NGDC_med_m),col=NA)+
  geom_sf(data=coast,fill='#dacf80',col='gray80')+
  xlim(bbox[1],bbox[3])+ylim(bbox[2],bbox[4])+
  scale_fill_viridis(option="turbo",direction=-1)
```

We make a spatially rotated grid, for better visualization when mapping the whole coastline. We use custom projections for this, which B. Feist designed. Based on that work, here we provide generic code for any given amount of rotation.

```{r rotate_map}
# this is the first wkt file ('well-known text') that B. Feist created, containing the info on the custom spatial projection
rotation_wkt <- read_file(here('spatial_data','map_rotation','Rectified_Skew_Orthomorphic_center_30deg_rotation.wkt'))

# working in this simple text format, we can find and replace to change the rotation of the map
crs_rotated <- function(deg){
  crs_out <- sub('Skew Grid\",30',paste0('Skew Grid\",',deg),rotation_wkt)
  crs_out
}

# let's use this to make a 15 deg rotation, which seems about right for our study domain
rot15 <- crs_rotated(15) %>% st_crs()

# use this crs to reprojec the 5k grid
grid10_rot15 <- grid10_final %>% st_transform(rot15)
bbox_rot15 <- st_bbox(grid10_rot15) +c(0,0,-20000,0)

# reproject for ca only
grid10_rot15_ca <- grid10_final %>% filter(State == 'CA') %>% st_transform(rot15)
bbox_rot15_ca <- st_bbox(grid10_rot15 %>% filter(State == 'CA')) +c(0,0,-20000,0)

# map
ggplot()+
  geom_sf(data=grid10_rot15, aes(fill=NGDC_med_m),col=NA)+
  geom_sf(data=coast,fill='#dacf80',col='gray80')+
  xlim(bbox_rot15[1],bbox_rot15[3])+ylim(bbox_rot15[2],bbox_rot15[4])+
  scale_fill_viridis(option="turbo",direction=-1)

# map ca only
ggplot()+
  geom_sf(data=grid10_rot15_ca, aes(fill=NGDC_med_m),col=NA)+
  geom_sf(data=coast,fill='#dacf80',col='gray80')+
  xlim(bbox_rot15_ca[1],bbox_rot15_ca[3])+ylim(bbox_rot15_ca[2],bbox_rot15_ca[4])+
  scale_fill_viridis(option="turbo",direction=-1)

```

start adding mapping code, see line 808 of Owen's make_map_coastwide()

```{r make_map_function}

make_map_ca <- function(df, column,scl_ord=1,...) {
  
  cols_to_check <- c("State", "State_trips", "State_pings")
  
  dat2 <- df %>% filter(if_any(any_of(cols_to_check), ~ . == 'CA')) %>% st_transform(rot15)
  
  #is the value a factor (i.e., categorical?)
  is_factor <- df %>% pull( {{column}} ) %>% is.factor()
  
  # ca map
  if(is_factor){
    ggplot() + 
      geom_sf(data=dat2,aes(fill= {{column}}),col=NA)+
      geom_sf(data=coast,fill='#dacf80',col='gray80')+
      xlim(bbox_rot15[1],bbox_rot15[3])+
      ylim(bbox_rot15[2],bbox_rot15[4])+
      scale_fill_viridis_d(na.value='#abd0d6',option='turbo',direction=scl_ord,...)+
      theme(panel.background=element_rect(fill="#abd0d6"),panel.grid=element_line(color='gray70',linewidth = 0.2))
  } else{
    ggplot() + 
      geom_sf(data=dat2,aes(fill= {{column}}),col=NA)+
      geom_sf(data=coast,fill='#dacf80',col='gray80')+
      xlim(bbox_rot15[1],bbox_rot15[3])+
      ylim(bbox_rot15[2],bbox_rot15[4])+
      scale_fill_viridis(option='turbo',na.value='#abd0d6',direction=scl_ord,...)+
      theme(panel.background=element_rect(fill="#abd0d6"),panel.grid=element_line(color='gray70',linewidth = 0.2))
  }
  
}

make_map_ca_sqrttransform <- function(df, column,scl_ord=1,...) {
  
  dat2 <- df %>% filter(State == 'CA') %>% st_transform(rot15)
  
  #is the value a factor (i.e., categorical?)
  is_factor <- df %>% pull( {{column}} ) %>% is.factor()
  
  # ca map
  if(is_factor){
    ggplot() + 
      geom_sf(data=dat2,aes(fill= {{column}}),col=NA)+
      geom_sf(data=coast,fill='#dacf80',col='gray80')+
      xlim(bbox_rot15_ca[1],bbox_rot15_ca[3])+
      ylim(bbox_rot15_ca[2],bbox_rot15_ca[4])+
      scale_fill_viridis_d(na.value='#abd0d6',option='turbo',direction=scl_ord,...)+
      theme(panel.background=element_rect(fill="#abd0d6"),panel.grid=element_line(color='gray70',linewidth = 0.2))
  } else{
    ggplot() + 
      geom_sf(data=dat2,aes(fill= sqrt({{column}})),col=NA)+
      geom_sf(data=coast,fill='#dacf80',col='gray80')+
      xlim(bbox_rot15_ca[1],bbox_rot15_ca[3])+
      ylim(bbox_rot15_ca[2],bbox_rot15_ca[4])+
      scale_fill_viridis(option='turbo',na.value='#abd0d6',direction=scl_ord,...)+
      theme(panel.background=element_rect(fill="#abd0d6"),panel.grid=element_line(color='gray70',linewidth = 0.2))
  }
  
}

# just plotting NGDC_med_m (depth) as the fill to make sure everything is oriented correctly
# make_map_coastwide(grid10_final, NGDC_med_m)+labs(fill="dept")

make_map_ca(grid10_final, NGDC_med_m)+labs(fill="dept")

# make_map_coastwide(grid5k_final, centro_lat,scl_ord = -1)+labs(fill="latitude")

```

Map functions for >=2 levels for temporal_breaks, continuous variables

```{r make_map_function_multi}

make_map_ca_multi <- function(df, column,scl_ord=1,...) {
  
  dat2 <- df %>% filter(State == 'CA') %>% st_transform(rot15)
  
  # scale limits for the color bar, to ensure they are common across subplots
  lims <- dat2 %>% pull( {{column}} ) %>% range()
  
  # make the list of maps, with the list members named by the unique values in "temporal breaks"
  plotnames <- unique(dat2$temporal_break)
  
  plotlist <- map(plotnames,function(x){
    
    dfsub <- dat2 %>% filter(temporal_break==x)
    
    # ca map
    p1 <- make_map_ca(dfsub, {{column}},scl_ord,limits=lims) + labs(subtitle = x)
    
    p1
    
    })
  
  return(plotlist %>% set_names(plotnames))
}

make_map_ca_sqrttransform_multi <- function(df, column,scl_ord=1,...) {
  
  dat2 <- df %>% filter(State == 'CA') %>% st_transform(rot15)
  
  # scale limits for the color bar, to ensure they are common across subplots
  lims <- sqrt(dat2 %>% pull( {{column}} ) %>% range())
  
  # make the list of maps, with the list members named by the unique values in "temporal breaks"
  plotnames <- unique(dat2$temporal_break)
  
  plotlist <- map(plotnames,function(x){
    
    dfsub <- dat2 %>% filter(temporal_break==x)
    
    # ca map
    p1 <- make_map_ca_sqrttransform(dfsub, {{column}},scl_ord,limits=lims) + labs(subtitle = x)
    
    p1
    
    })
  
  return(plotlist %>% set_names(plotnames))
}



```

#### Now make maps

Subset to CA only

```{r ca-only}

trips_allyrs_ca <- trips_allyrs %>%
  filter(State == 'CA')

pings_allyrs_ca <- pings_allyrs %>%
  filter(State == 'CA')

trips_pings_allyrs_ca <- trips_pings_allyrs %>%
  filter(State_trips == 'CA') %>%
  filter(State_pings == 'CA')

trips_monthsallyrs_ca <- trips_monthsallyrs %>%
  filter(State == 'CA')

pings_monthsallyrs_ca <- pings_monthsallyrs %>%
  filter(State == 'CA')

```

##### 1. 60-minute ping interval years

###### Compare means across all years - CDFW blocks
```{r 60_min_maps}

trips_60minpinginterval <- make_map_ca_sqrttransform(trips_allyrs_ca %>% filter(temporal_break == '60-minute ping intervals'), mean_allyrs_trips)+labs(fill="sqrt mean\nnumber of trips", title = 'CDFW small block-reported\nfishing locations', subtitle = 'Nov 2011 - Sep 2020\n 60-minute ping intervals')

trips_60minpinginterval

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "60-minute ping intervals reported fishing locations.png"))


pings_60minpinginterval <- make_map_ca_sqrttransform(pings_allyrs_ca %>% filter(temporal_break == '60-minute ping intervals'), mean_allyrs_pings)+labs(fill="sqrt mean\nnumber of pings", title = 'VMS-recorded\nfishing locations', subtitle = 'Nov 2011 - Sep 2020\n 60-minute ping intervals')

pings_60minpinginterval

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "60-minute ping intervals vms-recorded fishing locations.png"))

plot_grid(trips_60minpinginterval, pings_60minpinginterval,
          labels = "auto"
          )

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "60-minute ping intervals reported and vms-recorded fishing locations.png"))

```

###### Compare quantiles across all years - CDFW blocks
```{r 60_min_maps_quantiles}

quantile_trips_60minpinginterval <- make_map_ca(trips_allyrs_ca %>% filter(temporal_break == '60-minute ping intervals'), quant_trips)+labs(fill="quantile number of trips", title = 'CDFW small block-reported\nfishing locations', subtitle = 'Nov 2011 - Sep 2020\n 60-minute ping intervals')

quantile_trips_60minpinginterval

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "60-minute ping intervals quantiles of reported fishing locations.png"))


quantile_pings_60minpinginterval <- make_map_ca(pings_allyrs_ca %>% filter(temporal_break == '60-minute ping intervals'), quant_pings)+labs(fill="quantile number of pings", title = 'VMS-recorded\nfishing locations', subtitle = 'Nov 2011 - Sep 2020\n 60-minute ping intervals')

quantile_pings_60minpinginterval

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "60-minute ping intervals quantiles of vms-recorded fishing locations.png"))

diff_quants_t_minus_p_60minpinginterval <- make_map_ca(trips_pings_allyrs_ca %>% filter(temporal_break_trips == '60-minute ping intervals') %>% filter(temporal_break_pings == '60-minute ping intervals'), diff_quants_t_minus_p)+labs(fill="Quantile difference\n(trips - pings)", title = 'Difference between\nCDFW small block-reported\nand VMS-recorded\nfishing locations', subtitle = 'Nov 2011 - Sep 2020\n 60-minute ping intervals')

diff_quants_t_minus_p_60minpinginterval

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "60-minute ping intervals quantile differences (reported trips minus vms-recorded fishing locations).png"))

plot_grid(quantile_trips_60minpinginterval, quantile_pings_60minpinginterval,diff_quants_t_minus_p_60minpinginterval,
          labels = "auto", nrow=1
          )

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "60-minute ping intervals quantiles and quantile differences of reported and vms-recorded fishing locations.png"))

```

###### Compare means across all years - RAMP zones
```{r 60_min_maps_ramp}

trips_60minpinginterval_ramp <- make_map_ca_sqrttransform(
  trips_allyrs_ca %>% 
    filter(temporal_break == '60-minute ping intervals') %>%
    group_by(RAMP_zone_whales, State) %>%
    summarise(
      mean_allyrs_trips = mean(mean_allyrs_trips)
    ),
  mean_allyrs_trips) + 
  labs(fill="sqrt mean\nnumber of trips", title = 'CDFW small block-reported\nfishing locations\nby RAMP zone', subtitle = 'Nov 2011 - Sep 2020\n 60-minute ping intervals')

trips_60minpinginterval_ramp

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "60-minute ping intervals reported fishing locations by RAMP zone.png"))


pings_60minpinginterval_ramp <- make_map_ca_sqrttransform(
  pings_allyrs_ca %>% 
    filter(temporal_break == '60-minute ping intervals') %>%
    group_by(RAMP_zone_whales, State) %>%
    summarise(
      mean_allyrs_pings = mean(mean_allyrs_pings)
    ), 
  mean_allyrs_pings) +
  labs(fill="sqrt mean\nnumber of pings", title = 'VMS-recorded\nfishing locations\nby RAMP zone', subtitle = 'Nov 2011 - Sep 2020\n 60-minute ping intervals')

pings_60minpinginterval_ramp

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "60-minute ping intervals vms-recorded fishing locations by RAMP zone.png"))

plot_grid(trips_60minpinginterval_ramp, pings_60minpinginterval_ramp,
          labels = "auto"
          )

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "60-minute ping intervals reported and vms-recorded fishing locations by RAMP zone.png"))

```

###### Compare quantiles across all years - RAMP zones
```{r 60_min_maps_quantiles_ramp}

quantile_trips_60minpinginterval_ramp <- make_map_ca(
  trips_allyrs_ca %>% 
    filter(temporal_break == '60-minute ping intervals') %>%
    group_by(RAMP_zone_whales, State) %>%
    summarise(
      quant_trips = mean(quant_trips)
    ), 
  quant_trips) + 
  labs(fill="quantile number of trips", title = 'CDFW small block-reported\nfishing locations\nby RAMP zone', subtitle = 'Nov 2011 - Sep 2020\n 60-minute ping intervals')

quantile_trips_60minpinginterval_ramp

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "60-minute ping intervals quantiles of reported fishing locations by RAMP zone.png"))


quantile_pings_60minpinginterval_ramp <- make_map_ca(
  pings_allyrs_ca %>% 
    filter(temporal_break == '60-minute ping intervals') %>%
    group_by(RAMP_zone_whales, State) %>%
    summarise(
      quant_pings = mean(quant_pings)
    ),
  quant_pings) + 
  labs(fill="quantile number of pings", title = 'VMS-recorded\nfishing locations\nby RAMP zone', subtitle = 'Nov 2011 - Sep 2020\n 60-minute ping intervals')

quantile_pings_60minpinginterval_ramp

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "60-minute ping intervals quantiles of vms-recorded fishing locations by RAMP zone.png"))

diff_quants_t_minus_p_60minpinginterval_ramp <- make_map_ca(
  trips_pings_allyrs_ca %>%
    filter(temporal_break_trips == '60-minute ping intervals') %>% 
    filter(temporal_break_pings == '60-minute ping intervals') %>%
    group_by(RAMP_zone_whales, State_trips) %>%
    summarise(
      diff_quants_t_minus_p = mean(diff_quants_t_minus_p)
    ), 
  diff_quants_t_minus_p) + 
  labs(fill="Quantile difference\n(trips - pings)", title = 'Difference between\nCDFW small block-reported\nand VMS-recorded\nfishing locations\nby RAMP zone', subtitle = 'Nov 2011 - Sep 2020\n 60-minute ping intervals')

diff_quants_t_minus_p_60minpinginterval_ramp

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "60-minute ping intervals quantile differences (reported trips minus vms-recorded fishing locations) by RAMP zone.png"))

plot_grid(quantile_trips_60minpinginterval_ramp, quantile_pings_60minpinginterval_ramp,diff_quants_t_minus_p_60minpinginterval_ramp,
          labels = "auto", nrow=1
          )

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "60-minute ping intervals quantiles and quantile differences of reported and vms-recorded fishing locations by RAMP zone.png"))

```

##### 2. 15-minute ping interval years


###### Compare means across all years - CDFW blocks
```{r 15_min_maps}

trips_15minpinginterval <- make_map_ca_sqrttransform(trips_allyrs_ca %>% filter(temporal_break == '15-minute ping intervals'), mean_allyrs_trips)+labs(fill="sqrt mean\nnumber of trips", title = 'CDFW small block-reported\nfishing locations', subtitle = 'Sep 2020 - Dec 2023\n 15-minute ping intervals')

trips_15minpinginterval

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "15-minute ping intervals reported fishing locations.png"))

pings_15minpinginterval <- make_map_ca_sqrttransform(pings_allyrs_ca %>% filter(temporal_break == '15-minute ping intervals'), mean_allyrs_pings)+labs(fill="sqrt mean\nnumber of pings", title = 'VMS-recorded\nfishing locations', subtitle = 'Sep 2020 - Dec 2023\n 15-minute ping intervals')

pings_15minpinginterval

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "15-minute ping intervals vms-recorded fishing locations.png"))

plot_grid(trips_15minpinginterval, pings_15minpinginterval,
          labels = "auto"
          )

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "15-minute ping intervals reported and vms-recorded fishing locations.png"))

```

###### Compare quantiles across all years - CDFW blocks
```{r 15_min_maps_quantiles}

quantile_trips_15minpinginterval <- make_map_ca(trips_allyrs_ca %>% filter(temporal_break == '15-minute ping intervals'), quant_trips)+labs(fill="quantile number of trips", title = 'CDFW small block-reported\nfishing locations', subtitle = 'Sep 2020 - Dec 2023\n 15-minute ping intervals')

quantile_trips_15minpinginterval

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "15-minute ping intervals quantiles of reported fishing locations.png"))


quantile_pings_15minpinginterval <- make_map_ca(pings_allyrs_ca %>% filter(temporal_break == '15-minute ping intervals'), quant_pings)+labs(fill="quantile number of pings", title = 'VMS-recorded\nfishing locations', subtitle = 'Sep 2020 - Dec 2023\n 15-minute ping intervals')

quantile_pings_15minpinginterval

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "15-minute ping intervals quantiles of vms-recorded fishing locations.png"))

diff_quants_t_minus_p_15minpinginterval <- make_map_ca(trips_pings_allyrs_ca %>% filter(temporal_break_trips == '15-minute ping intervals') %>% filter(temporal_break_pings == '15-minute ping intervals'), diff_quants_t_minus_p)+labs(fill="Quantile difference\n(trips - pings)", title = 'Difference between\nCDFW small block-reported\nand VMS-recorded\nfishing locations', subtitle = 'Sep 2020 - Dec 2023\n 15-minute ping intervals')

diff_quants_t_minus_p_15minpinginterval

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "15-minute ping intervals quantile differences (reported trips minus vms-recorded fishing locations).png"))

plot_grid(quantile_trips_15minpinginterval, quantile_pings_15minpinginterval,diff_quants_t_minus_p_15minpinginterval,
          labels = "auto",nrow=1
          )

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "15-minute ping intervals quantiles and quantile differences of reported and vms-recorded fishing locations.png"))

```
###### Compare means across all years - RAMP zones

```{r 15_min_maps_ramp}

trips_15minpinginterval_ramp <- make_map_ca_sqrttransform(
  trips_allyrs_ca %>% 
    filter(temporal_break == '15-minute ping intervals') %>%
    group_by(RAMP_zone_whales, State) %>%
    summarise(
      mean_allyrs_trips = mean(mean_allyrs_trips)
    ),
  mean_allyrs_trips) + 
  labs(fill="sqrt mean\nnumber of trips", title = 'CDFW small block-reported\nfishing locations\nby RAMP zone', subtitle = 'Sep 2020 - Dec 2023\n 15-minute ping intervals')

trips_15minpinginterval_ramp

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "15-minute ping intervals reported fishing locations by RAMP zone.png"))


pings_15minpinginterval_ramp <- make_map_ca_sqrttransform(
  pings_allyrs_ca %>% 
    filter(temporal_break == '15-minute ping intervals') %>%
    group_by(RAMP_zone_whales, State) %>%
    summarise(
      mean_allyrs_pings = mean(mean_allyrs_pings)
    ), 
  mean_allyrs_pings) +
  labs(fill="sqrt mean\nnumber of pings", title = 'VMS-recorded\nfishing locations\nby RAMP zone', subtitle = 'Sep 2020 - Dec 2023\n 15-minute ping intervals')

pings_15minpinginterval_ramp

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "15-minute ping intervals vms-recorded fishing locations by RAMP zone.png"))

plot_grid(trips_15minpinginterval_ramp, pings_15minpinginterval_ramp,
          labels = "auto"
          )

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "15-minute ping intervals reported and vms-recorded fishing locations by RAMP zone.png"))

```

###### Compare quantiles across all years - RAMP zones
```{r 15_min_maps_quantiles_ramp}

quantile_trips_15minpinginterval_ramp <- make_map_ca(
  trips_allyrs_ca %>% 
    filter(temporal_break == '15-minute ping intervals') %>%
    group_by(RAMP_zone_whales, State) %>%
    summarise(
      quant_trips = mean(quant_trips)
    ), 
  quant_trips) + 
  labs(fill="quantile number of trips", title = 'CDFW small block-reported\nfishing locations\nby RAMP zone', subtitle = 'Sep 2020 - Dec 2023\n 15-minute ping intervals')

quantile_trips_15minpinginterval_ramp

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "15-minute ping intervals quantiles of reported fishing locations by RAMP zone.png"))


quantile_pings_15minpinginterval_ramp <- make_map_ca(
  pings_allyrs_ca %>% 
    filter(temporal_break == '15-minute ping intervals') %>%
    group_by(RAMP_zone_whales, State) %>%
    summarise(
      quant_pings = mean(quant_pings)
    ),
  quant_pings) + 
  labs(fill="quantile number of pings", title = 'VMS-recorded\nfishing locations\nby RAMP zone', subtitle = 'Sep 2020 - Dec 2023\n 15-minute ping intervals')

quantile_pings_15minpinginterval_ramp

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "15-minute ping intervals quantiles of vms-recorded fishing locations by RAMP zone.png"))

diff_quants_t_minus_p_15minpinginterval_ramp <- make_map_ca(
  trips_pings_allyrs_ca %>%
    filter(temporal_break_trips == '15-minute ping intervals') %>% 
    filter(temporal_break_pings == '15-minute ping intervals') %>%
    group_by(RAMP_zone_whales, State_trips) %>%
    summarise(
      diff_quants_t_minus_p = mean(diff_quants_t_minus_p)
    ), 
  diff_quants_t_minus_p) + 
  labs(fill="Quantile difference\n(trips - pings)", title = 'Difference between\nCDFW small block-reported\nand VMS-recorded\nfishing locations\nby RAMP zone', subtitle = 'Sep 2020 - Dec 2023\n 15-minute ping intervals')

diff_quants_t_minus_p_15minpinginterval_ramp

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "15-minute ping intervals quantile differences (reported trips minus vms-recorded fishing locations) by RAMP zone.png"))

plot_grid(quantile_trips_15minpinginterval_ramp, quantile_pings_15minpinginterval_ramp,diff_quants_t_minus_p_15minpinginterval_ramp,
          labels = "auto", nrow=1
          )

ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "15-minute ping intervals quantiles and quantile differences of reported and vms-recorded fishing locations by RAMP zone.png"),
       width = 10, height = 8)

```

##### 3. Comparisons for each month across all years

###### Compare means - CDFW blocks 
```{r maps_month}

trips_minpinginterval_months <- make_map_ca_sqrttransform_multi(trips_monthsallyrs_ca, mean_allyrs_trips)

trips_minpinginterval_months

# Use iwalk() to save each plot with automatic naming
iwalk(trips_minpinginterval_months,
      ~ {
  # Add a title using the name of the list element (.y)
  final_plot <- .x + labs(fill="sqrt mean\nnumber of trips", title = 'CDFW small block-reported\nfishing locations')
  
  ggsave(filename = here("Confidential", "hindcast_output", output_subdir_name, "figures", paste0(.y, " - trips.png")), plot = final_plot)
  }
  )

pings_minpinginterval_months <- make_map_ca_sqrttransform_multi(pings_monthsallyrs_ca, mean_allyrs_pings)

pings_minpinginterval_months

# Use iwalk() to save each plot with automatic naming
iwalk(pings_minpinginterval_months,
      ~ {
  # Add a title using the name of the list element (.y)
  final_plot <- .x + labs(fill="sqrt mean\nnumber of pings", title = 'VMS-recorded\nfishing locations')
  
  ggsave(filename = here("Confidential", "hindcast_output", output_subdir_name, "figures", paste0(.y, " - pings.png")), plot = final_plot)
  }
  )

```

###### Compare quantiles - CDFW blocks 
```{r maps_quantiles_month}

quantile_trips_pinginterval_months <- make_map_ca_multi(trips_monthsallyrs_ca, quant_trips)

quantile_trips_pinginterval_months

# Use iwalk() to save each plot with automatic naming
iwalk(quantile_trips_pinginterval_months,
      ~ {
  # Add a title using the name of the list element (.y)
  final_plot <- .x + labs(fill="quantile number of trips", title = 'CDFW small block-reported\nfishing locations')
  
  ggsave(filename = here("Confidential", "hindcast_output", output_subdir_name, "figures", paste0(.y, " - quantile trips.png")), plot = final_plot)
  }
  )

quantile_pings_pinginterval_months <- make_map_ca_multi(pings_monthsallyrs_ca, quant_pings)

quantile_pings_pinginterval_months

# Use iwalk() to save each plot with automatic naming
iwalk(quantile_pings_pinginterval_months,
      ~ {
  # Add a title using the name of the list element (.y)
  final_plot <- .x + labs(fill="quantile number of pings", title = 'VMS-recorded\nfishing locations')
  
  ggsave(filename = here("Confidential", "hindcast_output", output_subdir_name, "figures", paste0(.y, " - quantile pings.png")), plot = final_plot)
  }
  )


```
