---
title: "01_compare_cdfw_block_vms_data"
author: "Jameal Samhouri"
date: "Written beginning 2025-06-05. Last Run `r Sys.Date()`"
output: 
  html_document:
      toc: true
      toc_float: true
geometry: margin=1in
---

# Setup

```{r setup, include=FALSE}

# import libraries
library(tidyverse)
library(here)
library(fredr)
library(magrittr)
library(sf)
library(knitr)
library(lubridate)
library(glue)
library(cowplot)
library(rnaturalearth)
library(viridis)
library(knitr)

options(dplyr.summarise.inform = FALSE)

# change default ggplot theme for plotting
plot_theme <- theme_minimal()+theme(panel.border = element_rect(color='black',fill=NA),
                                    panel.background=element_rect(fill="#abd0d6"),panel.grid=element_line(color='gray70',linewidth = 0.2))
theme_set(plot_theme)

```

```{r create-output-directories}
# make hindcast output directory, if doesn't yet exist
output_dir <- here('Confidential', 'hindcast_output')
if (!dir.exists(output_dir)) dir.create(output_dir)

# create hindcast output subdirectory name based on system date
output_subdir_name <- paste0("hindcast_output_", Sys.Date())

# make hindcast output subdirectory, if doesn't yet exist
output_subdir <- here('Confidential', 'hindcast_output', output_subdir_name)
if (!dir.exists(output_subdir)) dir.create(output_subdir)

# make tables, figures, and maps subdirectories
for (temp_name in c('tables', 'figures')) {
  temp_subdir <- here('Confidential', 'hindcast_output', output_subdir_name, temp_name)
  if (!dir.exists(temp_subdir)) dir.create(temp_subdir)
}
rm(temp_name, temp_subdir)
```

# Introduction

This document compares the spatial distribution of CA Dungeness crab fishing activity based on self-reported CDFW block data from fish tickets and geolocations obtained from VMS data.

1) Using the VMS record-by-record df, calculate distance of VMS point from centroid or edge of CDFW block reported in fish ticket; 
- Make plot of frequency distribution showing distance of VMS point from CDFW block, with 50/75/90% labeled 

2) Summarize VMS record-by-record df by trip_ID, and calculate % of VMS points within 0/1/2-degree neighboring blocks across time periods. Try this package: https://link.springer.com/article/10.1007/s10980-024-01930-z 
- 0-degree neighbor = inside/outside of CDFW block
- Make plot of % of VMS points (y-axis) and degree neighbor (x-axis), each point estimate representing a time period (eg, yr)
- Facet by region, vessel size, etc
- Consider summarizing by vessel first

3) Spatial visualizations (maps)
- Summarize VMS record-by-record df by trip_ID and CDFW block
    - # tix per CDFW block vs # pings/day per CDFW block
    - # unique vessels per CDFW block based on tix vs # unique vessels/day per CDFW block based on VMS
- Follow this vignette to make a correlation map: https://statnmap.com/2018-01-27-spatial-correlation-between-rasters/ 


Reminder that VMS point frequency distributions need to be split into before and after that ping rate increase, or by year


# Load and join data frames for processing

I have not yet integrated DCRB permit information.
eg,

NWFSC_DSACDFW_DcrabPermitDetail2013-2023_08072023.xlsx
TrapTierNumberKey.csv

## Load data

### Fish tix, no VMS records

1. Load the fish ticket data for specified years, which will be used to grab self-reported fishing locations.

*** Note you can skip to the chunk load_dcrb_tix and avoid running chunks 3-18 

These data can be joined to VMS data

The code below borrows heavily from Brooke Hawkins' notebook "10_blh_dungeness-crab-time-series.Rmd" and the 01_process_fish_tickets.Rmd


```{r load-data}
# choose years of data to load
load_years <- 2011:2023

# the ticket df from the vms pipeline is missing some columns, so load the raw fish ticket data, downloaded by Brooke Hawkins
# From Brooke: Many columns are dropped in step 1 when the fish tickets are cleaned, and that drop propagates through the rest of the pipeline. The fastest workaround is to join the raw fish tickets with the fish ticket ID and get it from there.
# 06-11-2025 changed from read_csv to read.csv to avoid warning messages associated with the HILLE_PERMIT column in at least the 2013 data file. that didnt work. so on 06-12-2025, decided to just read in columns i need. that didnt work either and brooke addressed this issue more comprehensively here: https://github.com/noaa-nwfsc/VMS-pipeline/blob/main/code/notebooks/brooke/16_blh_check-new-fish-tickets.md

# 06-17-2025. made rds so that we can read the df in faster
# ticket_df2 <- purrr::map(load_years, function(ly) {
#   read_csv(here('Confidential', 'raw_data', 'fish_tickets', paste0('fish_tickets_', ly, '.csv')), #) %>%
#     #dplyr::select(FISH_TICKET_ID, CDFW_AREA_BLOCK)
#   #}) %>% bind_rows()
#            col_types = list(
#              HILLE_PERMIT = "c",
#              DEALER_NUM = "c",
#              FTID = "c",
#              NUM_OF_FISH = "n",
#              EFP_CODE = "c",
#              EFP_NAME = "c",
#              GF_PERMIT_NUMBER = "c",
#              TRIP_SEQ = "i",
#              GEAR_CODE = "d",
#              HSFCA_PERMIT = "c" # JS added this specification in addition to those BH recommended
#              )
#   )
#   }) %>% bind_rows() 
# 
# write_rds(ticket_df2, here('Confidential', 'processed_data', 'fish_tickets',
#                            paste0('fish tickets_',
#                                   paste0(load_years[1],'-',load_years[length(load_years)]),
#                                   '.rds')
#                            )
# )

# read in tickets as rds
ticket_df2 <- read_rds(here('Confidential', 'processed_data', 'fish_tickets',
                            paste0('fish tickets_',
                                   paste0(load_years[1],'-',load_years[length(load_years)]),
                                   '.rds')
)
)

# check to see how many fish tickets are in each df. they don't match. this remains a problem. 06-16-2025
# BH made this and things look ok between and old and new fish ticket pull. https://github.com/noaa-nwfsc/VMS-pipeline/blob/main/code/notebooks/brooke/16_blh_check-new-fish-tickets.qmd

# ticket_df2 %>% 
#   # remove missing vessel ID
#   filter(VESSEL_NUM != "UNKNOWN") %>%
#   filter(VESSEL_NUM != "") %>%
#   filter(vessel_id != "MISSING") %>%
#   # remove tickets landed in AK and transported to WC
#   filter(COUNCIL_CODE != "N") %>%
#   # remove tribal fishing
#   filter(FLEET_CODE != "TI") %>%
#   # remove aquaculture
#   filter(PARTICIPATION_GROUP_CODE != "A") %>%
#   group_by(LANDING_YEAR) %>% 
#   summarise(n_trips = n_distinct(FISH_TICKET_ID))

```

<!-- 2. Load inflation adjustment factors from FRED. -->

<!-- Adjust revenue for inflation using Personal Consumption Expenditures: Chain-Type Price Index, [PCEPI](https://fred.stlouisfed.org/series/PCEPI) from FRED. PCEPI index is chosen per Dan Holland's suggestion. The reference year is 2017 by default. This is a monthly time series that is used by multiplying the revenue by 100 / PCEPI. It is used at a monthly time scale (rather than taking an annual average). -->

<!-- Pre-requisite: Create a FRED account and [API key](https://fred.stlouisfed.org/docs/api/api_key.html). This product uses the FRED® API but is not endorsed or certified by the Federal Reserve Bank of St. Louis.  -->

<!-- ```{r inflation-adjustment} -->
<!-- # insert your FRED API key -->
<!-- fredr_set_key('ea19d3d447eb4dc773db00c5a2e61f22') -->

<!-- # download the monthly inflation adjustments from FRED from start of min load year to end of max load year -->
<!-- fred_pcepi <- fredr( -->
<!--     series_id = "PCEPI",  -->
<!--     observation_start = as.Date(paste0(min(load_years), "-01-01")), -->
<!--     observation_end = as.Date(paste0(max(load_years), "-12-31")) -->
<!--   ) %>% -->
<!--   rename(year_month = date, pcepi = value) %>%  -->
<!--   select(-c(realtime_start, realtime_end, series_id)) -->

<!-- # write inflation adjustment factors -->
<!-- write.csv(file = here("Confidential", "hindcast_output", output_subdir_name, "tables", "inflation_pcepi.csv"), x = fred_pcepi, row.names = FALSE) -->

<!-- # plot inflation adjustment factors -->
<!-- fred_pcepi %>% ggplot(aes(x = year_month, y = pcepi)) + geom_line() + scale_x_date(date_breaks = "1 year", date_labels = "%b %Y") -->
<!-- ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "inflation_pcepi.png"), create.dir = TRUE) -->
<!-- ``` -->

3. Clean up fish ticket data

```{r cleaning}

ticket_df2_clean <- ticket_df2 %>% 
  #filter(year == 2011) %>% # this line was used for troubleshooting many-to-many join
  # remove missing vessel ID
  filter(VESSEL_NUM != "UNKNOWN") %>%
  filter(VESSEL_NUM != "") %>%
  filter(VESSEL_NUM != "MISSING") %>%
  # remove tickets landed in AK and transported to WC
  filter(COUNCIL_CODE != "N") %>%
  # remove tribal fishing
  filter(FLEET_CODE != "TI") %>%
  # remove aquaculture
  filter(PARTICIPATION_GROUP_CODE != "A") %>%
  dplyr::select(
    FISH_TICKET_ID,
    FTID,
    PACFIN_PORT_CODE,
    PACFIN_GROUP_PORT_CODE,
    DEALER_ID,
    VESSEL_NUM,
    AGENCY_CODE,
    CDFW_AREA_BLOCK,
    GEAR_CODE,
    GEAR_NAME,
    PACFIN_GROUP_GEAR_CODE,
    REMOVAL_TYPE_CODE,
    REMOVAL_TYPE_NAME,
    LANDING_DATE,
    LANDING_MONTH,
    LANDING_YEAR,
    NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE,
    LANDED_WEIGHT_LBS,
    EXVESSEL_REVENUE
  ) %>%
  # change some column names
  set_colnames(c(
    "Rec_ID",
    "FTID",
    "pacfin_port_code",
    "port_group_code",
    "dealerID",
    "drvid",
    "agency_code",
    "CDFW_AREA_BLOCK",
    "gear_code",
    "gear_name",
    "gear_group",
    "removal_type_code",
    "removal_type_name",
    "date",
    "month",
    "year",
    "NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE",
    "LANDED_WEIGHT_LBS",
    "EXVESSEL_REVENUE"
  )) %>%
  # Adjust gear group codes for some uncategorized gears
  mutate(gear_group=case_when(
    gear_group=='MSC' & gear_name %in% c('SPEAR','DIVING - ABALONE IRON','DIVING - RAKE/HOOKS SEA URCHINS','DIVING', 'SHELLFISH DIVER') ~ 'DVG',
    gear_group=='MSC' & gear_name %in% c('UNKNOWN','UNKNOWN OR UNSPECIFIED GEAR') ~ 'USP',
    gear_group=='MSC' & gear_name %in% c('AQUACULTURE FARM','OYSTER FARM','CLAM FARM') ~ 'FRM',
      TRUE ~ gear_group
  )) 
  
  # mutate(date=as.Date(LANDING_DATE,"%d-%b-%y")) %>%
  
  # dplyr::select(FISH_TICKET_ID, FTID, date, VESSEL_NUM, AGENCY_CODE, CDFW_AREA_BLOCK) %>%
  # distinct()

dim(ticket_df2_clean)

# the code below was used for troubleshooting many-to-many join
# which(ticket_df2_clean$FTID == ticket_df_clean[3,]$FTID)
# View(ticket_df2_clean[which(ticket_df2_clean$FTID == ticket_df_clean[3,]$FTID),])

# ticket_df_clean[which(ticket_df_clean$Rec_ID == ticket_df2_clean[123140,]$FISH_TICKET_ID),]
# length(is.na(ticket_df_clean$FINAL_LENGTH))
# View(ticket_df_clean[which(is.na(ticket_df_clean$FINAL_LENGTH))[1:50],])

# dim(ticket_df %>% filter(year == 2015) %>% distinct(Rec_ID, FTID, date, drvid, agency_code))
# dim(ticket_df2 %>% filter(LANDING_YEAR == 2015) %>% distinct(FISH_TICKET_ID, FTID, LANDING_DATE, VESSEL_NUM, AGENCY_CODE))

```

#### Process fish tickets

##### Concatenate gear / catch information

Concatenate all species information for the fish ticket.

```{r concatenate_species}
# all.species <- ticket_df2_clean %>%
#   group_by(Rec_ID, removal_type_name) %>%
#   summarise(species_code_all = ifelse(length(unique(NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE)) > 1, paste(unique(NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE), collapse="/"), as.character(unique(NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE))))
all.species <- ticket_df2_clean %>%
  group_by(Rec_ID) %>%
  summarise(
    species_code_all = ifelse(
      length(unique(NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE)) > 1,
      paste(unique(NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE), collapse = "/"),
      as.character(unique(NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE))
    )
  ) %>%
  ungroup()
```

Concatenate the gear information for the fish ticket

```{r concatenate_gear}
# gear.info <- ticket_df2_clean %>%
#   group_by(Rec_ID, removal_type_name) %>%
#   summarise(gear_name_all = ifelse(length(unique(gear_name)) > 1, paste(unique(gear_name), collapse="/"), as.character(unique(gear_name))),
#             gear_code_all = ifelse(length(unique(gear_name)) > 1, paste(unique(gear_code), collapse="/"), as.character(unique(gear_code))))

gear.info <- ticket_df2_clean %>%
  group_by(Rec_ID) %>%
  summarise(
    gear_name_all = ifelse(
      length(unique(gear_name)) > 1,
      paste(unique(gear_name), collapse = "/"),
      as.character(unique(gear_name))
    ),
    gear_code_all = ifelse(
      length(unique(gear_name)) > 1,
      paste(unique(gear_code), collapse = "/"),
      as.character(unique(gear_code))
    )
  ) %>%
  ungroup()
```

##### Find Target Species

We need to define the target species for each landed ticket. We will do this by finding the species with the greatest landed revenue for each trip.

Right now, each row of the data is a landing amount for a particular gear/ticket/species combo. We want to collapse these tickets in order to just have one row for each ticket, with an associated amount of landings for the TARGET species.

```{r plot_recordsperticket}
ticket_df2_clean %>%
  count(Rec_ID) %>%
  ggplot(aes(n, after_stat(count))) +
  geom_histogram(fill = "seagreen", bins = 30) +
  labs(x = "Number of records per ticket", "kernel density")
```

Calculate landed pounds and revenue by species for each fish ticket. Then use these data to define proportions of pounds and revenue by species/ticket, in order to designate a "target" species for each fishing trip. We denote the target species for both revenue and pounds landed. If a ticket does not have a definitive target (classified as a proportion of landed pounds or revenue that is >10 percent more than the second-place species), denote "NONE" for the target.

```{r calc_targets}

# global options set in main_process , so it only gets read in if you are running the whole pipeline. You'll see the other variables that we carry through from main process  in the keepers vector at line 49. These are all inherited from the first chunks of main process , before ~L140 https://github.com/noaa-nwfsc/VMS-pipeline/blob/main/code/pipeline_steps/main_process.Rmd

# how much "more important" does your target need to be than the species with the second greatest catch? Expressed as a ratio.
target_cutoff <- 1.1
pacfin_weight_metric <- sym("LANDED_WEIGHT_LBS") #sym(pacfin_weight_metric)
pacfin_revenue_metric <- sym("EXVESSEL_REVENUE") # sym(pacfin_revenue_metric)

rawdat.w.targets <- ticket_df2_clean %>%
  distinct() %>%

  # NEW SEP 2024: REMOVE DEPENDENCY ON REMOVAL TYPE, i.e. JUST GROUP BY TICKET AND SPECIES
  group_by(Rec_ID,NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE) %>% 

  # OLD: Group by ticket, removal type, and species
  # group_by(Rec_ID,removal_type_code,removal_type_name,NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE) %>%

  # calculate landed pounds and revenue by species
  # NEW 09/2024- also calculate landed pounds with zero revenue (indicating personal or other use)
  summarise(spp_lbs=sum({{pacfin_weight_metric}}),
            spp_lbs_no_revenue=sum({{pacfin_weight_metric}}[{{pacfin_revenue_metric}}==0]),
            spp_revenue=sum({{pacfin_revenue_metric}})) %>% 

  # now, calculate total pounds per species across the entire ticket
  # group_by(Rec_ID,NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE) %>%
  # mutate(tot_lbs_spp=sum(spp_lbs),
  #        tot_revenue_spp=sum(spp_revenue)) %>%
  # ungroup() %>%

  # using these species totals, calculate proportions of total catch belonging to each species
  # by lbs landed and revenue
  group_by(Rec_ID) %>%
  mutate(prop_lbs_spp=spp_lbs/sum(spp_lbs),
         prop_revenue_spp=spp_revenue/sum(spp_revenue)) %>% 

  # mutate(prop_lbs_spp=tot_lbs_spp/sum(tot_lbs_spp),
  #        prop_revenue_spp=tot_revenue_spp/sum(tot_revenue_spp)) %>%
  #
  # finally, assign a TARGET to the trip, defined as the species with the
  # LARGEST proportion of revenue for that trip
  # If a species landed is not >10% more than the second species, target is NONE
  mutate(first_rev=dplyr::first(prop_revenue_spp,order_by = desc(prop_revenue_spp)),
         second_rev=dplyr::nth(prop_revenue_spp,n=2,order_by = desc(prop_revenue_spp)),
         first_rev_spp=dplyr::first(NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE,order_by= desc(prop_revenue_spp)),
         second_rev_spp=dplyr::nth(NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE,n=2,order_by= desc(prop_revenue_spp)),
         first_lbs=dplyr::first(prop_lbs_spp,order_by = desc(prop_lbs_spp)),
         second_lbs=dplyr::nth(prop_lbs_spp,2,order_by = desc(prop_lbs_spp)),
         first_lbs_spp=dplyr::first(NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE,order_by=desc(prop_lbs_spp)),
         second_lbs_spp=dplyr::nth(NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE,n=2,order_by= desc(prop_lbs_spp))) %>% 

  # check if first is >10% more than second, for revenue and landed lbs
  # or, if first and second species are the same (i.e. for a ticket with both commercial and personal use catch)
  # if so, assign that species as TARGET

  mutate(TARGET_rev=ifelse(is.na(first_rev/second_rev)|(first_rev/second_rev)>=target_cutoff|first_rev_spp==second_rev_spp,first_rev_spp,"NONE"),
         TARGET_lbs=ifelse(is.na(first_lbs/second_lbs)|(first_lbs/second_lbs)>=target_cutoff|first_lbs_spp==second_lbs_spp,first_lbs_spp,"NONE")) %>% 
  ungroup() %>%
  select(-(first_rev:second_lbs_spp))

# Add back in dates, vessel IDs, etc.##Added FTID in this list
# Tested a few different ways of doing this...
# 1. select all columns of interest and match by c(Rec_ID,NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE)
#   recID_attributes <- ticket_df2_clean %>%
#     select(Rec_ID,FTID,NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE,pacfin_port_code,port_group_code,dealerID,gear_code,gear_name,gear_group,drvid,agency_code,date,month,year)
# 2. select all columns and match by c(Rec_ID,NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE)
#   recID_attributes2 <- ticket_df2_clean %>%
#     distinct(Rec_ID,FTID,NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE,pacfin_port_code,port_group_code,dealerID,gear_code,gear_name,gear_group,drvid,agency_code,date,month,year)
# 3. select just the first set of attributes for each fish ticket/species combo
#   recID_attributes3 <- ticket_df2_clean %>%
#     group_by(Rec_ID,NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE) %>%
#     slice(1) %>%
#     ungroup()%>%
#     select(Rec_ID,FTID,NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE,pacfin_port_code,port_group_code,dealerID,gear_code,gear_name,gear_group,drvid,agency_code,date,month,year)
# As of 09/2024, we are moving forward with this last option
recID_attributes <- ticket_df2_clean %>%
  group_by(Rec_ID, NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE) %>%
  slice(1) %>%
  ungroup() %>%
  select(
    Rec_ID,
    FTID,
    NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE,
    pacfin_port_code,
    port_group_code,
    dealerID,
    CDFW_AREA_BLOCK,
    gear_code,
    gear_name,
    gear_group,
    drvid,
    agency_code,
    date,
    month,
    year
  )

# Test the join and count
# rawdat.w.targets.test1 <- rawdat.w.targets %>%
#   left_join(recID_attributes,by=c("Rec_ID","NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE"))
# rawdat.w.targets.test1 %>% count(Rec_ID,NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE) %>% pull(n) %>% range()
# for 2017 data- range is 1:40.

# rawdat.w.targets.test2 <- rawdat.w.targets %>%
#   left_join(recID_attributes2,by=c("Rec_ID","NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE"))
# rawdat.w.targets.test2 %>% count(Rec_ID,NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE) %>% pull(n) %>% range()
# for 2017 data- range is 1:2.

# rawdat.w.targets.test3 <- rawdat.w.targets %>%
#   left_join(recID_attributes3,by=c("Rec_ID","NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE"))
# rawdat.w.targets.test3 %>% count(Rec_ID,NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE) %>% pull(n) %>% range()
# for 2017 data- all combinations are unique (i.e., only 1 row per fish ticket/spp combo)
# As of 09/2024, we are moving forward with this last option
rawdat.w.targets %<>%
  left_join(recID_attributes,by=c("Rec_ID","NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE"))
# for 2017 data- all combinations are unique (i.e., only 1 row per fish ticket/spp combo)


# add all species and gear types for each ticket
rawdat.w.targets %<>%
  left_join(all.species,by = join_by(Rec_ID)) %>% 
  left_join(gear.info,by = join_by(Rec_ID))

## CAUTION: this subsection will subset all fish tix just to groundfish bottom trawl. Comment out for all other pipeline runs!!!
# rawdat.w.targets %<>% # <- rawdat.targets %>%
#   filter(gear_group == "TWL",
#                 !gear_name %in% c("MIDWATER TRAWL", "BEAM TRAWL"),
#                 !TARGET_lbs %in% c("CHLB", "PHLB", "UHLB", "BSRM", "PSHP", "OSRM", "USRM", "RPRW", "SPRW", "GPRW", "LOBS", "OCRB", "RCRB", "OMSK", "OURC", "USCU", "RURC", "SHAD", "SHPD",  "WBAS", "WCRK"),
#                   !TARGET_rev %in% c("CHLB", "PHLB", "UHLB", "BSRM", "PSHP", "OSRM", "USRM", "RPRW", "SPRW", "GPRW", "LOBS", "OCRB", "RCRB", "OMSK", "OURC", "USCU", "RURC", "SHAD", "SHPD",  "WBAS", "WCRK"))
```

##### Calculate "Other" Category

For each fish ticket, sum up an "Other" category based on the choice made in the beginning of the script (`gear_codes`)

```{r calc other}
gear_codes <- c("SHELLFISH POT (CRAB)","CRAB OR LOBSTER POT","CRAB POT","CRAB RING") # Dungeness crab

dat_targets_other <- rawdat.w.targets %>%
  group_by(Rec_ID) %>%
  # denote whether a record is in the Other category
  mutate(is_other = gear_name %in% gear_codes) %>%

  # sum the other category for each ticket
  mutate(
    all_species_rev = sum(spp_revenue[is_other]),
    all_species_lbs = sum(spp_lbs[is_other])
  ) %>%
  ungroup()
```

##### Add Species-Specific Columns

Here we add some species-specific columns based on the choice made in the beginning of the script (`species_cols`).

```{r}
# Use pivoting to define pounds and revenue for each species of interest

spp_codes <- c("DCRB") # Dungeness crab

RecID_spp_cols <- dat_targets_other %>%
  select(
    Rec_ID,
    NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE,
    spp_lbs,
    spp_lbs_no_revenue,
    spp_revenue
  ) %>%
  filter(NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE %in% spp_codes) %>%
  distinct() %>%
  pivot_longer(spp_lbs:spp_revenue) %>%
  mutate(
    type = case_when(
      name == "spp_lbs" ~ 'lbs',
      name == "spp_lbs_no_revenue" ~ 'lbs_no_revenue',
      name == "spp_revenue" ~ 'revenue'
    )
  ) %>%
  # mutate(type=ifelse(str_detect(name,'lbs'),'lbs','revenue')) %>%
  select(-name) %>%
  group_by(Rec_ID) %>%
  # spread to create the columns
  pivot_wider(
    names_from = c(NOMINAL_TO_ACTUAL_PACFIN_SPECIES_CODE, type),
    values_from = value,
    values_fill = list(value = 0)
  ) %>%
  ungroup()

# add back to main data
dat_targets_other %<>%
  left_join(RecID_spp_cols)
```

##### Format Dates

Change existing date columns to date objects and format.

```{r dates}
dat_targets_other %<>%
  mutate(date = as.Date(date, "%d-%b-%y")) %>%
  mutate(
    month = lubridate::month(date),
    month = month.name[month],
    year_mo = paste(year, lubridate::month(date), sep = "_"),
    jdate = yday(date),
    year_jdate = paste(year, jdate, sep = "_"),
    Week = week(date),
    year_Wk = ifelse(
      Week < 10,
      paste0(year, "_0", Week),
      paste0(year, "_", Week)
    )
  )
```

##### Duplicated Dates

We add two columns that denote how many tickets were submitted by a vessel in a given day. `n_tix` denotes the total number of tickets.

```{r}
dat_out <- dat_targets_other %>%

  # boolean indicating whether a record is for personal use
  # REMOVED 09/2024
  # mutate(is_personal=removal_type_name=="PERSONAL USE") %>%

  # for each vessel and date
  group_by(drvid,date) %>% 

  # calculate quantities of interest
  # mutate(n_tix=n_distinct(Rec_ID),
  #        n_pers=n_distinct(Rec_ID[is_personal]),
  #        n_nonpers_tix=n_tix-n_pers) %>%
  # select(-n_pers,-is_personal) %>%
  # ungroup()
  mutate(n_tix=n_distinct(Rec_ID)) %>% 
  # select(-n_pers,-is_personal) %>%
  ungroup()
```

##### Reorder Output

Clean up the output by reording the columns of interest

```{r reorder_dataframe}
#Added FTID to select list
dat_out %<>%
  select(-(spp_lbs:prop_revenue_spp), -is_other) %>%
  select(
    Rec_ID,
    FTID,
    pacfin_port_code,
    port_group_code,
    dealerID,
    drvid,
    agency_code,
    CDFW_AREA_BLOCK,
    gear_code_all,
    date,
    month,
    year,
    year_mo,
    jdate,
    year_jdate,
    Week,
    year_Wk,
    contains("_lbs"),
    contains("_rev"),
    species_code_all,
    all_species_lbs,
    all_species_rev
  ) %>%
  # the last distinct() removes records for which we no longer care about individual sub-trip species catches (we only care about catches of the focal species, plus total catch and TARGETs)
  distinct()
```


##### Save

Save the data!


```{r save}


write_rds(dat_out, here('Confidential', 'processed_data',
                        'fish_tickets',
                        paste0(
                          'fishtix_withFTID_DCRB_',
                          paste0(load_years[1],'-',load_years[length(load_years)]),
                          
                          '_',Sys.Date(),'.rds')
                           )
)


# create output subdirectory, if doesn't yet exist
# output_subdir <- here(
#   'Confidential',
#   'processed_data',
#   'fish_tickets'
# )
# if (!dir.exists(output_subdir)) dir.create(output_subdir)
# 
# # write output
# dat_out %>%
#   write_rds(here::here(
#     'Confidential',
#     'processed_data',
#     'fish_tickets',
#     paste0("fishtix_withFTID_", process_year, ".rds", collapse = "")
#   ))

# for(y in unique(rawdat$LANDING_YEAR)){
#   year.final.processdat <- filter(final.processdat, year == y )
#   write_csv(file = paste0(processdir, "fish_tickets ", y, " processed_multispecies03.csv"), x = year.final.processdat, row.names = FALSE)
#   cat("wrote out file for ", y, "\n")
# }
```

### Shortcut load fish tix

```{r load_alltix}

load_years <- 2011:2023

dat_out <- read_rds(here(
  'Confidential', 
  'processed_data', 
  'fish_tickets',
  paste0(
    'fishtix_withFTID_DCRB_',
    paste0(load_years[1],'-',load_years[length(load_years)]),
    '_2025-07-16.rds')
  )
)



```

#### Transform fish ticket data

Transform the fish ticket data:

1. Filter for dungeness crab related records.
2. Add temporal columns.
3. Adjust revenue for inflation with data from FRED.
4. add CDFW_AREA_BLOCK and permit tier info

Some commonly used acronyms for variable naming in the code include:

* `dcrb` dungeness crab
* `rev` revenue
* `lbs` landings
* `VMS` vessel monitoring system
* `afi` adjusted for inflation

```{r define-filters-ticket_df}
# define filters
target_rev <- "DCRB"         # revenue target
target_lbs <- "DCRB"         # landings target
state_agency_code <- "C"     # landings brought to California ports
min_depth <- 0               # minimum depth in meters
max_depth <- -150            # maximum depth in meters
min_speed <- 0               # minimum speed in m/s
max_speed <- 4.11556         # maximum speed in m/s (4.11556 m/s = 8 knots)
crab_year_start <- 11        # month defines start of crab year
winter_months <- c("November", "December", "January", "February", "March") # determine Winter or Spring-Summer season
```

```{r transform-ticket-data}
# apply same filters as VMS dataframe
dcrb_ticket_df <- dat_out %>%
  # add temporal columns
  mutate(
    year_numeric = year(date),
    month_numeric = month(date),
    week_numeric = week(date),
    day_numeric = yday(date),
    month_factor = month(date, label = TRUE, abbr = FALSE),
    year_month_character = paste0(year(date),"_", substr(ymd(date), 6, 7)),
    crab_year_character = ifelse(month_numeric >= crab_year_start, 
                                 paste0(year_numeric, "_", 1+year_numeric),
                                 paste0(year_numeric-1, "_", year_numeric)),
    season_character = as.character(ifelse(month_factor %in% winter_months, "Winter", "Spring-Summer")),
    year_month_date = ym(year_month_character)
  ) %>%
  # apply filters
  filter(TARGET_rev == target_rev | TARGET_lbs == target_lbs) %>%
  # join inflation adjustment factor and adjust revenue
  # left_join(fred_pcepi, by = join_by(year_month_date == year_month)) %>%
  # mutate(DCRB_revenue_afi = DCRB_revenue * 100 / pcepi) %>%
  # select columns
  dplyr::select(
    # identifiers
    Rec_ID,           # fish ticket ID
    FTID,
    drvid,            # vessel ID
    pacfin_port_code, # port ID
    port_group_code,  # port group ID
    agency_code,      # agency code
    CDFW_AREA_BLOCK,
    # temporal fields
    date,
    year_month_date,
    crab_year_character,
    year_month_character,
    month_factor,
    year_numeric,
    month_numeric,
    week_numeric,
    day_numeric,
    # # vessel length
    # FINAL_LENGTH,
    # # inflation adjustment index
    # pcepi,
    # dungeness crab fields
    DCRB_lbs,
    DCRB_revenue#,
    #DCRB_revenue_afi
  ) 

# jameal had to troubleshoot the CDFW_AREA_BLOCK join for a bit because of many to many relationship problems during the join. error messages and some troubleshooting comments below
 # add in CDFW block info 'CDFW_AREA_BLOCK' 06-16-2025. tried this -- many to many problem! leads to ~2,0000 extra rows. tried joining on Rec_ID == FISH_TICKET_ID, date == LANDING_YEAR, problem remains: Error in `left_join()`: ! Can't join `x$date` with `y$LANDING_YEAR` due to incompatible types. ℹ `x$date` is a <date>. ℹ `y$LANDING_YEAR` is a <double>. Backtrace:
# even after adding all of these additional specifications, still get a warning on many to many relationship in 2011 (row 3 of x, row 1223140 of y); similar problems for 2012-2014. but 2021-2022 are fine.

# take a peek at columns in the resulting dataframe
colnames(dcrb_ticket_df)

# write rds
write_rds(dcrb_ticket_df, here('Confidential', 'processed_data', 'fish_tickets',
                           paste0('fish tickets_DCRB_',
                                  paste0(load_years[1],'-',load_years[length(load_years)]),
                                  '_',Sys.Date(),'.rds')
                           )
)

```

```{r count_trips}
# count total records, trips and years
n_records <- nrow(dcrb_ticket_df)
n_trips   <- n_distinct(dcrb_ticket_df$Rec_ID)
n_years   <- n_distinct(dcrb_ticket_df$year_numeric)
```

If we want to skip the pre-processing above, we can simply read in the rds from the chunk above

```{r load_dcrb_tix}
# choose years of data to load
load_years <- 2011:2023

dcrb_ticket_df <- read_rds(
  here('Confidential', 'processed_data', 'fish_tickets',
       paste0('fish tickets_DCRB_',
              paste0(load_years[1],'-',load_years[length(load_years)]),
              '_2025-07-17.rds')
       )
)


```

The Dungeness crab fish ticket dataframe has `r n_records` records (fish tickets), `r n_trips` distinct trips (fish tickets), across `r n_years` years.

<!-- There are more records than trips due to the vessel registration processing code including the lookback window data in step 2 of the pipeline. I looked at some examples of duplicates, and they occur for tickets within the lookback window - for example, a ticket landed on Dec. 28, 2014 is processed in the 2014 and 2015 runs of the pipeline for vessel registration. I suspect this could be fixed in the pipeline itself, but I'm leaving it be for now. -->

<!-- This is fixed here in the plotting code by removing duplicates when one length is NA, and I didn't see any duplication within the VMS data. -->

<!-- ```{r} -->
<!-- # identify which records need to be de-duplicated, by finding RecIDs with 2 records, 1 of which has NA length -->
<!-- qa_df <- dcrb_ticket_df %>%  -->
<!--   group_by(Rec_ID) %>% -->
<!--   summarize(n = n(), -->
<!--             na_length = sum(is.na(FINAL_LENGTH))) %>% -->
<!--   mutate(remove_na = (na_length == 1) & (n == 2)) %>% -->
<!--   arrange(desc(n)) -->

<!-- # take a peek at columns in the resulting dataframe -->
<!-- colnames(qa_df) -->

<!-- # how many records need de-duplication (have 2 records, 1 of which has NA length)? -->
<!-- qa_df %>% group_by(n, na_length, remove_na) %>% summarise(n_records = n()) %>% arrange(desc(n_records)) # 1229 records -->

<!-- # join back to dcrb_ticket_df -->
<!-- qa_dcrb_ticket_df <- dcrb_ticket_df %>% -->
<!--   left_join(qa_df, by = join_by(Rec_ID)) %>% -->
<!--   filter(!(remove_na & is.na(FINAL_LENGTH))) -->

<!-- # count total records, trips and years -->
<!-- n_qa_records <- nrow(qa_dcrb_ticket_df) -->
<!-- n_qa_trips   <- n_distinct(qa_dcrb_ticket_df$Rec_ID) -->
<!-- n_qa_years   <- n_distinct(qa_dcrb_ticket_df$year_numeric) -->
<!-- ``` -->

<!-- The QA'ed Dungeness crab fish ticket dataframe has `r n_qa_records` records (fish tickets), `r n_qa_trips` distinct trips (fish tickets), across `r n_qa_years` years. Now there are no duplicate fish tickets. -->

### Fish tix matched to VMS data



1. Load Matched VMS and fish tix (matched_filtered_withFTID_length_yyyy.rds) 

Pre-requisite: Run the pipeline steps 1-5 for calendar years 2011-2023.

The code below borrows heavily from Brooke Hawkins' notebook: 03_blh_hindcasts-of-california-dungeness-crab-fishing-activity.Rmd

*** Note you can skip to the chunk load_dcrb_vms and avoid running chunks 20, 22-26, but you still need chunk 21, load_10arcminute_grid

```{r load vms_tix}

# choose years of data to load
load_years <- 2011:2023

# load fish ticket data joined to vessel length data, but before it was joined to VMS data
vms_df <- purrr::map(load_years, function(ly) {
  read_rds(here('Confidential', 'processed_data', 'processed_2025-03-19', 'filtered', paste0('matched_filtered_withFTID_length_', ly, '.rds')))
}) %>% bind_rows()

glimpse(vms_df)


```

2. Load grid

```{r load_10arcminute_grid}

# 10-arcminute grid shapefile
grd <- read_sf(here('spatial_data', 'grids', 'ten_arcminute_grid_shore_lamb.shp'))
names(grd)

# background map, from rnaturalearth
coast <- ne_states(country='United States of America',returnclass = 'sf') %>%
  filter(name %in% c('California','Oregon','Washington','Nevada')) %>%
  st_transform(crs = st_crs(grd))

```


3. Join VMS and fish ticket data to grid

identifies the CDFW block in which each VMS record falls

This chunk is inspired by overlay_points_on_grid.Rmd from the grid_overlay_edits branch

```{r match_points}

pt <- proc.time()

# convert vms to spatial object (longitude/latitude) and join to grid
vms_sf <- vms_df %>%
  st_as_sf(coords=c('LON','LAT'),crs=4326) %>% 
  # then, convert to planar projection to match the grid
  st_transform(st_crs(grd)) %>%
  st_join(grd) 

vms_grid10_df <- vms_sf %>%
  st_set_geometry(NULL)

proc.time()-pt

# 55s for 2022 data
# 764 s for 2011-2023 data

rm(vms_df)
rm(vms_sf)

```


4. Load inflation adjustment factors from FRED. Use annual instead of monthly

Adjust revenue for inflation using Fred GDP data, adapted from R code by Erin Steiner. This product uses the FRED® API but is not endorsed or certified by the Federal Reserve Bank of St. Louis.

Pre-requisite: Create a FRED account and API key. See https://fred.stlouisfed.org/docs/api/api_key.html.

```{r inflation-adjustment}
# insert your FRED API key
fredr_set_key('ea19d3d447eb4dc773db00c5a2e61f22') 

# download the quarterly inflation adjustments from Fred from 1985 to present
fred_gdpdefl <- fredr(
    series_id = "GDPDEF", 
    observation_start = as.Date(paste0(min(load_years)-1, "-01-01"))
  )

# generate the mean annual deflators based on the quarterly values
gdp_defl <- mutate(fred_gdpdefl, year = year(date)) %>%
  group_by(year) %>%
  summarize(defl = mean(value), .groups = 'drop') %>%
  mutate(inflation_adjustment_factor = defl / defl[year == max(year)]) %>%
  select(year, inflation_adjustment_factor)

# write inflation adjustment factors
write.csv(file = here("Confidential", "hindcast_output", "tables", "annual_inflation_adjustment.csv"), x = gdp_defl)

# plot inflation adjustment factors
gdp_defl %>% ggplot(aes(x = as.factor(year), y = inflation_adjustment_factor)) + geom_point()
ggsave(here("Confidential", "hindcast_output", "figures", "annual_inflation_adjustment.png"))

# clean up
rm(fred_gdpdefl)
```

#### Transform data

```{r define-filters-vms_df}
# define filters
target_rev <- "DCRB"         # revenue target
target_lbs <- "DCRB"         # landings target
state_agency_code <- "C"     # landings brought to California ports
min_depth <- 0               # minimum depth in meters
max_depth <- -150            # maximum depth in meters
min_speed <- 0               # minimum speed in m/s
max_speed <- 4.11556         # maximum speed in m/s (4.11556 m/s = 8 knots)
crab_year_start <- 11        # month defines start of crab year
winter_months <- c("November", "December", "January", "February", "March") # determine Winter or Spring-Summer season
```

```{r transform_vms_data}

# transform VMS and fish ticket data
dcrb_df <- vms_grid10_df %>%
  # use distinct to avoid duplicates caused by vessel length matching step (do not keep rows that are the same except for "FINAL_LENGTH")
  distinct(
    VMS_RECNO, Rec_ID, FTID, pacfin_port_code, port_group_code, dealerID, drvid, agency_code, date, TARGET_lbs, all_species_lbs, DCRB_lbs, DCRB_lbs_no_revenue, TARGET_rev, all_species_rev, DCRB_revenue, species_code_all, X_COORD, Y_COORD,  westcoastdate_notime, avg_speed_recalc, BLOCK10_ID,
     .keep_all = TRUE
  ) %>%
  # add temporal columns
  mutate(
    year = lubridate::year(westcoastdate_notime),
    month = lubridate::month(westcoastdate_notime, label=TRUE, abbr=FALSE),
    month_numeric = month(westcoastdate_notime),
    year_month = paste0(lubridate::year(westcoastdate_notime),"_", substr(lubridate::ymd(westcoastdate_notime), 6, 7)),
    crab_year = ifelse(month_numeric >= crab_year_start, paste0(year, "_", 1+year), paste0(year-1, "_", year)),
    season = as.character(ifelse(month %in% winter_months, "Winter", "Spring-Summer")),
    week_of_year = week(westcoastdate_notime),
    day_of_year = yday(westcoastdate_notime),
    # year_month_character = paste0(year(date),"_", substr(ymd(date), 6, 7)),
    year_month_date = floor_date(ymd(westcoastdate_notime), 'month')
  ) %>%
  # apply filters
  filter(TARGET_rev == target_rev | TARGET_lbs == target_lbs) %>%
  filter(agency_code == state_agency_code) %>%
  filter(NGDC_M <= min_depth & NGDC_M >= max_depth) %>%
  filter(avg_speed_recalc <= max_speed & avg_speed_recalc >= min_speed) %>% 
  # join inflation adjustment factor and adjust revenue
  left_join(gdp_defl, by = join_by(year == year)) %>%
  mutate(DCRB_revenue_afi = DCRB_revenue * 100 / inflation_adjustment_factor) %>% #at this point, 14,418,569 rows have been reduced to 1,077,651 rows
  # select columns
  # drop "gear_code_all", "end", "prev_ticket_dttm", "lookback_dttm", "begin", "trip_dur", "UTCDATETIME", "VESSEL_NAME", "AVG_SPEED", "AVG_COURSE", "DECLARATION_CODE", "westcoastdate",  "has_vms", "multispecies", "segment_dur",      "in_port", "Shoreline", "row_index", "col_index" 
  dplyr::select(
    # identifiers
    Rec_ID,           # fish ticket ID
    FTID,
    VMS_RECNO,        # VMS ping ID
    drvid,            # vessel ID
    agency_code,
    pacfin_port_code, # port ID
    port_group_code,  # port group ID
    BLOCK10_ID,       # grid cell ID
    BLOCK10_n,
    Offshrblok,
    Adj_cell_1, 
    Adj_cell_2, 
    Adj_cell_3, 
    Adj_Cell_4, 
    Adj_Cell_5, 
    Adj_Cell_6, 
    Adj_Cell_7, 
    Adj_Cell_8,
    Area_m2,
    WM_NGDC_m,
    # temporal fields
    westcoastdate,
    westcoastdate_notime,
    crab_year, 
    season,
    year,
    year_month,
    month, 
    month_numeric,
    week_of_year,
    day_of_year,
    # vessel length
    FINAL_LENGTH,
    # dungeness crab fields
    DCRB_lbs,
    DCRB_revenue_afi
  ) %>%
   left_join(
    dcrb_ticket_df %>% dplyr::select(year=year_numeric, FTID, CDFW_AREA_BLOCK)
    )

glimpse(dcrb_df)

dim(distinct(dcrb_df))

#length(which(is.na(dcrb_df$CDFW_AREA_BLOCK))) # 


```



```{r write dcrb_vms_grd}

# write rds for non-spatial version
write_rds(dcrb_df, here(
  'Confidential', 
  'processed_data', 
  'vms_fish_tickets',
  paste0(
    'dcrb_vms_w_grd_10_arcmin_',
    paste0(load_years[1],'-',load_years[length(load_years)]),
    '_',Sys.Date(),'.rds'
    )
)
)



```

# Analyze matched VMS and fish tix

## shortcut load

```{r load_dcrb_vms}

load_years <- 2011:2023

dcrb_df <- read_rds(here(
  'Confidential', 
  'processed_data', 
  'vms_fish_tickets',
  paste0(
    'dcrb_vms_w_grd_10_arcmin_',
    paste0(load_years[1],'-',load_years[length(load_years)]),
    '_2025-07-29.rds'
    )
  )
)



```

## Determine correspondence between reported and tracked fishing locations

Quantify % of pings that match reported block, and % of tix where at least one ping matches reported block

For now, I am not accounting for individual vessel level effects, but instead looking across all trips and all vessels all at once

```{r match_ping_reported}

# # give every ticket a reported an offshore block ID, either because it was reported as 1032:1042 in CDFW_AREA_BLOCK or by using the lookup provided by Blake in Offshrblok
# 
# 
# dcrb_df_match_long <- dcrb_df %>%
#   mutate(
#     match_ping_report = case_when(
#       BLOCK10_ID == CDFW_AREA_BLOCK ~ 1,
#       BLOCK10_ID != CDFW_AREA_BLOCK ~ 0#,
#       #.default = 9999 # 4225/1078180 NAs. ignore for now
#       ),
#     offshore_reported = case_when(
#       CDFW_AREA_BLOCK %in% offshore_blk_IDs ~ 1,
#       !CDFW_AREA_BLOCK %in% offshore_blk_IDs ~ 0#,
#       #.default = 9999
#     ),
#    offshore_block = case_when(
#       offshore_reported == 1 ~ offshore_reported,
#       offshore_reported == 0 ~ grd$Offshrblok[which(grd$BLOCK10_ID == CDFW_AREA_BLOCK)]
#   ) 

offshore_blk_IDs <- 1032:1042
medium_blk_IDs <- c(138:139, 240:241, 280:281, 484:487, 545:546, 568, 649:650, 697, 776:777, 895:897)

dcrb_df_match_long <- dcrb_df %>%
  mutate(
    match_ping_report = case_when(
      BLOCK10_ID == CDFW_AREA_BLOCK ~ 1,
      BLOCK10_ID != CDFW_AREA_BLOCK ~ 0#,
      #.default = 9999 # 4225/1078180 NAs. ignore for now
      ),
    offshore_reported = case_when(
      CDFW_AREA_BLOCK %in% offshore_blk_IDs ~ 1,
      !CDFW_AREA_BLOCK %in% offshore_blk_IDs ~ 0#,
      #.default = 9999
    ),
    medium_reported = case_when(
      CDFW_AREA_BLOCK %in% medium_blk_IDs ~ 1,
      !CDFW_AREA_BLOCK %in% medium_blk_IDs ~ 0#,
      #.default = 9999
    )
  )

# length(which(dcrb_df_match$offshore_reported == 1))
# length(which(dcrb_df_match$offshore_reported == 0))

dcrb_df_match_all_fishtix <- dcrb_df_match_long %>%
  # summarise by each fish ticket
  group_by( 
    crab_year, Rec_ID, FTID, drvid, agency_code, pacfin_port_code, port_group_code, FINAL_LENGTH #, DCRB_lbs, DCRB_revenue_afi. commented out DCRB_revenue_afi for now because for tix with vms records that span multiple months we end up with >1 row for each unique ticket. fix later
  ) %>%
  summarise(
    total_match_ping_report = sum(match_ping_report, na.rm=TRUE),
    num_vms_records = n(),
    match_ping_report_YN = ifelse(total_match_ping_report >=1, 1, 0),
    prop_match_ping_report = total_match_ping_report / num_vms_records
  ) %>%
  ungroup()

dcrb_df_match_all_crabyrs <- dcrb_df_match_all_fishtix %>%
  # summarise across all fish tix
  group_by(
    crab_year
  ) %>%
  summarise(
    prop_match_ping_report_YN = sum(match_ping_report_YN)/n(),
    mean_prop_match_ping_report = mean(prop_match_ping_report, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(
    year_numeric = substr(crab_year,6,9)
  ) %>%
  filter(year_numeric != 2024)

dcrb_df_match_block10_fishtix <- dcrb_df_match_long %>%
  filter(offshore_reported == 0) %>%
  filter(medium_reported == 0) %>%
  # summarise by each fish ticket
  group_by( 
    crab_year, Rec_ID, FTID, drvid, agency_code, pacfin_port_code, port_group_code, FINAL_LENGTH #, DCRB_lbs, DCRB_revenue_afi. commented out DCRB_revenue_afi for now because for tix with vms records that span multiple months we end up with >1 row for each unique ticket. fix later
  ) %>%
  summarise(
    total_match_ping_report = sum(match_ping_report, na.rm=TRUE),
    num_vms_records = n(),
    match_ping_report_YN = ifelse(total_match_ping_report >=1, 1, 0),
    prop_match_ping_report = total_match_ping_report / num_vms_records
  ) %>%
  ungroup()

dcrb_df_match_block10_crabyrs <- dcrb_df_match_block10_fishtix %>%
  # summarise across all fish tix
  group_by(
    crab_year
  ) %>%
  summarise(
    prop_match_ping_report_YN = sum(match_ping_report_YN)/n(),
    mean_prop_match_ping_report = mean(prop_match_ping_report, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(
    year_numeric = substr(crab_year,6,9)
  ) %>%
  filter(year_numeric != 2024)

```

## Make % match plots

Plot % of tickets with at least one match between reported and tracked blocks - include both reported offshore blocks and reported block10s

```{r low_bar_plot_all}

dcrb_df_match_all_crabyrs %>%
  ggplot(
    aes(
        x = year_numeric, y = prop_match_ping_report_YN, group = 1
        )
  ) +
  geom_point() +
  geom_line() +
  ylim(0,1) +
  labs(
    subtitle = 'includes both reported offshore blocks and reported block10s',
    y = 'Proportion of trips with >=1 VMS record\nmatching reported CDFW 10-arc minute block',
    x = 'Crab year'
       ) +
  theme_classic()

ggsave(filename = here("Confidential", "hindcast_output", "figures", "time_series_prop_dcrb_tix_match_ping_report_all.png"), width = 8, height = 6)

```

Plot % of tickets with at least one match between reported and tracked blocks - include reported block10s only

```{r low_bar_plot_block10}

dcrb_df_match_block10_crabyrs %>%
  ggplot(
    aes(
        x = year_numeric, y = prop_match_ping_report_YN, group = 1
        )
  ) +
  geom_point() +
  geom_line() +
  ylim(0,1) +
  labs(
    subtitle = 'includes only reported block10s',
    y = 'Proportion of trips with >=1 VMS record\nmatching reported CDFW 10-arc minute block',
    x = 'Crab year'
       ) +
  theme_classic()

ggsave(filename = here("Confidential", "hindcast_output", "figures", "time_series_prop_dcrb_tix_match_ping_report_block10.png"), width = 8, height = 6)

```

## Summarize data frame for mapping

Focus on fish tix with self-reported 10-arc minute blocks and associated VMS data

Use Owen's code as a basis. https://github.com/noaa-nwfsc/VMS-pipeline/blob/7fac4c72968f6235ad1f73975128821cf3ddd933/code/notebooks/owen/summary%20mapping%20dcrb.qmd 

```{r sum_for_maps}

# filter to CDFW small block tix only
# group_by by crab_year and BLOCK10_ID
# sum num Rec_ID reporting each block, num VMS_RECNO in each block
# repeat faceted by pacfin_port_code, season if desired

#offshore_blk_IDs <- 1032:1042
medium_blk_IDs <- c(138:139, 240:241, 280:281, 484:487, 545:546, 568, 649:650, 697, 776:777, 895:897)

dcrb_df_block10 <- dcrb_df %>%
  # mutate(
  #   offshore_reported = case_when(
  #     CDFW_AREA_BLOCK %in% offshore_blk_IDs ~ 1,
  #     !CDFW_AREA_BLOCK %in% offshore_blk_IDs ~ 0#,
  #     #.default = 9999
  #   )
  # ) %>%
  # filter(offshore_reported == 0) %>%
  filter(BLOCK10_ID < 1000) %>%
  filter(!BLOCK10_ID %in% medium_blk_IDs)

# function to aggregate. make sure not to compare post 2020 high ping rate to pre 2020
summarize_dcrb_attribute_block10 <- function(att,temporal_breaks){
  
  # first, establish the attribute to be summarized, and select it from the dataset
  if(att=="trips") df <- dcrb_df_block10 %>%
      rename(attribute=Rec_ID)
  if(att=="pings") df <- dcrb_df_block10 %>%
      rename(attribute=VMS_RECNO)
  
  # then, group the data into the appropriate temporal breaks and summarize based on the chosen measure
  # if the option is all years, sum across all months of the crab year, separating pre Sep 2020 from post due to an increase in ping rate from 60-mn intervals to 15-min intervals
  if(temporal_breaks=="all years") {
    df2 <- df %>% 
      filter(month_numeric %in% c(1,2,3,4,5,6,7,11,12)) %>% 
      mutate(temporal_break=ifelse(westcoastdate_notime < '2020-09-01',"60-minute ping intervals","15-minute ping intervals")) %>% 
      group_by(temporal_break,crab_year,BLOCK10_ID) %>% 
      summarise(attr_n=n_distinct(attribute,na.rm=T)) %>% 
      group_by(temporal_break,BLOCK10_ID) %>% 
      # calculate mean and sd
      summarise(
        mean_allyrs=mean(attr_n,na.rm=T),
        sd_allyrs=sd(attr_n,na.rm=T),
      ) %>% 
      # calculate rank and quantile
      ungroup() %>% 
      mutate(ranked=cume_dist(mean_allyrs)) %>% 
      arrange(desc(ranked)) %>%
      mutate(cume=cumsum(mean_allyrs)) %>%
      mutate(quant=cume/max(cume)) %>%
      # binned into 5% chunks
      mutate(quantbinned=ntile(quant,10)*10) %>%
      mutate(quantbinned=ordered(quantbinned)) %>% 
      dplyr::select(-cume)
  }
  
  #df_out <- df2
  
  # finally, join the spatial data
  df_out <- df2 %>%
    #left_join(grd_dissolved,by="BLOCK10_ID") # pick up here 07-29-2025. need to make grd_dissolved
    left_join(grd,by="BLOCK10_ID") %>% st_as_sf()
  
  df_out
}

```


## Compare metrics

### Example correlations between fish tix and VMS records

These are very ugly.

```{r vms_rec_map}

trips_allyrs <- summarize_dcrb_attribute_block10('trips', 'all years') %>% mutate(metric = 'trips') %>% rename(quant_trips = quant, mean_allyrs_trips = mean_allyrs)
pings_allyrs <- summarize_dcrb_attribute_block10('pings', 'all years') %>% mutate(metric = 'pings') %>% rename(quant_pings = quant, mean_allyrs_pings = mean_allyrs)

trips_allyrs %>% st_join(pings_allyrs) %>%
  ggplot()+
  geom_point(aes(x=quant_trips, y=quant_pings))+
  theme(panel.background=element_rect(fill='white'))

pings_allyrs %>% st_join(trips_allyrs) %>%
  ggplot()+
  geom_point(aes(x=mean_allyrs_trips, y=mean_allyrs_pings))+
  theme(panel.background=element_rect(fill='white'))

```

### Maps

#### Get set up to make maps.

We can filter the large 10-arc minute grid to only encompass the cells with some crab fishing effort. Do this with a filter-buffer-crop workflow (could also use, e.g., a distance from shore or depth cutoff):

```{r dcrb_cells}
# filter the overall grid using grid cells with DCRB effort in any month/year
cells_we_need <- unique(dcrb_df$BLOCK10_ID)
grid10_filt <- grd %>% filter(BLOCK10_ID %in% cells_we_need)

# buffer by 50km
grid10_filt_buff <- grid10_filt %>% summarise() %>% st_buffer(1e4)

# crop to make the final grid
grid10_final <- grd %>% st_filter(grid10_filt_buff)
bbox <- st_bbox(grid10_final)

# map this
ggplot()+
  geom_sf(data=grid10_final,aes(fill=NGDC_med_m),col=NA)+
  geom_sf(data=coast,fill='#dacf80',col='gray80')+
  xlim(bbox[1],bbox[3])+ylim(bbox[2],bbox[4])+
  scale_fill_viridis(option="turbo",direction=-1)
```

We make a spatially rotated grid, for better visualization when mapping the whole coastline. We use custom projections for this, which B. Feist designed. Based on that work, here we provide generic code for any given amount of rotation.

```{r rotate_map}
# this is the first wkt file ('well-known text') that B. Feist created, containing the info on the custom spatial projection
rotation_wkt <- read_file(here('spatial_data','map_rotation','Rectified_Skew_Orthomorphic_center_30deg_rotation.wkt'))

# working in this simple text format, we can find and replace to change the rotation of the map
crs_rotated <- function(deg){
  crs_out <- sub('Skew Grid\",30',paste0('Skew Grid\",',deg),rotation_wkt)
  crs_out
}

# let's use this to make a 15 deg rotation, which seems about right for our study domain
rot15 <- crs_rotated(15) %>% st_crs()

# use this crs to reprojec the 5k grid
grid10_rot15 <- grid10_final %>% st_transform(rot15)
bbox_rot15 <- st_bbox(grid10_rot15) +c(0,0,-20000,0)

# reproject for ca only
grid10_rot15_ca <- grid10_final %>% filter(State == 'CA') %>% st_transform(rot15)
bbox_rot15_ca <- st_bbox(grid10_rot15 %>% filter(State == 'CA')) +c(0,0,-20000,0)

# map
ggplot()+
  geom_sf(data=grid10_rot15, aes(fill=NGDC_med_m),col=NA)+
  geom_sf(data=coast,fill='#dacf80',col='gray80')+
  xlim(bbox_rot15[1],bbox_rot15[3])+ylim(bbox_rot15[2],bbox_rot15[4])+
  scale_fill_viridis(option="turbo",direction=-1)

# map ca only
ggplot()+
  geom_sf(data=grid10_rot15_ca, aes(fill=NGDC_med_m),col=NA)+
  geom_sf(data=coast,fill='#dacf80',col='gray80')+
  xlim(bbox_rot15_ca[1],bbox_rot15_ca[3])+ylim(bbox_rot15_ca[2],bbox_rot15_ca[4])+
  scale_fill_viridis(option="turbo",direction=-1)

```

start adding mapping code, see line 808 of Owen's make_map_coastwide()

```{r make_map_function}

make_map_coastwide <- function(df, column,scl_ord=1,...) {
  
  dat2 <- df %>% st_transform(rot15)
  
  #is the value a factor (i.e., categorical?)
  is_factor <- df %>% pull( {{column}} ) %>% is.factor()
  
  # coastwide map
  if(is_factor){
    ggplot() + 
      geom_sf(data=dat2,aes(fill= {{column}}),col=NA)+
      geom_sf(data=coast,fill='#dacf80',col='gray80')+
      xlim(bbox_rot15[1],bbox_rot15[3])+
      ylim(bbox_rot15[2],bbox_rot15[4])+
      scale_fill_viridis_d(na.value='#abd0d6',option='turbo',direction=scl_ord,...)+
      theme(panel.background=element_rect(fill="#abd0d6"),panel.grid=element_line(color='gray70',linewidth = 0.2))
  } else{
    ggplot() + 
      geom_sf(data=dat2,aes(fill= {{column}}),col=NA)+
      geom_sf(data=coast,fill='#dacf80',col='gray80')+
      xlim(bbox_rot15[1],bbox_rot15[3])+
      ylim(bbox_rot15[2],bbox_rot15[4])+
      scale_fill_viridis(option='turbo',na.value='#abd0d6',direction=scl_ord,...)+
      theme(panel.background=element_rect(fill="#abd0d6"),panel.grid=element_line(color='gray70',linewidth = 0.2))
  }
  
}

make_map_ca <- function(df, column,scl_ord=1,...) {
  
  dat2 <- df %>% filter(State == 'CA') %>% st_transform(rot15)
  
  #is the value a factor (i.e., categorical?)
  is_factor <- df %>% pull( {{column}} ) %>% is.factor()
  
  # ca map
  if(is_factor){
    ggplot() + 
      geom_sf(data=dat2,aes(fill= {{column}}),col=NA)+
      geom_sf(data=coast,fill='#dacf80',col='gray80')+
      xlim(bbox_rot15_ca[1],bbox_rot15_ca[3])+
      ylim(bbox_rot15_ca[2],bbox_rot15_ca[4])+
      scale_fill_viridis_d(na.value='#abd0d6',option='turbo',direction=scl_ord,...)+
      theme(panel.background=element_rect(fill="#abd0d6"),panel.grid=element_line(color='gray70',linewidth = 0.2))
  } else{
    ggplot() + 
      geom_sf(data=dat2,aes(fill= {{column}}),col=NA)+
      geom_sf(data=coast,fill='#dacf80',col='gray80')+
      xlim(bbox_rot15_ca[1],bbox_rot15_ca[3])+
      ylim(bbox_rot15_ca[2],bbox_rot15_ca[4])+
      scale_fill_viridis(option='turbo',na.value='#abd0d6',direction=scl_ord,...)+
      theme(panel.background=element_rect(fill="#abd0d6"),panel.grid=element_line(color='gray70',linewidth = 0.2))
  }
  
}

# just plotting NGDC_med_m (depth) as the fill to make sure everything is oriented correctly
make_map_coastwide(grid10_final, NGDC_med_m)+labs(fill="dept")
make_map_ca(grid10_final, NGDC_med_m)+labs(fill="dept")
# make_map_coastwide(grid5k_final, centro_lat,scl_ord = -1)+labs(fill="latitude")

```

#### Now make maps

Subset to CA only

```{r ca-only}

trips_allyrs_ca <- trips_allyrs %>%
  filter(State == 'CA')

pings_allyrs_ca <- pings_allyrs %>%
  filter(State == 'CA')

```

1. 60-minute ping interval years

```{r 60_min_maps}

make_map_ca(trips_allyrs_ca %>% filter(temporal_break == '60-minute ping intervals'), mean_allyrs_trips)+labs(fill="mean number of trips", title = 'CDFW small block-reported fishing locations', subtitle = 'Nov 2011 - Sep 2020\n 60-minute ping intervals')
ggsave(here("Confidential", "hindcast_output", "figures", "60-minute ping intervals reported fishing locations.png"))


make_map_ca(pings_allyrs_ca %>% filter(temporal_break == '60-minute ping intervals'), mean_allyrs_pings)+labs(fill="mean number of pings", title = 'VMS-recorded fishing locations', subtitle = 'Nov 2011 - Sep 2020\n 60-minute ping intervals')
ggsave(here("Confidential", "hindcast_output", "figures", "60-minute ping intervals vms-recorded fishing locations.png"))

```

2. 15-minute ping interval years

```{r 15_min_maps}

make_map_ca(trips_allyrs_ca %>% filter(temporal_break == '15-minute ping intervals'), mean_allyrs_trips)+labs(fill="mean number of trips", title = 'CDFW small block-reported fishing locations', subtitle = 'Sep 2020 - Dec 2023\n 15-minute ping intervals')
ggsave(here("Confidential", "hindcast_output", "figures", "15-minute ping intervals reported fishing locations.png"))

make_map_ca(pings_allyrs_ca %>% filter(temporal_break == '15-minute ping intervals'), mean_allyrs_pings)+labs(fill="mean number of pings", title = 'VMS-recorded fishing locations', subtitle = 'Sep 2020 - Dec 2023\n 15-minute ping intervals')
ggsave(here("Confidential", "hindcast_output", "figures", "15-minute ping intervals vms-recorded fishing locations.png"))

```