---
title: "Step by Step VMS Processing"
author: "Owen Liu"
date: Last Run "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup}
library(knitr)
library(here)
```

# Master VMS and Fish Ticket Processing Workflow

This document collates all individual steps of VMS/Fish Ticket processing for a chosen year.

```{r timer, include=FALSE}
alltime <- proc.time()
```

## Choose Fishery

Indicate the PacFIN codes for the fish ticket species you are interested in tracking, as well as the gears. Landings and revenue for these species and gears will be tracked in the pipeline, for both targeted and non-targeted trips.

```{r}
spp_codes <- c("ALBC")
# here's your options
#https://pacfin.psmfc.org/pacfin_pub/data_rpts_pub/code_lists/sp.txt

gear_codes <- c("SEN","POL")
# https://pacfin.psmfc.org/pacfin_pub/data_rpts_pub/code_lists/agency_gears.txt
```

Indicate the metric to use in determining targets of fishing trips, as well as the threshold for determining the target of each fishing trip.

```{r}
# afi_revenue for now
pacfin_weight_metric <- "LANDED_WEIGHT_LBS" # another option is "LANDED_WEIGHT_MTONS"
pacfin_revenue_metric <- "AFI_EXVESSEL_REVENUE" # another option is AFI_EXVESSEL_REVENUE
# target_metric <- "revenue" # afi_revenue is an alternative option for newer fish tickets

## how much "more important" does your target need to be than the species with the second greatest catch? Expressed as a ratio.

target_cutoff <- 1

## do you want to report revenue and lbs for the given target species? (dynamic)
# include_target <- TRUE
```

## Choose Year to Process

This parameter will get passed through all processing steps

```{r choose year}
# Leena will be working through all steps for 2014 as a test
alltime <- proc.time()
process_year <- 2009
```

## Lookback Window

We enforce a "lookback window" when matching the VMS and fish tickets such that if there are more than X days [an option that can be changed] between successive tickets for the same vessel, we only include data for those X days. Look for the lookback window option in the master process script.

Create lookback object. Here is where the lookback window duration can be changed. Default is 7 days.

```{r lookback}
lookback_window <- 7 # in days
year_begin <- lubridate::ymd(paste0(process_year,"-01-01"), tz= "America/Los_Angeles")
year_end <- lubridate::ymd(paste0(process_year,"-12-31"), tz= "America/Los_Angeles")

if(process_year == 2009){
  lookback_begin <- year_begin
} else{
  lookback_begin <- year_begin - lubridate::days(lookback_window)
}
```

# Step 1: Process Raw Fish Ticket Data

```{r step1,message=F,warning=F,child=here::here('scripts updated for OWE','01_process_fish_tickets_owe.Rmd')}
```
```{r}
x<-proc.time()-alltime
```

**So far, this pipeline for `r process_year` VMS data has taken `r round(x[3]/60,2)` minutes to run.**

# Step 2: Report Vessel Lengths

```{r step2, message=F,warning=F,child=here::here('scripts updated for OWE','02_calculate_vessel_lengths_owe.Rmd')}
```
```{r}
x<-proc.time()-alltime
```

**So far, this pipeline for `r process_year` VMS data has taken `r round(x[3]/60,2)` minutes to run.**

# Step 3: Process Raw VMS Data

```{r step3, message=F,warning=F,child=here::here('scripts updated for OWE','03_process_vms_owe.Rmd')}
```
```{r}
x<-proc.time()-alltime
```

**So far, this pipeline for `r process_year` VMS data has taken `r round(x[3]/60,2)` minutes to run.**

# Step 4: Match VMS and Fish Ticket Data

```{r step5, message=F,warning=F,child=here::here('scripts updated for OWE','04_match_vms_fishtix_owe.Rmd')}
```
```{r}
x<-proc.time()-alltime
# cat('This step for',process_year,'VMS data took',round(x[3]/60,2),'minutes to run.')
```

**So far, this pipeline for `r process_year` VMS data has taken `r round(x[3]/60,2)` minutes to run.**

# Step 5: Filter Matched Data by Speed and Proximity to Ports

```{r step5, message=F,warning=F,child=here::here('scripts updated for OWE','05_filter_matched_data_owe.Rmd')}
```
```{r}
x<-proc.time()-alltime
```

**So far, this pipeline for `r process_year` VMS data has taken `r round(x[3]/60,2)` minutes to run.**
