---
title: "Match VMS data to 3km grid and prep for DCRB analysis for Feist et al. wind-fisheries tradeoffs paper"
output: 
  html_document:
    toc: true
    toc_float: true
---

## Purpose

Matching the output VMS data to a multi-region grid to use for summarizing fishing effort. Then, filtering data for crab fishing events and assigning revenue and pounds of DCRB to each geolocation.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libs,message=FALSE,warning=FALSE}
library(tidyverse)
library(magrittr)
library(sf)
library(here)
library(knitr)
library(lubridate)
```


## Sample Year

Import files for an example year to match. We use the matched, filtered, interpolated VMS data

```{r}
# clean, interpolated vms data from 2020. produced by leena riekkola september 2022
vms2020 <- read_rds("/Users/jameal.samhouri/Documents/RAIMBOWT/Processed Data/VMS/interpolation/2020interpolated.rds")

# 3x3 grid shapefile
grd <- read_sf("/Users/jameal.samhouri/Documents/RAIMBOWT/Processed Data/3x3 Grid/owem_grid_eez_poly_lamb.shp")
names(grd)

```

## Match Points

```{r}
# convert vms to spatial object (longitude/latitude)
vms_sf <- vms2020 %>%
  st_as_sf(coords=c('LONGITUDE','LATITUDE'),crs=4326) %>% 
  # then, convert to planar projection to match the grid
  st_transform(st_crs(grd))

# do the join
pt <- proc.time()
vms_grd_match <- vms_sf %>%
  st_join(grd)
proc.time()-pt
```

## Check

Did the join work? How many unique 5km, 25km, and 50km cells are represented in the 2009 data?

```{r checks}
names(vms_grd_match)

length(unique(vms_grd_match$OWEM_ID))

```

## All Data

We now match all of the VMS points from all years. We process each year just as one above.

### Import

```{r import all}
# clear wrkspace
rm(list=setdiff(ls(),'grd'))

# load 2009-2019 data. 
# 2009-2019 fish ticket and VMS data as used for Samhouri et al. 2021 PRSB. Available as one rds for each year on kiteworks/VMS_finaldat_08-28-2019/New Pipeline Data/ processed/matched/interpolation
# clean, interpolated vms data from 2020 and 2021 produced by leena riekkola september 2022

# load rds files
vms_all <- purrr::map_df(2009:2021, function(yr){
  read_rds(paste0('/Users/jameal.samhouri/Documents/RAIMBOWT/Processed Data/VMS/interpolation/',yr,"interpolated.rds")) 
})
##here::here('data','processed','matched','interpolation'),'/'

glimpse(vms_all)

```

### Match

```{r}
# convert VMS to spatial object
pt <- proc.time()
vms_all_sf <- vms_all %>%
  st_as_sf(coords=c('LONGITUDE','LATITUDE'),crs=4326) %>% 
  # then, convert to planar projection to match the grid
  st_transform(st_crs(grd))
x<-proc.time()-pt
```

Took `r round(x[3]/60,2)` minutes to do the conversion. Now for the join...

```{r join to grid}
# do the join
pt <- proc.time()
vms_all_grd_match <- vms_all_sf %>%
  st_join(grd)
x<-proc.time()-pt
```

The join took`r round(x[3]/60,2)` minutes.

<!-- ### Add BIAs -->

<!-- We can add whale BIAs into the data as well. They have previously been matched to the 5km grid, courtesy of Sam Woodman. -->

<!-- ```{r add BIAs} -->
<!-- BIA <- read_csv(here::here('data','raw','grid','Grid5km_BIA_overlap.csv'),col_types = cols()) -->
<!-- BIA %<>% select(GRID5KM_ID,BIA_mn,BIA_bm) -->

<!-- vms_all_grd_match %<>% left_join(BIA,by='GRID5KM_ID') -->
<!-- ``` -->

### Save and Write

```{r}
# non-spatial version
glimpse(vms_all_grd_match)
vms_all_grd_match %>% 
  st_set_geometry(NULL) %>% 
  write_rds('/Users/jameal.samhouri/Documents/RAIMBOWT/Processed Data/VMS/vms_all_interpolated_w_grd_3km.rds')
```

## Prep data for DCRB focused analysis

Ok now that we have the VMS data on the 3km grid, we need to prepare it for analysis as in Samhouri et al. 2021 PRSB, with a focus on DCRB fishing effort.

The code below is based on prep_data_for_scenario_df_function.R

Start by choosing which variables to use to filter to fishing locations only.

```{r fishing data prep 1}

vms_all_grd_match <- read_rds('/Users/jameal.samhouri/Documents/RAIMBOWT/Processed Data/VMS/vms_all_interpolated_w_grd_3km.rds')
glimpse(vms_all_grd_match)

vms_all_grd_match <- vms_all_grd_match %>% mutate(
  year = lubridate::year(westcoastdate_notime)
)

# check years included
unique(vms_all_grd_match$year)

#####################################################
# filter the data using the control variables below
# years <- seq(2009,2019,1)

state_agency_code <- "C"
target_rev <- "DCRB"
target_lbs <- "DCRB"
winter_months <- c("November", "December", "January", "February", "March")
removal_types <- c("COMMERCIAL (NON-EFP)", "COMMERCIAL(DIRECT SALES)", "UNKNOWN")
max_speed <- 4.11556
min_speed <- 0


```

Now filter the data based on control variables above: California, DCRB trips only, commercial removal types only, slow vessel speeds. Need to add DEPTH_CATM, but for now ask Blake to filter based on NGDC_M or the depth of each grid cell.

```{r filter fishing data}

# subset the data based on above queries. CA only
dcrb_ca_vms_tix_analysis <- vms_all_grd_match %>%
  filter(agency_code == state_agency_code) %>%
  filter(TARGET_rev == target_rev | TARGET_lbs == target_lbs) %>%
  filter(removal_type_name %in% removal_types) %>%
  #filter(DEPTH_CATM == "0-100m" | DEPTH_CATM == "100-150m") %>%
  #filter(NGDC_M <= 0 & NGDC_M >= -12000) %>% # considered adding 052220. also see "evaluate fishing depths.R". probably unnecessary because NGDC_M refers to centroid depth and we can match depths by 5km grid cell later
  filter(avg_speed_recalc <= max_speed & avg_speed_recalc >= min_speed) %>%
  #filter(is.na(in_port) == TRUE) %>% # only removes ~4000 records
  #filter(port_group_code %in% ports) %>%
  mutate(
         year_month = paste0(lubridate::year(westcoastdate_notime),"_", substr(lubridate::ymd(westcoastdate_notime),6,7)), # substr() ensures month is a 2 digit value
         month = lubridate::month(westcoastdate_notime, label=TRUE, abbr = FALSE),
         month_as_numeric = month(westcoastdate_notime),
         week_of_year = week(westcoastdate_notime),
         day_of_year = yday(westcoastdate_notime),
         season = as.character(ifelse(month %in% winter_months, "Winter", "Spring-Summer")),
         crab_year = ifelse(
           month_as_numeric >= 11, paste0(year,"_",1+year), paste0(year - 1,"_",year)
         )
  ) %>%
  # filter(westcoastdate_notime >= as.Date("2009-11-01") & westcoastdate_notime <= as.Date("2019-08-01")) %>%
  # #filter(month_as_numeric %in% crab_months) %>%
  dplyr::select(
    Rec_ID, VMS_RECNO, drvid,
    westcoastdate_notime, year, crab_year, year_month, month, month_as_numeric, week_of_year, day_of_year, season,
    OWEM_ID, pacfin_port_code, port_group_code, NGDC_M,
    TARGET_lbs, TARGET_rev, DCRB_lbs, DCRB_revenue
    ) # DEPTH_CATM, 

glimpse(dcrb_ca_vms_tix_analysis)

```

Distribute rev and lbs to grid cells. 

```{r distribute $ and lbs to cells}

### create new columns that apportion pings/lbs/$ from each fish ticket based on proportion of pings related to each ticket in each grid cell

# total records per trip (total number of VMS records associated with each fish ticket)
VMSrecords_per_trip <- dcrb_ca_vms_tix_analysis %>%
  group_by(Rec_ID) %>%
  summarise(trip_VMSrecords = n(),
            .groups = 'drop') #%>%
  #filter(trip_VMSrecords > 1)

# add to fish ticket / vms data, make columns vessels, lbs, and $ per VMS location. note that columns with "...." in them, like ""TARGET...." reflect revenue 
dcrb_ca_vms_tix_analysis_TripInfo <- left_join(VMSrecords_per_trip, dcrb_ca_vms_tix_analysis, by="Rec_ID") %>%
  mutate(
    DCRB_lbs_per_VMSlocation = DCRB_lbs/trip_VMSrecords,
    DCRB_rev_per_VMSlocation = DCRB_revenue/trip_VMSrecords,
    Num_DCRB_Vessels_per_VMSlocation = 1/trip_VMSrecords
  )
glimpse(dcrb_ca_vms_tix_analysis_TripInfo)

# could add in depth filter by excluding all NGDC_M values >-1500

# write out for Blake
write_rds(dcrb_ca_vms_tix_analysis_TripInfo, 
          "/Users/jameal.samhouri/Documents/RAIMBOWT/Processed Data/VMS/CA_DCRB_vms_tix_daily_2009_2021_lbs_rev_per_geolocation_nodepthfilter.RDS")

#####################################################

#####################################################

### at long last, make the df we want

### all vessels

start.time <- Sys.time()

con_df_daily_years_5km_CA <- dcrb_ca_vms_tix_analysis_TripInfo %>%
  group_by(year, crab_year, year_month, season, month, month_as_numeric, week_of_year, day_of_year, OWEM_ID) %>%
  summarise(
    DCRB_lbs = sum(DCRB_lbs_per_VMSlocation),
    DCRB_rev =sum(DCRB_rev_per_VMSlocation),
    Num_DCRB_VMS_pings = n(),
    Num_DCRB_Vessels = sum(Num_DCRB_Vessels_per_VMSlocation),
    Num_Unique_DCRB_Vessels = length(unique(as.character(drvid))),
    mean_depth = mean(NGDC_M)
  ) %>%
  ungroup() 
Sys.time() - start.time

glimpse(con_df_daily_years_5km_CA)

### if and when satisfied, write out rds

write_rds(con_df_daily_years_5km_CA, 
          "/Users/jameal.samhouri/Documents/RAIMBOWT/Processed Data/VMS/CA_DCRB_vms_tix_daily_2009_2021_lbs_rev_per_3kmgrid_nodepthfilter.RDS")

```

Did not yet, but could summarise by yr, yr_month, etc