---
title: "Hindcasts of California Dungeness crab fishing activity, 2011-2023"
author: "Brooke Hawkins"
date: "`r Sys.Date()`"
output: html_document
---

## Purpose
* Integrate interpolated VMS and fish ticket data to produce time series, rasters, and maps of fishing activity, landed weight, and revenue.
Deliverable:
* Monthly maps of fishing activity; tables summarizing monthly fishing activity

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prep data

```{r import-libraries}
library(here)
library(readr)
library(tidyverse)
library(sf)
library(rnaturalearth)
library(viridis)

# adjust ggplot theme
theme_replace(axis.text.x=element_text(angle=45, vjust=1, hjust=1),
              axis.ticks.x=element_blank(),
              axis.ticks.y=element_blank())
```
Load the joined, cleaned, interpolated VMS and fish ticket data for specified years. Then load and join the spatial 5km grid.

```{r load-data}
# TODO should yearly filters and summaries be by year or by crab_year?
# choose years of data to load
load_years <- c(2014, 2015, 2016)

# load VMS and fish ticket data
file_name <- here('Confidential', 'processed', 'matched', 'interpolation',
                  paste0(load_years[1],'_interpolated.rds'))
# read the first year of data
vms_df <- read_rds(file_name)
# read the remaining years of data
if (length(load_years) > 1) {
  for (ly in load_years[2:length(load_years)]) {
    file_name <- here('Confidential', 'processed', 'matched', 'interpolation',
                      paste0(ly,'_interpolated.rds'))
    vms_df <- bind_rows(vms_df, read_rds(file_name))
  }
}

# load 5km grid shape file
grid_5km <- read_sf(here('GIS_layers', 'master_5km_grid_tmer.shp'))
# join VMS and fish ticket data to 5km grid
vms_df <- vms_df %>%
  st_as_sf(coords = c('LON', 'LAT'), crs=4326) %>%
  st_transform(st_crs(grid_5km)) %>%
  st_join(grid_5km) %>%
  st_set_geometry(NULL)

# clean up
remove(file_name, ly)
```

Some commonly used acronyms in variable naming are:

* `dcrb` dungeness crab
* `rev` revenue
* `lbs` pounds
* `VMS` vessel monitoring system

```{r transform-data}
# define filters
target_rev <- "DCRB"         # revenue target
target_lbs <- "DCRB"         # landings target
state_agency_code <- "C"     # landings brought to California ports
min_depth <- 0               # minimum depth in meters
max_depth <- -150            # maximum depth in meters
min_speed <- 0               # minimum speed in m/s
max_speed <- 4.11556         # maximum speed in m/s (4.11556 m/s = 8 knots)
crab_year_start <- 11        # month defines start of crab year
winter_months <- c("November", "December", "January", "February", "March") # determine Winter or Spring-Summer season

# transform data
dcrb_df <- vms_df %>%
  # add temporal columns
  mutate(
    year = lubridate::year(westcoastdate_notime),
    month = lubridate::month(westcoastdate_notime, label=TRUE, abbr=FALSE),
    month_numeric = month(westcoastdate_notime),
    year_month = paste0(lubridate::year(westcoastdate_notime),"_", substr(lubridate::ymd(westcoastdate_notime), 6, 7)),
    crab_year = ifelse(month_numeric >= crab_year_start, paste0(year, "_", 1+year), paste0(year-1, "_", year)),
    season = as.character(ifelse(month %in% winter_months, "Winter", "Spring-Summer")),
    week_of_year = week(westcoastdate_notime),
    day_of_year = yday(westcoastdate_notime)
  ) %>%
  # apply filters
  filter(TARGET_rev == target_rev | TARGET_lbs == target_lbs) %>%
  filter(agency_code == state_agency_code) %>%
  filter(NGDC_M <= min_depth & NGDC_M >= max_depth) %>% # TODO should this be for WM_NGDC_M instead?
  filter(avg_speed_recalc <= max_speed & avg_speed_recalc >= min_speed) %>% 
  # select columns
  dplyr::select(
    # identifiers
    Rec_ID,           # fish ticket ID
    VMS_RECNO,        # VMS ping ID
    drvid,            # vessel ID
    pacfin_port_code, # port ID
    port_group_code,  # port group ID
    GRID5KM_ID,       # grid cell ID
    # temporal fields
    westcoastdate,
    westcoastdate_notime,
    crab_year, 
    season,
    year,
    year_month,
    month, 
    month_numeric,
    week_of_year,
    day_of_year,
    # vessel length
    FINAL_LENGTH,
    # dungeness crab fields
    DCRB_lbs,
    DCRB_revenue
  )

# count total records, trips and years
n_records <- nrow(vms_df)
n_trips   <- n_distinct(vms_df$Rec_ID)
n_years   <- n_distinct(lubridate::year(vms_df$westcoastdate_notime))

# take a peek at the resulting dataframe
glimpse(dcrb_df)
```

The dungeness crab dataframe has `r n_records` records (interpolated VMS pings), `r n_trips` trips (fish tickets), across `r n_years` years.

```{r}
# clean up
remove(vms_df, n_records, n_trips, n_years)
```

```{r gridded-summary}
# summarize trip-level attributes
trip_df <- dcrb_df %>%
  group_by(Rec_ID) %>%
  summarise(trip_n_vms_records = n(),
            trip_start = min(westcoastdate),
            trip_end = max(westcoastdate))

# join trip-level attributes back to dungeness crab dataframe
dcrb_trip_df <- left_join(dcrb_df, trip_df, by="Rec_ID") %>%
  mutate(
    dcrb_lbs_per_vms_record     = DCRB_lbs / trip_n_vms_records,
    dcrb_rev_per_vms_record     = DCRB_revenue / trip_n_vms_records,
    dcrb_vessels_per_vms_record = 1 / trip_n_vms_records
  )

# summarize grid cell and month-level attributes
confidential_df <- dcrb_df %>%
  group_by(GRID5KM_ID, year_month) %>%
  summarise(n_unique_vessels = n_distinct(drvid),
            # set confidential_flag TRUE for confidential, FALSE for non-confidential
            confidential_flag = (n_distinct(drvid) < 3), .groups="keep") %>%
  ungroup()

# join grid cell and month-level attributes back to dungeness crab dataframe
dcrb_confidential_df <- dcrb_trip_df %>%
  left_join(confidential_df, by=join_by(GRID5KM_ID, year_month))

# create monthly gridded summary
dcrb_5km_summary_df <- dcrb_confidential_df %>%
  group_by(
    # grid cell ID, makes this a gridded summary
    GRID5KM_ID,
    # temporal columns, makes this a monthly summary
    year,
    crab_year,
    year_month,
    season,
    month,
    month_numeric
  ) %>%
  # sum landings, revenue; count unique vessels and VMS records; determine confidentiality filter
  summarise(
    dcrb_lbs = sum(dcrb_lbs_per_vms_record),
    dcrb_rev = sum(dcrb_rev_per_vms_record),
    n_vms_records = n(),
    n_unique_vessels = n_distinct(drvid),
    # set confidential_flag TRUE for confidential, FALSE for non-confidential
    confidential_flag = max(confidential_flag) == 1,
    .groups = "keep"
  ) %>% 
  ungroup()

# write monthly gridded summary
write.csv(file=here("Confidential", "tables", "dcrb_5km_summary_df.csv"), x=dcrb_5km_summary_df)
```

```{r}
# clean up
remove(trip_df, dcrb_trip_df, confidential_df)
```


## Create outputs

If a grid cell in a given month has < 3 unique vessels, it cannot be included in non-confidential maps summarizing the data.

```{r}
# set whether to use the non-confidential or confidential gridded summary for output
confidential_output_flag <- TRUE

if (confidential_output_flag) {
  # if confidential output is ok, then don't remove any VMS records
  dcrb_5km_output_df <- dcrb_5km_summary_df
} else {
  # if non-confidential output is needed, then remove any VMS records where the grid cell and month had < 3 unique vessels, where confidential_flag is FALSE
 dcrb_5km_output_df <- dcrb_5km_summary_df %>% filter(!confidential_flag)
}
```

Regardless of whether you're using confidential or non-confidential output, report how many VMS records are filtered out by the confidentiality flag.

```{r confidentiality-stats}
# calculate the % of VMS records dropped by confidentiality filter, for the entire time period
numerator_drop_confidential <- sum(dcrb_confidential_df$confidential_flag)
denominator_drop_confidential <- dim(dcrb_confidential_df)[1]
percent_drop_confidential <- substr(as.character(numerator_drop_confidential / denominator_drop_confidential * 100), 1, 5)

# repeat for each month
confidential_monthly_table <- dcrb_confidential_df %>% group_by(year, month_numeric) %>%
  summarise(confidential_vms_records = sum(confidential_flag),
            total_vms_records = n(),
            percent_vms_records_confidential = trunc(confidential_vms_records / total_vms_records * 10000) / 100,
            .groups = 'keep') %>%
  ungroup() %>%
  arrange(year, month_numeric)

# write monthly result
write.csv(file = here("Confidential", "tables", "table_confidentiality_monthly.csv"),
          x = confidential_monthly_table)

# repeat for each year
confidential_yearly_table <- dcrb_confidential_df %>% group_by(year) %>%
  summarise(confidential_vms_records = sum(confidential_flag),
            total_vms_records = n(),
            percent_vms_records_confidential = trunc(confidential_vms_records / total_vms_records * 10000) / 100,
            .groups = 'keep') %>%
  ungroup() %>%
  arrange(year)

# write yearly result
write.csv(file = here("Confidential", "tables", "table_confidentiality_yearly.csv"),
          x = confidential_yearly_table)
```

The non-confidential data has `r numerator_drop_confidential` (`r percent_drop_confidential`%) fewer records (VMS pings), where grid cells had <3 unique vessels in a given month.

```{r}
# clean up
remove(numerator_drop_confidential, denominator_drop_confidential, percent_drop_confidential)
```


```{r tables}
# create monthly summary of gridded fishing activity
fishing_activity_monthly_table <- dcrb_5km_output_df %>%
  group_by(year, month_numeric) %>%
  summarise(
    n_grid_cells = n(),
    min_vms_records = min(n_vms_records),
    median_vms_records = median(n_vms_records),
    mean_vms_records = mean(n_vms_records),
    max_vms_records = max(n_vms_records),
    iqr_vms_records = IQR(n_vms_records),
    min_unique_vessels = min(n_unique_vessels),
    median_unique_vessels = median(n_unique_vessels),
    mean_unique_vessels = mean(n_unique_vessels),
    max_unique_vessels = max(n_unique_vessels),
    iqr_unique_vessels = IQR(n_unique_vessels),
    .groups = 'keep'
  ) %>%
  arrange(year, month_numeric)
# write result
write.csv(file=here("Confidential", "tables", "table_fishing_activity_monthly.csv"),
          x=fishing_activity_monthly_table)

# repeat for yearly
fishing_activity_yearly_table <- dcrb_5km_output_df %>%
  group_by(year) %>%
  summarise(
    n_grid_cell_months = n(),
    n_grid_cells = n_distinct(GRID5KM_ID),
    min_vms_records = min(n_vms_records),
    median_vms_records = median(n_vms_records),
    mean_vms_records = mean(n_vms_records),
    max_vms_records = max(n_vms_records),
    iqr_vms_records = IQR(n_vms_records),
    min_unique_vessels = min(n_unique_vessels),
    median_unique_vessels = median(n_unique_vessels),
    mean_unique_vessels = mean(n_unique_vessels),
    max_unique_vessels = max(n_unique_vessels),
    iqr_unique_vessels = IQR(n_unique_vessels),
    .groups = 'keep'
  ) %>%
  arrange(year)
# write result
write.csv(file=here("Confidential", "tables", "table_fishing_activity_yearly.csv"),
          x=fishing_activity_yearly_table)
```

```{r maps}
# map fishing activity, apply to each month of data
#   data_sf: sf object
#   data_col: vector from data_sf to log-transform and map
#   log_transform: boolean to log-transform data_col on map
#   fill_range: range for the fill of data_col vector
#   bbox: bounding box to use for map
#   title: map title
#   data_col_name: name of data_col vector, used to label map legend and name output PNG
#   png_suffix: optional string, used to name output PNG
#   background_sf: sf object
fishing_activity_map <- function (data_sf, data_col, bbox, log_transform, fill_range, title, data_col_name, png_suffix='', background_sf) {
  # log-transform data, if applicable
  if (log_transform) {
    data_col <- log10(data_col)
    fill_range <- log10(fill_range)
    data_col_name <- paste0("log10_", data_col_name)
  }
  # create map
  map_output <- ggplot() +
    # plot background
    geom_sf(data=background_sf, fill='gray80') +
    # plot log-10 transformed data
    geom_sf(data=data_sf, aes(fill=data_col)) +
    # label plot
    labs(title=title,
         fill=paste(data_col_name, "per 5x5 cell")) +
    # set bounding box
    xlim(bbox[1], bbox[3]) + ylim(bbox[2], bbox[4]) +
    # set fill color and legend limits
    scale_fill_viridis(limits=fill_range) +
    # adjust legend position
    theme(legend.position="bottom")
  # save the output map
  ggsave(filename=here("Confidential", "figures", paste0("map_", 
                                                         data_col_name,
                                                         "_",
                                                         png_suffix, 
                                                         ".png")))
  # return the output map
  return(map_output)
}

# load west coast states simple features object for background
west_coast_states <- rnaturalearth::ne_states(country='United States of America', returnclass='sf') %>% 
  filter(name %in% c('California','Oregon','Washington','Nevada')) %>% 
  st_transform(st_crs(grid_5km))

# join geometry from 5km grid to gridded fishing summary
dcrb_5km_output_sf <- dcrb_5km_output_df %>%
  left_join(grid_5km, by=join_by(GRID5KM_ID)) %>%
  st_as_sf()

# set bounding box
bbox <- st_bbox(dcrb_5km_output_sf)

# set limits for min and max fill for # VMS pings and # unique vessels
vms_fill_range <- range(dcrb_5km_output_sf$n_vms_records)
vessels_fill_range <- range(dcrb_5km_output_sf$n_unique_vessels)

# get unique and ordered year, year_month and month from gridded fishing summary
month_df <- dcrb_5km_output_sf %>% 
  st_drop_geometry() %>% 
  select(year, year_month, month) %>% 
  unique() %>% 
  arrange(year_month)

# iterate over months
for (ym in month_df$year_month) {
  # print progress
  print(ym)
  # get title as a combination of year and month
  y <- month_df$year[which(month_df$year_month==ym)]
  m <- month_df$month[which(month_df$year_month==ym)]
  title <- paste(y, m)
  # filter data for given month
  ym_data_sf <- dcrb_5km_output_sf %>% filter(year_month == ym)
  # make and save map of # VMS pings
  vms_map <- fishing_activity_map(data_sf = ym_data_sf,
                                  data_col = ym_data_sf$n_vms_records,
                                  bbox = bbox,
                                  log_transform = TRUE,
                                  fill_range = vms_fill_range,
                                  title = title,
                                  data_col_name = 'n_vms_records',
                                  png_suffix = ym,
                                  background_sf = west_coast_states)
  # make and save map of # unique vessels
  vessels_map <- fishing_activity_map(data_sf = ym_data_sf,
                                      data_col = ym_data_sf$n_unique_vessels,
                                      bbox = bbox,
                                      log_transform = FALSE,
                                      fill_range = vessels_fill_range,
                                      title = title,
                                      data_col_name = 'n_unique_vessels',
                                      png_suffix = ym,
                                      background_sf = west_coast_states)
}

# view last one of each map
vms_map
vessels_map
```
```{r}
# clean up
remove(ym, y, m, ym_data_sf, month_df, bbox, title, vessels_fill_range, vms_fill_range)
```

```{r time-series}
# TODO does this section need a confidentiality filter, since not spatial?

# this uses the confidential data, using non-gridded summary to not double count unique vessels
time_series_df <- dcrb_confidential_df %>%
  group_by(crab_year, year, year_month, month) %>%
  summarise(total_vms_records = n(),
            n_unique_vessels = n_distinct(drvid),
            .groups = 'keep')

# plot # VMS pings per month as bar chart
vms_time_series <- time_series_df %>% 
  ggplot() + 
  geom_col(aes(x=year_month, y=total_vms_records, fill=crab_year))
ggsave(filename=here("Confidential", "figures", "time_series_vms.png"))
vms_time_series

# plot # unique vessels per month as bar chart
vessels_time_series <- time_series_df %>% 
  ggplot() + 
  geom_col(aes(x=year_month, y=n_unique_vessels, fill=crab_year))
ggsave(filename=here("Confidential", "figures", "time_series_vessels.png"))
vessels_time_series

# plot # VMS pings per month as heatmap
vms_heatmap <- time_series_df %>%
  ggplot() +
  geom_tile(aes(x=month, y=as.factor(year), fill=total_vms_records))
ggsave(filename=here("Confidential", "figures", "heatmap_vms.png"))
vms_heatmap

# plot # unique vessels per month as heatmap
vessels_heatmap <- time_series_df %>%
  ggplot() +
  geom_tile(aes(x=month, y=as.factor(year), fill=n_unique_vessels))
ggsave(filename=here("Confidential", "figures", "heatmap_vessels.png"))
vessels_heatmap
```

