---
title: "Hindcasts of California Dungeness crab fishing activity, 2011-2023"
author: "Brooke Hawkins"
date: "`r Sys.Date()`"
output: html_document
---

## Purpose
* Integrate interpolated VMS and fish ticket data to produce time series, rasters, and maps of fishing activity, landed weight, and revenue.
Deliverable:
* Monthly maps of fishing activity; tables summarizing monthly fishing activity

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prep data

```{r import-libraries}
library(tidyverse)
library(here)
library(fredr)
library(sf)
library(rnaturalearth)
library(viridis)
library(magick)

# adjust ggplot theme
theme_replace(axis.text.x=element_text(angle=45, vjust=1, hjust=1),
              axis.ticks.x=element_blank(),
              axis.ticks.y=element_blank())
```
Load the joined, cleaned, interpolated VMS and fish ticket data for specified years. Also load the cleaned fish ticket data for specified years, which will be used to check how representative the VMS dataset is of all dungeness crab landings. Then load and join the spatial 5km grid, which will be used to generate spatial summaries of fishing activity. Joining the spatial 5km gird is one of the slowest parts of the script.

```{r load-data}
# TODO should yearly filters and summaries be by year or by crab_year?
# choose years of data to load
load_years <- 2021:2023

# load VMS and fish ticket data
file_name <- here('Confidential', 'processed', 'matched', 'interpolation',
                  paste0(load_years[1],'_interpolated.rds'))
# read the first year of data
vms_df <- read_rds(file_name)
# read the remaining years of data
if (length(load_years) > 1) {
  for (ly in load_years[2:length(load_years)]) {
    file_name <- here('Confidential', 'processed', 'matched', 'interpolation',
                      paste0(ly,'_interpolated.rds'))
    vms_df <- bind_rows(vms_df, read_rds(file_name))
  }
}

# load fish ticket data before it was joined to VMS data, used to check VMS representativeness
file_name <- here('Confidential', 'processed', 'fish tickets',
                  paste0(load_years[1],'fishtix_vlengths_withFTID.rds'))

# read the first year of data
ticket_df <- read_rds(file_name)
# read the remaining years of data
if (length(load_years) > 1) {
  for (ly in load_years[2:length(load_years)]) {
    file_name <- here('Confidential', 'processed', 'fish tickets',
                      paste0(ly,'fishtix_vlengths_withFTID.rds'))
    ticket_df <- bind_rows(ticket_df, read_rds(file_name))
  }
}

# clean up
remove(file_name, ly)

# load 5km grid shape file
grid_5km <- read_sf(here('GIS_layers', 'master_5km_grid_tmer.shp'))
# join VMS and fish ticket data to 5km grid
vms_df <- vms_df %>%
  st_as_sf(coords = c('LON', 'LAT'), crs=4326) %>%
  st_transform(st_crs(grid_5km)) %>%
  st_join(grid_5km) %>%
  st_set_geometry(NULL)
```

Adjust revenue for inflation using Fred GDP data, adapted from R code by Erin Steiner. This product uses the FRED® API but is not endorsed or certified by the Federal Reserve Bank of St. Louis.

Pre-requisite: Create a FRED account and API key. See https://fred.stlouisfed.org/docs/api/api_key.html.

```{r inflation-adjustment}
# insert your FRED API key
fredr_set_key('') 

# download the quarterly inflation adjustments from Fred from 1985 to present
fred_gdpdefl <- fredr(
    series_id = "GDPDEF", 
    observation_start = as.Date(paste0(min(load_years), "-01-01"))
  )

# TODO should a different reference year be used? Currently set to year the script is run.
# update by replacing the denominator for mutating inflation_adjustment_factor

# generate the mean annual deflators based on the quarterly values
gdp_defl <- mutate(fred_gdpdefl, year = year(date)) %>%
  group_by(year) %>%
  summarize(defl = mean(value), .groups = 'drop') %>%
  mutate(inflation_adjustment_factor = defl / defl[year == max(year)]) %>%
  select(year, inflation_adjustment_factor)

# write inflation adjustment factors
write.csv(file = here("Confidential", "hindcast_output", "tables", "inflation_adjustment.csv"), x = gdp_defl)

# plot inflation adjustment factors
gdp_defl %>% ggplot(aes(x = as.factor(year), y = inflation_adjustment_factor)) + geom_point()
ggsave(here("Confidential", "hindcast_output", "figures", "inflation_adjustment.png"))

# clean up
rm(fred_gdpdefl)
```

The reference year for inflation adjustment was `r max(gdp_defl$year)`.

Some commonly used acronyms in variable naming are:

* `dcrb` dungeness crab
* `rev` revenue
* `lbs` pounds
* `VMS` vessel monitoring system
* `afi` adjusted for inflation

```{r transform-vms-data}
# define filters
target_rev <- "DCRB"         # revenue target
target_lbs <- "DCRB"         # landings target
state_agency_code <- "C"     # landings brought to California ports
min_depth <- 0               # minimum depth in meters
max_depth <- -150            # maximum depth in meters
min_speed <- 0               # minimum speed in m/s
max_speed <- 4.11556         # maximum speed in m/s (4.11556 m/s = 8 knots)
crab_year_start <- 11        # month defines start of crab year
winter_months <- c("November", "December", "January", "February", "March") # determine Winter or Spring-Summer season

# transform VMS and fish ticket data
dcrb_df <- vms_df %>%
  # add temporal columns
  mutate(
    year = lubridate::year(westcoastdate_notime),
    month = lubridate::month(westcoastdate_notime, label=TRUE, abbr=FALSE),
    month_numeric = month(westcoastdate_notime),
    year_month = paste0(lubridate::year(westcoastdate_notime),"_", substr(lubridate::ymd(westcoastdate_notime), 6, 7)),
    crab_year = ifelse(month_numeric >= crab_year_start, paste0(year, "_", 1+year), paste0(year-1, "_", year)),
    season = as.character(ifelse(month %in% winter_months, "Winter", "Spring-Summer")),
    week_of_year = week(westcoastdate_notime),
    day_of_year = yday(westcoastdate_notime)
  ) %>%
  # apply filters
  filter(TARGET_rev == target_rev | TARGET_lbs == target_lbs) %>%
  filter(agency_code == state_agency_code) %>%
  filter(NGDC_M <= min_depth & NGDC_M >= max_depth) %>% # TODO should this be for WM_NGDC_M instead?
  filter(avg_speed_recalc <= max_speed & avg_speed_recalc >= min_speed) %>% 
  # join inflation adjustment factor and adjust revenue
  left_join(gdp_defl, by = join_by(year)) %>%
  mutate(DCRB_revenue_afi = DCRB_revenue / inflation_adjustment_factor) %>%
  # select columns
  dplyr::select(
    # identifiers
    Rec_ID,           # fish ticket ID
    VMS_RECNO,        # VMS ping ID
    drvid,            # vessel ID
    pacfin_port_code, # port ID
    port_group_code,  # port group ID
    GRID5KM_ID,       # grid cell ID
    # temporal fields
    westcoastdate,
    westcoastdate_notime,
    crab_year, 
    season,
    year,
    year_month,
    month, 
    month_numeric,
    week_of_year,
    day_of_year,
    # vessel length
    FINAL_LENGTH,
    # dungeness crab fields
    DCRB_lbs,
    DCRB_revenue_afi
  )

# count total records, trips and years
n_records <- nrow(dcrb_df)
n_trips   <- n_distinct(dcrb_df$Rec_ID)
n_years   <- n_distinct(lubridate::year(dcrb_df$westcoastdate_notime))

# take a peek at the resulting dataframe
glimpse(dcrb_df)
```

The dungeness crab VMS dataframe has `r n_records` records (interpolated VMS pings), `r n_trips` trips (fish tickets), across `r n_years` years.

```{r}
# clean up
remove(vms_df, n_records, n_trips, n_years)
```

Create a monthly gridded summary of fishing activity. 

Trip-level attributes are added at this step:

* `trip_n_vms_records`: the number of VMS records associated with a fishing trip (fish ticket)

VMS-record attributes are then calculated using the trip-level attribute:

* `dcrb_lbs_per_vms_record`: the total landings for a fishing trip, divided by the number of VMS records associated with that trip (e.g. a fish ticket with 1200 lbs landed and 20 VMS records would have 60 lbs per VMS record)
* `dcrb_rev_per_vms_record`: the total revenue for a fishing trip, divided by the number of VMS records associated with that trip (e.g. a fish ticket with $4000 revenue and 20 VMS pings would have \$200 revenue per VMS record)

Then, the activity is summarized at a monthly gridded level:

* `dcrb_lbs`: the total landings per VMS record in a given grid cell in a given month
* `dcrb_rev`: the total revenue per VMS record in a given grid cell in a given month
* `n_vms_records`: the number of VMS records in a given grid cell in a given month
* `n_unique_vessels`: the number of unique vessels in a given grid cell in a given month
* `confidential_flag`: determines whether the grid cell month has <3 unique vessels; in which case, the data in this grid cell needs to be excluded from non-confidential reports

```{r gridded-summary}
# summarize trip-level attributes
trip_df <- dcrb_df %>%
  group_by(Rec_ID) %>%
  summarise(trip_n_vms_records = n())

# join trip-level attributes back to dungeness crab dataframe
dcrb_trip_df <- left_join(dcrb_df, trip_df, by="Rec_ID") %>%
  mutate(
    dcrb_lbs_per_vms_record     = DCRB_lbs / trip_n_vms_records,
    dcrb_rev_per_vms_record     = DCRB_revenue_afi / trip_n_vms_records,
    dcrb_vessels_per_vms_record = 1 / trip_n_vms_records
  )

# summarize grid cell and month-level attributes
confidential_df <- dcrb_df %>%
  group_by(GRID5KM_ID, year_month) %>%
  summarise(n_unique_vessels = n_distinct(drvid),
            # set confidential_flag TRUE for confidential, FALSE for non-confidential
            confidential_flag = (n_distinct(drvid) < 3), .groups="keep") %>%
  ungroup()

# join grid cell and month-level attributes back to dungeness crab dataframe
dcrb_confidential_df <- dcrb_trip_df %>%
  left_join(confidential_df, by=join_by(GRID5KM_ID, year_month))

# create monthly gridded summary
dcrb_5km_summary_df <- dcrb_confidential_df %>%
  group_by(
    # grid cell ID, makes this a gridded summary
    GRID5KM_ID,
    # temporal columns, makes this a monthly summary
    year,
    crab_year,
    year_month,
    season,
    month,
    month_numeric
  ) %>%
  # sum landings, revenue; count unique vessels and VMS records; determine confidentiality filter
  summarise(
    dcrb_lbs = sum(dcrb_lbs_per_vms_record),
    dcrb_rev = sum(dcrb_rev_per_vms_record),
    dcrb_vessels = sum(dcrb_vessels_per_vms_record),
    n_vms_records = n(),
    n_unique_vessels = n_distinct(drvid),
    # set confidential_flag TRUE for confidential, FALSE for non-confidential
    confidential_flag = max(confidential_flag) == 1,
    .groups = "keep"
  ) %>% 
  ungroup()

# write monthly gridded summary
write.csv(file=here("Confidential", "hindcast_output", "tables", "dcrb_5km_summary_df.csv"), x=dcrb_5km_summary_df)
```

```{r}
# clean up
remove(trip_df, dcrb_trip_df, confidential_df)
```

Repeat the data transformation steps for fish ticket data, which will be used to later on to evaluate the representativeness of the VMS data for dungeness crab fishing activity.

```{r transform-ticket-data}
# select distinct fish ticket IDs from filtered VMS dataframe
dcrb_vms_rec_ids <- dcrb_confidential_df %>% 
  distinct(Rec_ID) %>% 
  mutate(vms_represented_flag = TRUE)

# apply same filters as VMS dataframe (except for depth and speed filters, which are from VMS data)
dcrb_ticket_df <- ticket_df %>%
  # add temporal columns
  mutate(
    year = lubridate::year(date),
    month = lubridate::month(date, label=TRUE, abbr=FALSE),
    month_numeric = month(date),
    year_month = paste0(lubridate::year(date),"_", substr(lubridate::ymd(date), 6, 7)),
    crab_year = ifelse(month_numeric >= crab_year_start, paste0(year, "_", 1+year), paste0(year-1, "_", year)),
    season = as.character(ifelse(month %in% winter_months, "Winter", "Spring-Summer")),
    week_of_year = week(date),
    day_of_year = yday(date)
  ) %>%
  # apply filters
  filter(TARGET_rev == target_rev | TARGET_lbs == target_lbs) %>%
  filter(agency_code == state_agency_code) %>%
  mutate(
    year_month = paste0(lubridate::year(date),"_", substr(lubridate::ymd(date),6,7)),
    month_as_numeric = month(date),
    month_as_date = round_date(date, unit='month'),
    season = as.character(ifelse(month %in% winter_months, "Winter", "Spring-Summer")),
    crab_year = ifelse(
      month_as_numeric >= 11, paste0(year, "_", 1+year), paste0(year-1, "_", year)
    )
  ) %>%
  # join inflation adjustment factor and adjust revenue
  left_join(gdp_defl, by = join_by(year)) %>%
  mutate(DCRB_revenue_afi = DCRB_revenue / inflation_adjustment_factor) %>%
  # select columns
  dplyr::select(
    # identifiers
    Rec_ID,           # fish ticket ID
    drvid,            # vessel ID
    pacfin_port_code, # port ID
    port_group_code,  # port group ID
    # temporal fields
    date,
    crab_year,
    season,
    year,
    year_month,
    month,
    month_numeric,
    week_of_year,
    day_of_year,
    # vessel length
    FINAL_LENGTH,
    # dungeness crab fields
    DCRB_lbs,
    DCRB_revenue_afi
  ) %>%
  # add column to ticket dataframe for whether that ticket was represented by VMS data
  left_join(dcrb_vms_rec_ids, by = join_by(Rec_ID)) %>%
  mutate(vms_represented_flag = if_else(is.na(vms_represented_flag), FALSE, TRUE),
         vms_represented_drvid = if_else(vms_represented_flag, drvid, NA))

# take a peek at the resulting dataframe
glimpse(dcrb_ticket_df)
```

```{r}
# clean up
remove(ticket_df)
```

## Create outputs

Output files will be written to `Confidential/hindcast_output`, into three subdirectories: `figures`, `maps`, and `tables`.

If a grid cell in a given month has < 3 unique vessels, it cannot be included in non-confidential maps summarizing the data.

```{r decide-on-confidential-output}
# set whether to use the non-confidential or confidential gridded summary for output
confidential_output_flag <- TRUE

if (confidential_output_flag) {
  # if confidential output is ok, then don't remove any VMS records
  dcrb_5km_output_df <- dcrb_5km_summary_df
} else {
  # if non-confidential output is needed, then remove any VMS records where the grid cell and month had < 3 unique vessels, where confidential_flag is FALSE
 dcrb_5km_output_df <- dcrb_5km_summary_df %>% filter(!confidential_flag)
}
```

Regardless of whether you're using confidential or non-confidential output, report how many VMS records are filtered out by the confidentiality flag.

```{r confidentiality-table}
# calculate the % of VMS records dropped by confidentiality filter, for the entire time period
numerator_drop_confidential <- sum(dcrb_confidential_df$confidential_flag)
denominator_drop_confidential <- dim(dcrb_confidential_df)[1]
percent_drop_confidential <- substr(as.character(numerator_drop_confidential / denominator_drop_confidential * 100), 1, 5)

# repeat for each month
confidential_monthly_table <- dcrb_confidential_df %>% group_by(year, month_numeric) %>%
  summarise(confidential_vms_records = sum(confidential_flag),
            total_vms_records = n(),
            percent_vms_records_confidential = trunc(confidential_vms_records / total_vms_records * 10000) / 100,
            .groups = 'keep') %>%
  ungroup() %>%
  arrange(year, month_numeric)

# write monthly result
write.csv(file = here("Confidential", "hindcast_output", "tables", "table_confidentiality_monthly.csv"),
          x = confidential_monthly_table)

# repeat for each year
confidential_yearly_table <- dcrb_confidential_df %>% group_by(year) %>%
  summarise(confidential_vms_records = sum(confidential_flag),
            total_vms_records = n(),
            percent_vms_records_confidential = trunc(confidential_vms_records / total_vms_records * 10000) / 100,
            .groups = 'keep') %>%
  ungroup() %>%
  arrange(year)

# write yearly result
write.csv(file = here("Confidential", "hindcast_output", "tables", "table_confidentiality_yearly.csv"),
          x = confidential_yearly_table)
```

The non-confidential data has `r numerator_drop_confidential` (`r percent_drop_confidential`%) fewer records (VMS pings), where grid cells had <3 unique vessels in a given month.

```{r}
# clean up
remove(numerator_drop_confidential, denominator_drop_confidential, percent_drop_confidential)
```

```{r representativeness-tables}
# TODO use inflation-adjusted revenue

# check monthly VMS representativeness for tickets, landings, revenue and vessels
representativeness_monthly_table <- dcrb_ticket_df %>%
  group_by(year, month_numeric) %>%
  summarise(
    # check for tickets
    vms_represented_tickets          = sum(vms_represented_flag),
    total_tickets                    = n(),
    percent_vms_represented_tickets  = trunc(vms_represented_tickets / total_tickets * 10000) / 100,
    # repeat for landings
    vms_represented_landings         = sum(vms_represented_flag * DCRB_lbs),
    total_landings                   = sum(DCRB_lbs),
    percent_vms_represented_landings = trunc(vms_represented_landings / total_landings * 10000) / 100,
    # repeat for revenue
    vms_represented_revenue          = sum(vms_represented_flag * DCRB_revenue_afi),
    total_revenue                    = sum(DCRB_revenue_afi),
    percent_vms_represented_revenue  = trunc(vms_represented_revenue / total_revenue * 10000) / 100,
    # repeat for vessels
    vms_represented_vessels          = n_distinct(vms_represented_drvid, na.rm=TRUE),
    total_vessels                    = n_distinct(drvid),
    percent_vms_represented_vessels  = trunc(vms_represented_vessels / total_vessels * 10000) / 100,
    .groups = 'keep'
  ) %>%
  ungroup() %>%
  arrange(year, month_numeric)

# write monthly result
write.csv(file = here("Confidential", "hindcast_output", "tables", "table_vms_representativeness_monthly.csv"),
          x = representativeness_monthly_table)

# repeat check yearly
representativeness_yearly_table <- dcrb_ticket_df %>%
  group_by(year) %>%
  summarise(
    # check for tickets
    vms_represented_tickets          = sum(vms_represented_flag),
    total_tickets                    = n(),
    percent_vms_represented_tickets  = trunc(vms_represented_tickets / total_tickets * 10000) / 100,
    # repeat for landings
    vms_represented_landings         = sum(vms_represented_flag * DCRB_lbs),
    total_landings                   = sum(DCRB_lbs),
    percent_vms_represented_landings = trunc(vms_represented_landings / total_landings * 10000) / 100,
    # repeat for revenue
    vms_represented_revenue          = sum(vms_represented_flag * DCRB_revenue_afi),
    total_revenue                    = sum(DCRB_revenue_afi),
    percent_vms_represented_revenue  = trunc(vms_represented_revenue / total_revenue * 10000) / 100,
    # repeat for vessels
    vms_represented_vessels          = n_distinct(vms_represented_drvid, na.rm=TRUE),
    total_vessels                    = n_distinct(drvid),
    percent_vms_represented_vessels  = trunc(vms_represented_vessels / total_vessels * 10000) / 100,
    .groups = 'keep'
  ) %>%
  ungroup() %>%
  arrange(year)

# write yearly result
write.csv(file = here("Confidential", "hindcast_output", "tables", "table_vms_representativeness_yearly.csv"),
          x = representativeness_yearly_table)
```

```{r representativeness-time-series}
# plot % represented by VMS for all four metrics, yearly
vms_representativeness_time_series_yearly <- representativeness_yearly_table %>%
  select(year,
         contains("percent_vms_represented")) %>%
  pivot_longer(cols = contains("percent_vms_represented"),
               names_to = "metric",
               names_prefix = "percent_vms_represented_",
               values_to ="percent_vms_represented") %>%
  ggplot(aes(x = as.factor(year), y = percent_vms_represented)) +
  geom_col() +
  xlab("year") +
  facet_grid(metric ~ ., scales = "free_y")
ggsave(filename=here("Confidential", "hindcast_output", "figures", "time_series_representativeness_yearly.png"))
vms_representativeness_time_series_yearly

# plot % represented by VMS for all four metrics, monthly
vms_representativeness_time_series_monthly <- representativeness_monthly_table %>%
  mutate(year_month = paste0(year,"_", ifelse(nchar(month_numeric)>1, month_numeric, paste0(0, month_numeric)))) %>%
  select(year_month,
         contains("percent_vms_represented")) %>%
  pivot_longer(cols = contains("percent_vms_represented"),
               names_to = "metric",
               names_prefix = "percent_vms_represented_",
               values_to ="percent_vms_represented") %>%
  ggplot(aes(x = year_month, y = percent_vms_represented)) +
  geom_col() +
  facet_grid(metric ~ ., scales = "free_y")
ggsave(filename=here("Confidential", "hindcast_output", "figures", "time_series_representativeness_monthly.png"))
vms_representativeness_time_series_monthly
```
Repeat the representativeness table summaries and visualizations, now stratified by vessel size. Vessel size will be grouped into small vessels (with vessel length less than or equal to `vessel_size_cutoff` feet) and large vessels (greater than `vessel_size_cutoff` feet).

```{r size-stratified-representativeness-tables}

# set a cutoff (in feet) for vessel size stratification in VMS representativeness tables and figures
vessel_size_cutoff <- 40

# check monthly VMS representativeness for tickets, landings, revenue and vessels
stratified_representativeness_monthly_table <- dcrb_ticket_df %>%
  mutate(vessel_size_above_cutoff = FINAL_LENGTH > vessel_size_cutoff) %>%
  group_by(year, month_numeric, vessel_size_above_cutoff) %>%
  summarise(
    # check for tickets
    vms_represented_tickets          = sum(vms_represented_flag),
    total_tickets                    = n(),
    percent_vms_represented_tickets  = trunc(vms_represented_tickets / total_tickets * 10000) / 100,
    # repeat for landings
    vms_represented_landings         = sum(vms_represented_flag * DCRB_lbs),
    total_landings                   = sum(DCRB_lbs),
    percent_vms_represented_landings = trunc(vms_represented_landings / total_landings * 10000) / 100,
    # repeat for revenue
    vms_represented_revenue          = sum(vms_represented_flag * DCRB_revenue_afi),
    total_revenue                    = sum(DCRB_revenue_afi),
    percent_vms_represented_revenue  = trunc(vms_represented_revenue / total_revenue * 10000) / 100,
    # repeat for vessels
    vms_represented_vessels          = n_distinct(vms_represented_drvid, na.rm=TRUE),
    total_vessels                    = n_distinct(drvid),
    percent_vms_represented_vessels  = trunc(vms_represented_vessels / total_vessels * 10000) / 100,
    .groups = 'keep'
  ) %>%
  ungroup() %>%
  arrange(year, month_numeric, vessel_size_above_cutoff)

# write monthly result
write.csv(file = here("Confidential", "hindcast_output", "tables", "table_vms_representativeness_monthly_stratified.csv"),
          x = stratified_representativeness_monthly_table)

# repeat for yearly
stratified_representativeness_yearly_table <- dcrb_ticket_df %>%
  mutate(vessel_size_above_cutoff = FINAL_LENGTH > vessel_size_cutoff) %>%
  group_by(year, vessel_size_above_cutoff) %>%
  summarise(
    # check for tickets
    vms_represented_tickets          = sum(vms_represented_flag),
    total_tickets                    = n(),
    percent_vms_represented_tickets  = trunc(vms_represented_tickets / total_tickets * 10000) / 100,
    # repeat for landings
    vms_represented_landings         = sum(vms_represented_flag * DCRB_lbs),
    total_landings                   = sum(DCRB_lbs),
    percent_vms_represented_landings = trunc(vms_represented_landings / total_landings * 10000) / 100,
    # repeat for revenue
    vms_represented_revenue          = sum(vms_represented_flag * DCRB_revenue_afi),
    total_revenue                    = sum(DCRB_revenue_afi),
    percent_vms_represented_revenue  = trunc(vms_represented_revenue / total_revenue * 10000) / 100,
    # repeat for vessels
    vms_represented_vessels          = n_distinct(vms_represented_drvid, na.rm=TRUE),
    total_vessels                    = n_distinct(drvid),
    percent_vms_represented_vessels  = trunc(vms_represented_vessels / total_vessels * 10000) / 100,
    .groups = 'keep'
  ) %>%
  ungroup() %>%
  arrange(year, vessel_size_above_cutoff)

# write yearly result
write.csv(file = here("Confidential", "hindcast_output", "tables", "table_vms_representativeness_yearly_stratified.csv"),
          x = stratified_representativeness_yearly_table)
```

```{r size-stratified-representativeness-time-series}
# plot % represented by VMS for all four metrics stratified by vessel size category, yearly
stratified_vms_representativeness_time_series_yearly <- stratified_representativeness_yearly_table %>%
  select(year,
         vessel_size_above_cutoff,
         contains("percent_vms_represented")) %>%
  pivot_longer(cols = contains("percent_vms_represented"),
               names_to = "metric",
               names_prefix = "percent_vms_represented_",
               values_to ="percent_vms_represented") %>%
  ggplot(aes(x = as.factor(year), y = percent_vms_represented, fill = vessel_size_above_cutoff)) +
  geom_col(position = "dodge") +
  xlab("year") +
  facet_grid(metric ~ ., scales = "free_y")
ggsave(filename=here("Confidential", "hindcast_output", "figures", "time_series_representativeness_yearly_stratified.png"))
stratified_vms_representativeness_time_series_yearly

# plot % represented by VMS for all four metrics stratified by vessel size category, monthly
stratified_vms_representativeness_time_series_monthly <- stratified_representativeness_monthly_table %>%
  mutate(year_month = paste0(year,"_", ifelse(nchar(month_numeric)>1, month_numeric, paste0(0, month_numeric)))) %>%
  select(year_month,
         vessel_size_above_cutoff,
         contains("percent_vms_represented")) %>%
  pivot_longer(cols = contains("percent_vms_represented"),
               names_to = "metric",
               names_prefix = "percent_vms_represented_",
               values_to ="percent_vms_represented") %>%
  ggplot(aes(x = year_month, y = percent_vms_represented, fill = vessel_size_above_cutoff)) +
  geom_col(position = "dodge") +
  facet_grid(metric ~ ., scales = "free_y")
ggsave(filename=here("Confidential", "hindcast_output", "figures", "time_series_representativeness_monthly_stratified.png"))
stratified_vms_representativeness_time_series_monthly
```

```{r gridded-summary-tables}
# create monthly summary of gridded fishing activity
fishing_activity_monthly_table <- dcrb_5km_output_df %>%
  group_by(year, month_numeric) %>%
  summarise(
    # grid cell count
    n_grid_cells = n(),
    # VMS record count
    min_vms_records     = min(n_vms_records),
    median_vms_records  = median(n_vms_records),
    mean_vms_records    = mean(n_vms_records),
    max_vms_records     = max(n_vms_records),
    iqr_vms_records     = IQR(n_vms_records),
    # vessel count, based on vessels per VMS record (e.g. 1 trip with 3 pings would count as 1/3 vessel per VMS record)
    min_dcrb_vessels    = min(dcrb_vessels),
    median_dcrb_vessels = median(dcrb_vessels),
    mean_dcrb_vessels   = mean(dcrb_vessels),
    max_dcrb_vessels    = max(dcrb_vessels),
    iqr_dcrb_vessels    = IQR(dcrb_vessels),
    # unique vessel count
    min_unique_vessels    = min(n_unique_vessels),
    median_unique_vessels = median(n_unique_vessels),
    mean_unique_vessels   = mean(n_unique_vessels),
    max_unique_vessels    = max(n_unique_vessels),
    iqr_unique_vessels    = IQR(n_unique_vessels),
    .groups = 'keep'
  ) %>%
  arrange(year, month_numeric)
# write result
write.csv(file=here("Confidential", "hindcast_output", "tables", "table_fishing_activity_monthly.csv"),
          x=fishing_activity_monthly_table)

# repeat for yearly
fishing_activity_yearly_table <- dcrb_5km_output_df %>%
  group_by(year) %>%
  summarise(
    # grid cell month and grid cell count
    n_grid_cell_months = n(),
    n_grid_cells       = n_distinct(GRID5KM_ID),
    # VMS record count
    min_vms_records     = min(n_vms_records),
    median_vms_records  = median(n_vms_records),
    mean_vms_records    = mean(n_vms_records),
    max_vms_records     = max(n_vms_records),
    iqr_vms_records     = IQR(n_vms_records),
    # vessel count, based on vessels per VMS record (e.g. 1 trip with 3 pings would count as 1/3 vessel per VMS record)
    min_dcrb_vessels    = min(dcrb_vessels),
    median_dcrb_vessels = median(dcrb_vessels),
    mean_dcrb_vessels   = mean(dcrb_vessels),
    max_dcrb_vessels    = max(dcrb_vessels),
    iqr_dcrb_vessels    = IQR(dcrb_vessels),
    # unique vessel count
    min_unique_vessels    = min(n_unique_vessels),
    median_unique_vessels = median(n_unique_vessels),
    mean_unique_vessels   = mean(n_unique_vessels),
    max_unique_vessels    = max(n_unique_vessels),
    iqr_unique_vessels    = IQR(n_unique_vessels),
    .groups = 'keep'
  ) %>%
  arrange(year)
# write result
write.csv(file=here("Confidential", "hindcast_output", "tables", "table_fishing_activity_yearly.csv"),
          x=fishing_activity_yearly_table)
```

Create maps of fishing activity, based on the # of VMS pings, # of vessels (based on # vessels per VMS ping, e.g. 1 trip with 3 pings would count as 1/3 vessel per VMS record), and # of unique vessels in the gridded summary. This is one of the slowest parts of the script.

```{r fishing-activity-maps}
# map fishing activity, apply to each month of data
#   data_sf: sf object
#   data_col: vector from data_sf to log-transform and map
#   log_transform: boolean to log-transform data_col on map
#   fill_range: range for the fill of data_col vector
#   bbox: bounding box to use for map
#   title: map title
#   data_col_name: name of data_col vector, used to label map legend and name output PNG
#   png_suffix: optional string, used to name output PNG
#   background_sf: sf object
fishing_activity_map <- function (data_sf, data_col, bbox, log_transform, fill_range, title, data_col_name, png_suffix='', background_sf) {
  # log-transform data, if applicable
  if (log_transform) {
    data_col <- log10(data_col)
    fill_range <- log10(fill_range)
    data_col_name <- paste0("log10_", data_col_name)
  }
  # create map
  map_output <- ggplot() +
    # plot background
    geom_sf(data=background_sf, fill='gray80') +
    # plot log-10 transformed data
    geom_sf(data=data_sf, aes(fill=data_col)) +
    # label plot
    labs(title=title,
         fill=paste(data_col_name, "per 5x5 cell")) +
    # set bounding box
    xlim(bbox[1], bbox[3]) + ylim(bbox[2], bbox[4]) +
    # set fill color and legend limits
    scale_fill_viridis(limits=fill_range) +
    # adjust legend position
    theme(legend.position="bottom")
  # save the output map
  ggsave(filename=here("Confidential", "hindcast_output", "maps", paste0(data_col_name, "_", png_suffix, ".png")))
  # return the output map
  return(map_output)
}

# load west coast states simple features object for background
west_coast_states <- rnaturalearth::ne_states(country='United States of America', returnclass='sf') %>% 
  filter(name %in% c('California','Oregon','Washington','Nevada')) %>% 
  st_transform(st_crs(grid_5km))

# join geometry from 5km grid to gridded fishing summary
dcrb_5km_output_sf <- dcrb_5km_output_df %>%
  left_join(grid_5km, by=join_by(GRID5KM_ID)) %>%
  st_as_sf()

# set bounding box
bbox <- st_bbox(dcrb_5km_output_sf)

# set limits for min and max fill for # VMS pings and # unique vessels
vms_fill_range            <- range(dcrb_5km_output_sf$n_vms_records)
dcrb_vessels_fill_range   <- range(dcrb_5km_output_sf$dcrb_vessels)
unique_vessels_fill_range <- range(dcrb_5km_output_sf$n_unique_vessels)

# get unique and ordered year, year_month and month from gridded fishing summary
month_df <- dcrb_5km_output_sf %>% 
  st_drop_geometry() %>% 
  select(year, year_month, month) %>% 
  unique() %>% 
  arrange(year_month)

# iterate over months
for (ym in month_df$year_month) {
  # print progress
  print(ym)
  # get title as a combination of year and month
  y <- month_df$year[which(month_df$year_month==ym)]
  m <- month_df$month[which(month_df$year_month==ym)]
  title <- paste(y, m)
  # filter data for given month
  ym_data_sf <- dcrb_5km_output_sf %>% filter(year_month == ym)
  # make and save map of # VMS pings
  vms_map <- fishing_activity_map(data_sf = ym_data_sf,
                                  data_col = ym_data_sf$n_vms_records,
                                  bbox = bbox,
                                  log_transform = TRUE,
                                  fill_range = vms_fill_range,
                                  title = title,
                                  data_col_name = 'n_vms_records',
                                  png_suffix = ym,
                                  background_sf = west_coast_states)
  # make and save map of # vessels, based on vessels per VMS record
  # (e.g. 1 trip with 3 pings would count as 1/3 vessel per VMS record)
  dcrb_vessels_map <- fishing_activity_map(data_sf = ym_data_sf,
                                                  data_col = ym_data_sf$dcrb_vessels,
                                                  bbox = bbox,
                                                  log_transform = FALSE,
                                                  fill_range = dcrb_vessels_fill_range,
                                                  title = title,
                                                  data_col_name = 'dcrb_vessels',
                                                  png_suffix = ym,
                                                  background_sf = west_coast_states)
  # make and save map of # unique vessels
  unique_vessels_map <- fishing_activity_map(data_sf = ym_data_sf,
                                             data_col = ym_data_sf$n_unique_vessels,
                                             bbox = bbox,
                                             log_transform = FALSE,
                                             fill_range = unique_vessels_fill_range,
                                             title = title,
                                             data_col_name = 'n_unique_vessels',
                                             png_suffix = ym,
                                             background_sf = west_coast_states)
}

# view last one of each map
vms_map
dcrb_vessels_map
unique_vessels_map
```

```{r}
# clean up
remove(ym, y, m, ym_data_sf, month_df, bbox, title, vms_fill_range, dcrb_vessels_fill_range, vessels_fill_range)
```

Take static maps created above, and stitch together into an animated gif. Currently, the gif will skip months with no fishing data, rather than show a blank map for that month. It will stitch together all maps in the `Confidential/maps` output directory, even if they weren't generated with this particular run of the script. This is one of the slowest parts of the script.

```{r fishing-activity-gifs}
# create animation of fishing activity maps
#   map_directory: character, directory where map PNG files to animate are stored
#   map_metric_name: character, metric to map, used to search for PNG files in map_directory containing this metric name
fishing_activity_gif <- function (map_directory, map_metric_name) {
  # get list of files in the map directory
  file_list <- list.files(map_directory)
  # look for PNG files with metric name
  map_file_list <- file_list[grepl(map_metric_name, file_list) & grepl('.png', file_list)]
  # read in all images
  image_list <- lapply(paste0(map_directory, "/", map_file_list), image_read)
  # join and animate images
  image_gif <- image_animate(image_join(image_list), fps = 2)
  # write GIF
  image_write(image = image_gif, path = paste0(map_directory, "/animation_", map_metric_name, ".gif"))
  # return GIF
  return(image_gif)
}

# create animation for # VMS pings and # unique vessels
fishing_activity_gif(map_directory = here("Confidential", "hindcast_output", "maps"), map_metric_name = "n_vms_records")
fishing_activity_gif(map_directory = here("Confidential", "hindcast_output", "maps"), map_metric_name = "dcrb_vessels")
fishing_activity_gif(map_directory = here("Confidential", "hindcast_output", "maps"), map_metric_name = "n_unique_vessels")
```

```{r fishing-activity-time-series}
# TODO does this section need a confidentiality filter, since not spatial?
# TODO write these two time series sections as functions

# this uses the confidential data, using non-gridded summary to not double count unique vessels
time_series_df <- dcrb_confidential_df %>%
  group_by(crab_year, year, year_month, month) %>%
  summarise(n_vms_records = n(),
            dcrb_vessels = sum(dcrb_vessels_per_vms_record),
            n_unique_vessels = n_distinct(drvid),
            .groups = 'keep')

# plot # VMS pings per month as bar chart
n_vms_records_time_series <- time_series_df %>% 
  ggplot() + 
  geom_col(aes(x = year_month, y = n_vms_records, fill = crab_year))
ggsave(filename = here("Confidential", "hindcast_output", "figures", "time_series_n_vms_records.png"))
n_vms_records_time_series

# plot # vessels per VMS ping as bar chart
dcrb_vessels_time_series <- time_series_df %>% 
  ggplot() + 
  geom_col(aes(x = year_month, y = dcrb_vessels, fill = crab_year))
ggsave(filename = here("Confidential", "hindcast_output", "figures", "time_series_dcrb_vessels.png"))
dcrb_vessels_time_series

# plot # unique vessels per month as bar chart
n_unique_vessels_time_series <- time_series_df %>% 
  ggplot() + 
  geom_col(aes(x = year_month, y = n_unique_vessels, fill = crab_year))
ggsave(filename = here("Confidential", "hindcast_output", "figures", "time_series_n_unique_vessels.png"))
n_unique_vessels_time_series
```

```{r fishing-activity-heatmaps}
# plot # VMS pings per month as heatmap
n_vms_records_heatmap <- time_series_df %>%
  ggplot() +
  geom_tile(aes(x = month, y = as.factor(year), fill = n_vms_records))
ggsave(filename = here("Confidential", "hindcast_output", "figures", "heatmap_n_vms_records.png"))
n_vms_records_heatmap

# plot # vessels per VMS record as heatmap
dcrb_vessels_heatmap <- time_series_df %>%
  ggplot() +
  geom_tile(aes(x = month, y = as.factor(year), fill = dcrb_vessels))
ggsave(filename = here("Confidential", "hindcast_output", "figures", "heatmap_dcrb_vessels.png"))
dcrb_vessels_heatmap

# plot # unique vessels per month as heatmap
n_unique_vessels_heatmap <- time_series_df %>%
  ggplot() +
  geom_tile(aes(x = month, y = as.factor(year), fill = n_unique_vessels))
ggsave(filename = here("Confidential", "hindcast_output", "figures", "heatmap_n_unique_vessels.png"))
n_unique_vessels_heatmap
```
