---
title: "Overlay VMS points grid of choice and calcailate value of each 5km x 5km grid cell per year for dungeness crab"
output: 
  html_document:
    toc: true
    toc_float: true
---

### Set up

Attribute the VMS and fish ticket pipeline output for dungeness crab to each grid cell, then summarize for two baseline periods and one implementation period. Created for CDFW data request by Oct. 1, 2024.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(magrittr)
library(sf)
library(here)
library(knitr)
library(lubridate)
library(glue)
library(readr)
library(rnaturalearth) # for mapping states
library(viridis) # for good color scales
```

### Import data

Load VMS pings with fish tickets for all years.

```{r import data}
# clear workspace of everything but interpolation flag
rm(list = ls())

# load interpolated data
vms_interpolated <- read_rds(here('Confidential', 'processed', 'matched', 'interpolation', '2014_interpolated.rds')) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'matched', 'interpolation', '2015_interpolated.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'matched', 'interpolation', '2016_interpolated.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'matched', 'interpolation', '2017_interpolated.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'matched', 'interpolation', '2018_interpolated.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'matched', 'interpolation', '2019_interpolated.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'matched', 'interpolation', '2020_interpolated.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'matched', 'interpolation', '2021_interpolated.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'matched', 'interpolation', '2022_interpolated.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'matched', 'interpolation', '2023_interpolated.rds')))

# load non-interpolated data
vms_all <- read_rds(here('Confidential', 'processed', 'matched', 'filtering', '2014_matched_filtered_withFTID_length.rds')) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'matched', 'filtering', '2015_matched_filtered_withFTID_length.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'matched', 'filtering', '2016_matched_filtered_withFTID_length.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'matched', 'filtering', '2017_matched_filtered_withFTID_length.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'matched', 'filtering', '2018_matched_filtered_withFTID_length.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'matched', 'filtering', '2019_matched_filtered_withFTID_length.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'matched', 'filtering', '2020_matched_filtered_withFTID_length.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'matched', 'filtering', '2021_matched_filtered_withFTID_length.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'matched', 'filtering', '2022_matched_filtered_withFTID_length.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'matched', 'filtering', '2023_matched_filtered_withFTID_length.rds')))
```

``` {r basic counts before filter}
# count total records, trips and years in non-interpolated dataset
n_records_all <- nrow(vms_all)
n_trips_all <- n_distinct(vms_all$Rec_ID)
n_years_all <- n_distinct(lubridate::year(vms_all$westcoastdate_notime))

# count total records, trips and years in interpolated dataset
n_records_interpolated <- nrow(vms_interpolated)
n_trips_interpolated <- n_distinct(vms_interpolated$Rec_ID)
n_years_interpolated <- n_distinct(lubridate::year(vms_interpolated$westcoastdate_notime))

# count interpolated records and trips in interpolated dataset
n_interpolated_records_interpolated <- vms_interpolated %>% filter(is_interpolated == 1) %>% select(VMS_RECNO) %>% n_distinct()
n_interpolated_trips_interpolated <- vms_interpolated %>% filter(is_interpolated == 1) %>% select(Rec_ID) %>% n_distinct()
```

Before post-processing filtering, the non-interpolated data includes `r n_records_all` records (VMS pings), `r n_trips_all` trips (fish tickets), across `r n_years_all` years.
The interpolated data includes `r n_records_interpolated` records (VMS pings), `r n_trips_interpolated` trips (fish tickets), across `r n_years_interpolated` years.

`r n_interpolated_records_interpolated` records (`r round(n_interpolated_records_interpolated / n_records_interpolated * 100, 0)`%) are interpolated, and
`r n_interpolated_trips_interpolated` trips (`r round(n_interpolated_trips_interpolated / n_trips_interpolated * 100, 0)`%) have at least one interpolated record.

### Match VMS pings and fish tickets to grid cells

```{r}
# remove counts from above
rm(list = setdiff(ls(), c('vms_all', 'vms_interpolated')))

# load 5x5 grid shapefile
grd <- read_sf(here('GIS_layers', 'master_5km_grid_tmer.shp'))

# start timing
pt <- proc.time()

# convert VMS datasets to spatial object, then to planar projection to match the grid
vms_all_sf <- vms_all %>%
  st_as_sf(coords= c('LON','LAT'), crs=4326) %>% 
  st_transform(st_crs(grd))
vms_interpolated_sf <- vms_interpolated %>%
  st_as_sf(coords= c('LON','LAT'), crs=4326) %>% 
  st_transform(st_crs(grd))

# stop timing
x <- proc.time() - pt
```

Took `r round(x[3]/60,2)` minutes to do the conversion. Now for the join...

```{r join VMS with spatial object to grid}
# start timing
pt <- proc.time()

# do the join
vms_all_grd_match <- vms_all_sf %>%
  st_join(grd)
vms_interpolated_grd_match <- vms_interpolated_sf %>%
  st_join(grd)

# stop timing
x <- proc.time() - pt
```

The join took `r round(x[3]/60,2)` minutes.

```{r}
# save without geometry while still including grid cell ID
vms_all_grd_match %>% 
  st_set_geometry(NULL) %>% 
  write_rds(here('Confidential', 'processed', 'interim', 'vms_all_w_grd_5km.rds'))
# save without geometry while still including grid cell ID
vms_interpolated_grd_match %>% 
  st_set_geometry(NULL) %>% 
  write_rds(here('Confidential', 'processed', 'interim', 'vms_interpolated_w_grd_5km.rds'))

# load version without geometry with grid cell ID written above
vms_all_grd_match <- read_rds(here('Confidential', 'processed', 'interim', 'vms_all_w_grd_5km.rds'))
vms_interpolated_grd_match <- read_rds(here('Confidential', 'processed', 'interim', 'vms_interpolated_w_grd_5km.rds'))
```

### Filter data and add date columns

Now filter the data based on control variables below: dungeness crab target, landed in California, depth <= 150m, and vessel speeds <= 4 knots.
Then add date columns and select subset of columns.

Notes:

* *crab_year* is Nov.-Oct. (e.g. Nov. 2019 to Oct. 2020 for 2019_2020), rather than the precise dates of the crab fishing season.
* *agency_code* filters for VMS pings joined to fish tickets that were landed in California. You can see in the map at the end of the file this means some pings extend beyond California.
* *removal_type* to filter for commercial landings is not included here, so personal take is included in *DCRB_lbs*. This does not impact *DCRB_VMS_pings* or *DCRB_rev*.

```{r filter data}

# set filter parameters
state_agency_code = "C" # filter for landings brought to California ports
target_rev <- "DCRB" # I suspect this filter is redundant since Blake ran the pipeline for only DCRB, but I left it for now
target_lbs <- "DCRB" # again, I suspect this filter is redundant, but left it for now
winter_months <- c("November", "December", "January", "February", "March") # used to set season as Winter or Spring-Summer
min_depth <- 0
max_depth <- -150  # filter for 0-150 M depth
max_speed <- 4.11556 # units of m/s. 4.11556m/s = 8knots (4DCRB) and 1.54333 m/s = 3 knots (4CHNK)
min_speed <- 0 # units = m/s

# set fishing seasons to include
crab_years_to_vis = c("2014_2015", "2015_2016", "2016_2017", "2017_2018", "2018_2019",
                      "2019_2020", "2020_2021", "2021_2022", "2022_2023")

# filter data, add year/season columns, and select columns for all data
dcrb_vms_all <- vms_all_grd_match %>%
  filter(agency_code == state_agency_code) %>%
  filter(TARGET_rev == target_rev | TARGET_lbs == target_lbs) %>%
  filter(NGDC_M <= min_depth & NGDC_M >= max_depth) %>%
  filter(avg_speed_recalc <= max_speed & avg_speed_recalc >= min_speed) %>%
  mutate(
    year = lubridate::year(westcoastdate_notime),
    year_month = paste0(lubridate::year(westcoastdate_notime),"_", substr(lubridate::ymd(westcoastdate_notime),6,7)),
    # substr() ensures month is a 2 digit value, e.g. February is "02" not "2"
    month = lubridate::month(westcoastdate_notime, label=TRUE, abbr = FALSE),
    month_as_numeric = month(westcoastdate_notime),
    season = as.character(ifelse(month %in% winter_months, "Winter", "Spring-Summer")),
    crab_year = ifelse(
      month_as_numeric >= 11, paste0(year, "_", 1+year), paste0(year-1, "_", year)
    )
  ) %>%
  filter(crab_year %in% crab_years_to_vis) %>%
  dplyr::select(
      GRID5KM_ID, # grid cell ID
      Rec_ID, # fish ticket ID
      VMS_RECNO, # VMS ping ID
      drvid, # vessel ID
      FINAL_LENGTH,
      westcoastdate,
      westcoastdate_notime,
      year,
      crab_year, 
      year_month,
      month, 
      month_as_numeric,
      season,
      pacfin_port_code, 
      port_group_code, 
      NGDC_M, # depth
      TARGET_lbs, # target species by weight (filtered to "DCRB")
      TARGET_rev, # target species by revenue (filtered to "DCRB")
      DCRB_lbs, # dungeness crab catch weight
      DCRB_revenue # dungeness crab catch revenue
    )

# repeat for interpolated data
# only difference is to keep the is_interpolated column
dcrb_vms_interpolated <- vms_interpolated_grd_match %>%
  filter(agency_code == state_agency_code) %>%
  filter(TARGET_rev == target_rev | TARGET_lbs == target_lbs) %>%
  filter(NGDC_M <= min_depth & NGDC_M >= max_depth) %>%
  filter(avg_speed_recalc <= max_speed & avg_speed_recalc >= min_speed) %>%
  mutate(
    year = lubridate::year(westcoastdate_notime),
    year_month = paste0(lubridate::year(westcoastdate_notime),"_", substr(lubridate::ymd(westcoastdate_notime),6,7)),
    # substr() ensures month is a 2 digit value, e.g. February is "02" not "2"
    month = lubridate::month(westcoastdate_notime, label=TRUE, abbr = FALSE),
    month_as_numeric = month(westcoastdate_notime),
    season = as.character(ifelse(month %in% winter_months, "Winter", "Spring-Summer")),
    crab_year = ifelse(
      month_as_numeric >= 11, paste0(year, "_", 1+year), paste0(year-1, "_", year)
    )
  ) %>%
  filter(crab_year %in% crab_years_to_vis) %>%
  dplyr::select(
      GRID5KM_ID, # grid cell ID
      Rec_ID, # fish ticket ID
      VMS_RECNO, # VMS ping ID
      drvid, # vessel ID
      FINAL_LENGTH,
      westcoastdate,
      westcoastdate_notime,
      year,
      crab_year, 
      year_month,
      month, 
      month_as_numeric,
      season,
      pacfin_port_code, 
      port_group_code, 
      NGDC_M, # depth
      TARGET_lbs, # target species by weight (filtered to "DCRB")
      TARGET_rev, # target species by revenue (filtered to "DCRB")
      DCRB_lbs, # dungeness crab catch weight
      DCRB_revenue, # dungeness crab catch revenue
      is_interpolated
    )
```

``` {r basic counts after filter}
# count total records, trips and years in non-interpolated dataset
n_records_all <- nrow(dcrb_vms_all)
n_trips_all <- n_distinct(dcrb_vms_all$Rec_ID)
n_years_all <- n_distinct(lubridate::year(dcrb_vms_all$westcoastdate_notime))

# count total records, trips and years in interpolated dataset
n_records_interpolated <- nrow(dcrb_vms_interpolated)
n_trips_interpolated <- n_distinct(dcrb_vms_interpolated$Rec_ID)
n_years_interpolated <- n_distinct(lubridate::year(dcrb_vms_interpolated$westcoastdate_notime))

# count interpolated records and trips in interpolated dataset
n_interpolated_records_interpolated <- dcrb_vms_interpolated %>% filter(is_interpolated == 1) %>% select(VMS_RECNO) %>% n_distinct()
n_interpolated_trips_interpolated <- dcrb_vms_interpolated %>% filter(is_interpolated == 1) %>% select(Rec_ID) %>% n_distinct()
```

After post-processing filtering, the non-interpolated data includes `r n_records_all` records (VMS pings), `r n_trips_all` trips (fish tickets), across `r n_years_all` years.
The interpolated VMS data includes `r n_records_interpolated` records (VMS pings), `r n_trips_interpolated` trips (fish tickets), across `r n_years_interpolated` years.

`r n_interpolated_records_interpolated` records (`r round(n_interpolated_records_interpolated / n_records_interpolated * 100, 0)`%) are interpolated, and
`r n_interpolated_trips_interpolated` trips (`r round(n_interpolated_trips_interpolated / n_trips_interpolated * 100, 0)`%) have at least one interpolated record.

### Normalize pings and attribute landings and revenue to grid cells

If not using interpolated pings, also adjust pings based on # of pings / # hours per trip.

* For example, in a 38 hour trip with 38 pings, `avg_trip_ping_rate` = 38/38 = 1 ping/hour, and `VMSrecords_adjusted` = 1/1 = 1 adjusted ping.
* For example, in a 12 hour trip with 48 pings, `avg_trip_ping_rate` = 48/12 = 4 pings/hour, and `VMSrecords_adjusted` = 1/4 = 0.25 adjusted pings.

 Then, normalize interpolated or adjusted VMS pings by year-month and grid cell, and attribute landings and revenue across grid cells.
 
```{r distribute $ and lbs to cells, and adjust ping rate}

# remove counts from above and imported datasets, will use filtered versions for rest of the script
# keep filters, which will be used in VMS representation plots later on
rm(list = setdiff(ls(), c('dcrb_vms_all', 'dcrb_vms_interpolated', 'grd', 'state_agency_code', 'target_rev', 'target_lbs', 'winter_months', 'min_depth', 'max_depth', 'min_speed', 'max_speed', 'crab_years_to_vis')))

# create new columns that attribute pings, lbs, and $ from each fish ticket 
# across all VMS pings matched to that ticket

# total records per trip (total number of VMS records associated with each fish ticket)
# one record = one fish ticket
dcrb_trips_all <- dcrb_vms_all %>%
  group_by(Rec_ID) %>%
  summarise(trip_VMSrecords = n(),
            trip_start = min(westcoastdate),
            trip_end = max(westcoastdate),
            trip_duration_hours = if_else(as.numeric(trip_end - trip_start, units = 'hours') > 0,
                                          as.numeric(trip_end - trip_start, units = 'hours'),
                                          1),
            avg_trip_ping_rate = trip_VMSrecords / trip_duration_hours,
            .groups = 'drop')
# repeat for interpolated data
dcrb_trips_interpolated <- dcrb_vms_interpolated %>%
  group_by(Rec_ID) %>%
  summarise(trip_VMSrecords = n(),
            trip_start = min(westcoastdate),
            trip_end = max(westcoastdate),
            trip_duration_hours = if_else(as.numeric(trip_end - trip_start, units = 'hours') > 0,
                                          as.numeric(trip_end - trip_start, units = 'hours'),
                                          1),
            avg_trip_ping_rate = trip_VMSrecords / trip_duration_hours,
            .groups = 'drop')

# join total records per trip to VMS and fish ticket data
# adjust ping rate for non-interpolated data
# calculate vessels, lbs, and $ per VMS ping
dcrb_df_all <- left_join(dcrb_trips_all, dcrb_vms_all, by="Rec_ID") %>%
  mutate(
    VMSrecords_adjusted = 1 / avg_trip_ping_rate, # adjusted record-level pings (e.g. for 1 ping from trip with 4 pings per hour, this value is 0.25)
    DCRB_lbs_per_VMSlocation = DCRB_lbs / trip_VMSrecords,
    DCRB_rev_per_VMSlocation = DCRB_revenue / trip_VMSrecords,
    DCRB_Vessels_per_VMSlocation = 1 / trip_VMSrecords
  )
# repeat for interpolated data
dcrb_df_interpolated <- left_join(dcrb_trips_interpolated, dcrb_vms_interpolated, by="Rec_ID") %>%
  mutate(
    # no trip-wise adjustment for interpolated pings
    DCRB_lbs_per_VMSlocation = DCRB_lbs / trip_VMSrecords,
    DCRB_rev_per_VMSlocation = DCRB_revenue / trip_VMSrecords,
    DCRB_Vessels_per_VMSlocation = 1 / trip_VMSrecords
  )

# summarize and normalize fishery effort at year, month and grid cell level
# based on Jameal's prep_data_for_scenario_df_function.R
dcrb_year_month_5km_df_all <- dcrb_df_all %>%
  group_by(crab_year, year, year_month, month_as_numeric, month, GRID5KM_ID) %>%
  summarise(
    DCRB_lbs = sum(DCRB_lbs_per_VMSlocation),
    DCRB_rev = sum(DCRB_rev_per_VMSlocation),
    DCRB_VMS_pings = n(),
    DCRB_VMS_pings_adjusted = sum(VMSrecords_adjusted),
    DCRB_Vessels = sum(DCRB_Vessels_per_VMSlocation),
    Unique_DCRB_Vessels = length(unique(as.character(drvid)))
  ) %>%
  ungroup() %>%
  mutate(
    # rescale adjusted trip pings to 0-1
    normalized_adjusted_DCRB_VMS_pings = as.vector(scale(DCRB_VMS_pings_adjusted, center = min(DCRB_VMS_pings_adjusted), scale = diff(range(DCRB_VMS_pings_adjusted)))),
    # above rescaling converts minimum adjusted ping to 0, so change the minimum scaled value to one half of the rescaled value of the 2nd lowest adjusted ping value
    normalized_adjusted_DCRB_VMS_pings = ifelse(
      normalized_adjusted_DCRB_VMS_pings == 0, 
      0.5 * min(normalized_adjusted_DCRB_VMS_pings[normalized_adjusted_DCRB_VMS_pings != min(normalized_adjusted_DCRB_VMS_pings)]),
      normalized_adjusted_DCRB_VMS_pings
    )
  )
# repeat for interpolated data
dcrb_year_month_5km_df_interpolated <- dcrb_df_interpolated %>%
  group_by(crab_year, year, year_month, month_as_numeric, month, GRID5KM_ID) %>%
  summarise(
    DCRB_lbs = sum(DCRB_lbs_per_VMSlocation),
    DCRB_rev = sum(DCRB_rev_per_VMSlocation),
    DCRB_VMS_pings = n(),
    DCRB_Vessels = sum(DCRB_Vessels_per_VMSlocation),
    Unique_DCRB_Vessels = length(unique(as.character(drvid)))
  ) %>%
  ungroup() %>%
  mutate(
    # rescale adjusted trip pings to 0-1
    normalized_interpolated_DCRB_VMS_pings = as.vector(scale(DCRB_VMS_pings, center = min(DCRB_VMS_pings), scale = diff(range(DCRB_VMS_pings)))),
    # above rescaling converts minimum adjusted ping to 0, so change the minimum scaled value to one half of the rescaled value of the 2nd lowest adjusted ping value
    normalized_interpolated_DCRB_VMS_pings = ifelse(
      normalized_interpolated_DCRB_VMS_pings == 0, 
      0.5 * min(normalized_interpolated_DCRB_VMS_pings[normalized_interpolated_DCRB_VMS_pings != min(normalized_interpolated_DCRB_VMS_pings)]),
      normalized_interpolated_DCRB_VMS_pings
    )
  )

# join gridded summaries for trip-wise adjusted and interpolated pings
# dropping landings, revenue, and vessels; can add back in later if needed
dcrb_year_month_5km_df <- full_join(x = dcrb_year_month_5km_df_all, 
                                    y = dcrb_year_month_5km_df_interpolated,
                                    by = c('crab_year', 'year', 'year_month', 'month_as_numeric', 'month', 'GRID5KM_ID')) %>%
  rename(DCRB_VMS_pings_non_interpolated = DCRB_VMS_pings.x,
        DCRB_VMS_pings_interpolated = DCRB_VMS_pings.y) %>%
  select(crab_year, year, year_month, month_as_numeric, month, GRID5KM_ID,
         DCRB_VMS_pings_non_interpolated, # pings without interpolation, tripwise adjustment, or normalization
         DCRB_VMS_pings_adjusted, # pings with tripwise adjustment, without interpolation or normalization
         normalized_adjusted_DCRB_VMS_pings, # pings without interpolation, with tripwise adjustment and normalization
         DCRB_VMS_pings_interpolated, # pings without tripwise adjustment or normalization, with interpolation
         normalized_interpolated_DCRB_VMS_pings) # pings with interpolation and normalization, without tripwise adjustment

# save outputs before splitting into baseline vs. implementation periods
write_rds(dcrb_df_all, here::here('Confidential', 'processed', 'summaries', 'dcrb_df_all.rds'))
write_rds(dcrb_df_interpolated, here::here('Confidential', 'processed', 'summaries', 'dcrb_df_interpolated.rds'))
write_rds(dcrb_year_month_5km_df_all, here::here('Confidential', 'processed', 'summaries', 'dcrb_year_month_5km_df_all.rds'))
write_rds(dcrb_year_month_5km_df_interpolated, here::here('Confidential', 'processed', 'summaries', 'dcrb_year_month_5km_df_interpolated.rds'))
write_rds(dcrb_year_month_5km_df, here::here('Confidential', 'processed', 'summaries', 'dcrb_year_month_5km_df.rds'))
```

### Summarize fishing effort in baseline and implementation periods

* *Baseline 2014-2019* includes the start of Nov. 2014 to the end of Oct. 2019.
* *Baseline 2017-2019* includes the start of Nov. 2017 to the end of Oct. 2019.
* *Implementation 2019-2023* includes the start of Nov. 2019 to the end of Oct. 2023.

For baseline periods, each record represents one grid cell for one month, across all years. Fishing effort measured by # of pings is averaged across years for a given month.

For implementation period, each record represents one grid cell for one month in one year. Fishing effort measured by # of pings is not averaged across years for a given month.

```{r create summaries}
# set baseline and implementation years
baseline_2014_2019_crab_years <- c("2014_2015", "2015_2016", "2016_2017", "2017_2018", "2018_2019")
baseline_2017_2019_crab_years <- c("2017_2018", "2018_2019")
implementation_2019_2023_crab_years <- c("2019_2020", "2020_2021", "2021_2022", "2022_2023")

# summarize baseline data (2014-2018) at monthly temporal grain and grid cell ID spatial grain
baseline_2014_2019 <- dcrb_year_month_5km_df %>%
  filter(crab_year %in% baseline_2014_2019_crab_years) %>%
  group_by(GRID5KM_ID, month, month_as_numeric) %>%
  summarise(
    # take average by summing across years and dividing by # years
    # use sum / length instead of mean, because grid cells with 0 pings have no records, rather than a record with '0'
    avg_normalized_adjusted_DCRB_VMS_pings = sum(normalized_adjusted_DCRB_VMS_pings) / length(baseline_2014_2019_crab_years),
    avg_normalized_interpolated_DCRB_VMS_pings = sum(normalized_interpolated_DCRB_VMS_pings) / length(baseline_2014_2019_crab_years),
    .groups = 'drop'
  )
# write output
write_rds(baseline_2014_2019, here::here('Confidential', 'processed', 'summaries', 'baseline_2014_2019.rds'))

# summarize baseline data (2017-2018) at monthly temporal grain and grid cell ID spatial grain
baseline_2017_2019 <- dcrb_year_month_5km_df %>%
  filter(crab_year %in% baseline_2017_2019_crab_years) %>%
  group_by(GRID5KM_ID, month, month_as_numeric) %>%
  summarise(
    # take average by summing across years and dividing by # years
    # use sum / length instead of mean, because grid cells with 0 pings have no records, rather than a record with '0'
    avg_normalized_adjusted_DCRB_VMS_pings = sum(normalized_adjusted_DCRB_VMS_pings) / length(baseline_2017_2019_crab_years),
    avg_normalized_interpolated_DCRB_VMS_pings = sum(normalized_interpolated_DCRB_VMS_pings) / length(baseline_2017_2019_crab_years),
    .groups = 'drop'
  )
# write output
write_rds(baseline_2017_2019, here::here('Confidential', 'processed', 'summaries', 'baseline_2017_2019.rds'))

# write real-time data (2019-2023) at yearly and monthly temporal grain and grid cell ID spatial grain
implementation_2019_2023 <- dcrb_year_month_5km_df %>%
  filter(crab_year %in% implementation_2019_2023_crab_years)
# write output
write_rds(implementation_2019_2023, here::here('Confidential', 'processed', 'summaries', 'implementation_2019_2023.rds'))
```

### Visualize fishing effort

For a common sense check, see how the baseline periods and implementation period fishing effort look on a map. Visualize average # of pings for baseline periods, and # of pings for implementation period.

```{r map fishing effort}

# based on Owen's compare_hw_bh.R

# West coast, using the rnaturalearth package (for a background map)
coaststates <- ne_states(country='United States of America',returnclass = 'sf') %>% 
  filter(name %in% c('California','Oregon','Washington','Nevada')) %>% 
  # spatially transform to the same coordinate ref system as the 5k grid
  st_transform(st_crs(grd))
# join summary back to geometry by GRID5KM_ID
dcrb_year_month_5km_df_with_grid <- dcrb_year_month_5km_df %>% 
  # calculate percent difference between adjusted and interpolated pings
  mutate(DCRB_VMS_pings_diff = DCRB_VMS_pings_interpolated - DCRB_VMS_pings_adjusted,
         percent_DCRB_VMS_pings_diff = DCRB_VMS_pings_diff / DCRB_VMS_pings_adjusted * 100,
         normalized_DCRB_VMS_pings_diff = normalized_interpolated_DCRB_VMS_pings - normalized_adjusted_DCRB_VMS_pings,
         percent_normalized_DCRB_VMS_pings_diff = normalized_DCRB_VMS_pings_diff / normalized_adjusted_DCRB_VMS_pings * 100) %>%
  # pivot from wide to long to enable facet grids by transformation in charts later
  pivot_longer(cols = DCRB_VMS_pings_non_interpolated:percent_normalized_DCRB_VMS_pings_diff, names_to = "transformation", values_to = "pings") %>%
  # join grid
  left_join(grd, by = join_by(GRID5KM_ID)) %>%
  st_as_sf()
# set bounding box 
bbox <- st_bbox(dcrb_year_month_5km_df_with_grid)

# plot non-normalized values together
map_non_normalized_pings <- dcrb_year_month_5km_df_with_grid %>% 
  filter(transformation %in% c('DCRB_VMS_pings_adjusted', 'DCRB_VMS_pings_interpolated')) %>%
  ggplot() +
  geom_sf(data = coaststates, fill = 'gray80') +
  geom_sf(aes(fill = log10(pings)), color = NA) +
  facet_grid(transformation ~ month) +
  scale_fill_viridis() +
  xlim(bbox[1], bbox[3]) + ylim(bbox[2], bbox[4]) +
  labs(fill="log10 pings variable per 5x5 cell") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="bottom")
map_non_normalized_pings
ggsave(filename = here::here("Confidential", "figures", "map_non_normalized_pings.png"),
       width = 10, height = 5)

# plot normalized values together
map_normalized_pings <- dcrb_year_month_5km_df_with_grid %>% 
  filter(transformation %in% c('normalized_adjusted_DCRB_VMS_pings', 'normalized_interpolated_DCRB_VMS_pings')) %>%
  ggplot() +
  geom_sf(data = coaststates, fill = 'gray80') +
  geom_sf(aes(fill = log10(pings)), color = NA) +
  facet_grid(transformation ~ month) +
  scale_fill_viridis() +
  xlim(bbox[1], bbox[3]) + ylim(bbox[2], bbox[4]) +
  labs(fill="log10 pings variable per 5x5 cell") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="bottom")
map_normalized_pings
ggsave(filename = here::here("Confidential", "figures", "map_normalized_pings.png"),
       width = 10, height = 5)

# plot normalized values separately
# for adjusted pings
map_adjusted_normalized_pings <- dcrb_year_month_5km_df_with_grid %>% 
  filter(transformation %in% c('normalized_adjusted_DCRB_VMS_pings')) %>%
  ggplot() +
  geom_sf(data = coaststates, fill = 'gray80') +
  geom_sf(aes(fill = log10(pings)), color = NA) +
  facet_grid(. ~ month) +
  scale_fill_viridis() +
  xlim(bbox[1], bbox[3]) + ylim(bbox[2], bbox[4]) +
  labs(fill="log10 normalized_adjusted_DCRB_VMS_pings per 5x5 cell") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="bottom")
map_adjusted_normalized_pings
ggsave(filename = here::here("Confidential", "figures", "map_normalized_adjusted_DCRB_VMS_pings.png"))
# for interpolated pings
map_interpolated_normalized_pings <- dcrb_year_month_5km_df_with_grid %>% 
  filter(transformation %in% c('normalized_interpolated_DCRB_VMS_pings')) %>%
  ggplot() +
  geom_sf(data = coaststates, fill = 'gray80') +
  geom_sf(aes(fill = log10(pings)), color = NA) +
  facet_grid(. ~ month) +
  scale_fill_viridis() +
  xlim(bbox[1], bbox[3]) + ylim(bbox[2], bbox[4]) +
  labs(fill="log10 normalized_interpolated_DCRB_VMS_pings per 5x5 cell") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="bottom")
map_interpolated_normalized_pings
ggsave(filename = here::here("Confidential", "figures", "map_normalized_interpolated_DCRB_VMS_pings.png"))

# plot absolute difference between interpolated and adjusted pings (before and after normalization)
dcrb_year_month_5km_df_with_grid %>% 
  filter(transformation %in% c('DCRB_VMS_pings_diff', 'normalized_DCRB_VMS_pings_diff')) %>%
  ggplot() +
  geom_sf(data = coaststates, fill = 'gray80') +
  geom_sf(aes(fill = log10(pings)), color = NA) +
  facet_grid(transformation ~ month) +
  scale_fill_viridis() +
  xlim(bbox[1], bbox[3]) + ylim(bbox[2], bbox[4]) +
  labs(fill="log10 absolute difference for pings (before and after normalization) per 5x5 cell") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="bottom")
ggsave(filename = here::here("Confidential", "figures", "map_diff_DCRB_VMS_pings.png"))

# plot % difference between interpolated and adjusted pings (before and after normalization)
dcrb_year_month_5km_df_with_grid %>% 
  filter(transformation %in% c('percent_DCRB_VMS_pings_diff', 'percent_normalized_DCRB_VMS_pings_diff')) %>%
  ggplot() +
  geom_sf(data = coaststates, fill = 'gray80') +
  geom_sf(aes(fill = log10(pings)), color = NA) +
  facet_grid(transformation ~ month) +
  scale_fill_viridis() +
  xlim(bbox[1], bbox[3]) + ylim(bbox[2], bbox[4]) +
  labs(fill="log10 % difference for pings (before and after normalization) per 5x5 cell") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="bottom")
ggsave(filename = here::here("Confidential", "figures", "map_percent_diff_DCRB_VMS_pings.png"))

# plot % difference between normalized interpolated and adjusted pings by month
dcrb_year_month_5km_df_with_grid %>% 
  filter(transformation %in% c('percent_normalized_DCRB_VMS_pings_diff')) %>%
  ggplot() +
  geom_sf(data = coaststates, fill = 'gray80') +
  geom_sf(aes(fill = log10(pings)), color = NA) +
  facet_grid(. ~ month) +
  scale_fill_viridis() +
  xlim(bbox[1], bbox[3]) + ylim(bbox[2], bbox[4]) +
  labs(fill="log10 % difference for normalized pings per 5x5 cell") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="bottom")
ggsave(filename = here::here("Confidential", "figures", "map_percent_diff_DCRB_VMS_pings_month.png"))

# plot % difference between normalized interpolated and adjusted pings by crab year
dcrb_year_month_5km_df_with_grid %>% 
  filter(transformation %in% c('percent_normalized_DCRB_VMS_pings_diff')) %>%
  ggplot() +
  geom_sf(data = coaststates, fill = 'gray80') +
  geom_sf(aes(fill = log10(pings)), color = NA) +
  facet_grid(. ~ crab_year) +
  scale_fill_viridis() +
  xlim(bbox[1], bbox[3]) + ylim(bbox[2], bbox[4]) +
  labs(fill="log10 % difference for normalized pings per 5x5 cell") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="bottom")
ggsave(filename = here::here("Confidential", "figures", "map_percent_diff_DCRB_VMS_pings_crab_year.png"))

# quick histogram of % difference in fishing effort per grid cell (each point = 1 grid cell year month)
dcrb_year_month_5km_df_with_grid %>% 
  filter(transformation %in% c('percent_normalized_DCRB_VMS_pings_diff')) %>%
  ggplot(aes(x = pings)) + 
  labs(x="% difference for normalized pings per 5x5 cell for a year and month") +
  geom_histogram(binwidth = 5)+ 
  xlim(c(-200,500)) # note limited axes from -2x (-200%) to +5x (+500%)

# boxplot of fishing effort per grid cell (each point = 1 gird cell year month)
dcrb_year_month_5km_df_with_grid %>%
  filter(transformation %in% c('DCRB_VMS_pings_adjusted', 'DCRB_VMS_pings_interpolated')) %>%
  ggplot(aes(fill = transformation, y = pings)) +
  geom_boxplot() + 
  facet_grid(. ~ crab_year) + 
  ylim(c(0,25)) # note limited axes

# quick summary of each transformation
summary(dcrb_year_month_5km_df$DCRB_VMS_pings_non_interpolated)
summary(dcrb_year_month_5km_df$DCRB_VMS_pings_adjusted)
summary(dcrb_year_month_5km_df$normalized_adjusted_DCRB_VMS_pings)
summary(dcrb_year_month_5km_df$DCRB_VMS_pings_interpolated)
summary(dcrb_year_month_5km_df$normalized_interpolated_DCRB_VMS_pings)

# scatterplot of adjusted vs. interpolated pings
dcrb_year_month_5km_df %>%
  ggplot(aes(x = DCRB_VMS_pings_adjusted, y = DCRB_VMS_pings_interpolated)) + 
  geom_point() +
  geom_abline(slope=1, intercept=0) + 
  geom_rug()
ggsave(filename = here::here("Confidential", "figures", "scatterplot_adjusted_interpolated_pings.png"))

# scatterplot of normalized adjusted vs. normalized interpolated pings
dcrb_year_month_5km_df %>%
  ggplot(aes(x = normalized_adjusted_DCRB_VMS_pings, y = normalized_interpolated_DCRB_VMS_pings)) + 
  geom_point() +
  geom_abline(slope=1, intercept=0) + 
  geom_rug()

# correlation between adjusted and interpolated pings
cor(x = dcrb_year_month_5km_df$DCRB_VMS_pings_adjusted,
    y = dcrb_year_month_5km_df$DCRB_VMS_pings_interpolated,
    use = "pairwise.complete.obs")

# correlation between normalized adjusted and normalized interpolated pings
cor(x = dcrb_year_month_5km_df$normalized_adjusted_DCRB_VMS_pings,
    y = dcrb_year_month_5km_df$normalized_interpolated_DCRB_VMS_pings,
    use = "pairwise.complete.obs")
```

### Visualize basic time series

```{r plot monthly basic counts}
# summarize counts of vessels, tickets, days with VMS pings, and # pings
monthly_df_all <- dcrb_df_all %>% 
  mutate(month_as_date = round_date(westcoastdate_notime, unit='month')) %>%
  group_by(year, crab_year, month_as_date) %>%
  summarise(n_vessels = n_distinct(drvid),
            n_tickets = n_distinct(Rec_ID),
            n_days = n_distinct(westcoastdate_notime),
            n_pings = n())
# visualize count represented vs. not represented for vessels, landings, and revenue
monthly_df_all %>% 
  pivot_longer(cols = n_vessels:n_pings, names_to = "variable", values_to = "value") %>%
  ggplot(aes(x = month_as_date, y = value, color = crab_year, fill = crab_year)) +
  geom_col(position = "dodge") + 
  facet_grid(variable ~ ., scales = "free_y") +
  expand_limits(y = 0)
ggsave(filename = here::here("Confidential", "figures", "counts_non_interpolated.png"))

# repeat for interpolated data
# summarize counts of vessels, tickets, days with VMS pings, and # pings
monthly_df_interpolated <- dcrb_df_interpolated %>% 
  mutate(month_as_date = round_date(westcoastdate_notime, unit='month')) %>%
  group_by(year, crab_year, month_as_date) %>%
  summarise(n_vessels = n_distinct(drvid),
            n_tickets = n_distinct(Rec_ID),
            n_days = n_distinct(westcoastdate_notime),
            n_pings = n())
# visualize count represented vs. not represented for vessels, landings, and revenue
monthly_df_interpolated %>% 
  pivot_longer(cols = n_vessels:n_pings, names_to = "variable", values_to = "value") %>%
  ggplot(aes(x = month_as_date, y = value, color = crab_year, fill = crab_year)) +
  geom_col(position = "dodge") + 
  facet_grid(variable ~ ., scales = "free_y") +
  expand_limits(y = 0)
ggsave(filename = here::here("Confidential", "figures", "counts_interpolated.png"))
```

### Visualize ping interval and rate

```{r plot ping interval}

# visualize for non-interpolated data
# each point = 1 VMS ping
ping_interval_df_all <- dcrb_df_all %>%
  group_by(Rec_ID) %>%
  arrange(westcoastdate) %>%
  mutate(
    next_time = lag(westcoastdate, 1),
    ping_interval_min = as.numeric(difftime(westcoastdate, next_time, units = 'mins')),
    ping_interval_hours = as.numeric(difftime(westcoastdate, next_time, units = 'hours'))
  )
# summarize by year
ping_interval_df_all %>%
  group_by(crab_year) %>%
  summarise(min = min(ping_interval_hours, na.rm=TRUE),
            median = median(ping_interval_hours, na.rm=TRUE),
            mean = mean(ping_interval_hours, na.rm=TRUE),
            max = max(ping_interval_hours, na.rm=TRUE))
# visualize by year
ping_interval_df_all %>%
  ggplot(aes(x = as.factor(crab_year), y = ping_interval_hours)) + geom_boxplot() + ylim(c(0,2)) # note limited axes
# visualize by month in 2019-2021 (expect shift in Sept. 2020 based on what Blake saw in a quick check in the raw VMS data)
ping_interval_df_all %>%
  filter(crab_year %in% c("2019_2020", "2020_2021")) %>%
  ggplot(aes(x = as.factor(year_month), y = ping_interval_hours)) + geom_boxplot() + ylim(c(0,2)) + # note limited axes
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
# visualize by day in Dec. 2020-Jan. 2021
ping_interval_df_all %>%
  filter(year_month %in% c("2020_12", "2021_01")) %>%
  ggplot(aes(x = as.factor(westcoastdate_notime), y = ping_interval_hours)) + geom_boxplot() + ylim(c(0,2)) + # note limited axes
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
# I'm seeing the shift in ping rates between Dec. 30, 2020 and Jan. 7, 2021

# repeat for interpolated data
# each point = 1 VMS ping
ping_interval_df_interpolated <- dcrb_df_interpolated %>%
  group_by(Rec_ID) %>%
  arrange(westcoastdate) %>%
  mutate(
    next_time = lag(westcoastdate, 1),
    ping_interval_min = as.numeric(difftime(westcoastdate, next_time, units = 'mins')),
    ping_interval_hours = as.numeric(difftime(westcoastdate, next_time, units = 'hours'))
  )
# summarize by year
ping_interval_df_interpolated %>%
  group_by(crab_year) %>%
  summarise(min = min(ping_interval_hours, na.rm=TRUE),
            median = median(ping_interval_hours, na.rm=TRUE),
            mean = mean(ping_interval_hours, na.rm=TRUE),
            max = max(ping_interval_hours, na.rm=TRUE))
# visualize by year
ping_interval_df_interpolated %>%
  ggplot(aes(x = as.factor(crab_year), y = ping_interval_hours)) + geom_boxplot() + ylim(c(0,2)) # note limited axes
# visualize by month in 2019-2021 (expect shift in Sept. 2020 based on what Blake saw in a quick check in the raw VMS data)
ping_interval_df_interpolated %>%
  filter(crab_year %in% c("2019_2020", "2020_2021")) %>%
  ggplot(aes(x = as.factor(year_month), y = ping_interval_hours)) + geom_boxplot() + ylim(c(0,2)) + # note limited axes
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
# visualize by day in Dec. 2020-Jan. 2021
ping_interval_df_interpolated %>%
  filter(year_month %in% c("2020_12", "2021_01")) %>%
  ggplot(aes(x = as.factor(westcoastdate_notime), y = ping_interval_hours)) + geom_boxplot() + ylim(c(0,2)) + # note limited axes
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
# I'm seeing some outliers that aren't perfectly one hour intervals
```

```{r plot ping rate}
# start with non-interpolated data
# avg. ping rate summary
dcrb_trips_all %>%
  group_by(year(trip_start)) %>%
  summarise(min = min(avg_trip_ping_rate, na.rm=TRUE),
            median = median(avg_trip_ping_rate, na.rm=TRUE),
            mean = mean(avg_trip_ping_rate, na.rm=TRUE),
            max = max(avg_trip_ping_rate, na.rm=TRUE))
# avg. ping rate  boxplot
dcrb_trips_all %>%
  ggplot(aes(x=as.factor(year(trip_start)), y=avg_trip_ping_rate)) + geom_boxplot() + ylim(c(0,10)) # note limited axes

# repeat with interpolated data
dcrb_trips_interpolated %>%
  group_by(year(trip_start)) %>%
  summarise(min = min(avg_trip_ping_rate, na.rm=TRUE),
            median = median(avg_trip_ping_rate, na.rm=TRUE),
            mean = mean(avg_trip_ping_rate, na.rm=TRUE),
            max = max(avg_trip_ping_rate, na.rm=TRUE))
# avg. ping rate  boxplot
dcrb_trips_interpolated %>%
  ggplot(aes(x=as.factor(year(trip_start)), y=avg_trip_ping_rate)) + geom_boxplot()
```

### Visualize VMS representativeness

```{r plot vessel, revenue, and landings representation by VMS data}

# load all fish tickets
fish_ticket_df <- read_rds(here('Confidential', 'processed', 'fish tickets', '2014fishtix_vlengths_withFTID.rds')) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'fish tickets', '2015fishtix_vlengths_withFTID.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'fish tickets', '2016fishtix_vlengths_withFTID.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'fish tickets', '2017fishtix_vlengths_withFTID.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'fish tickets', '2018fishtix_vlengths_withFTID.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'fish tickets', '2019fishtix_vlengths_withFTID.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'fish tickets', '2020fishtix_vlengths_withFTID.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'fish tickets', '2021fishtix_vlengths_withFTID.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'fish tickets', '2022fishtix_vlengths_withFTID.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'fish tickets', '2023fishtix_vlengths_withFTID.rds')))

# apply same filters as dcrb_vms_tix_analysis, except for depth and speed which are from VMS
dcrb_ticket_df <- fish_ticket_df %>%
  filter(agency_code == state_agency_code) %>%
  filter(TARGET_rev == target_rev | TARGET_lbs == target_lbs) %>%
  mutate(
    year_month = paste0(lubridate::year(date),"_", substr(lubridate::ymd(date),6,7)),
    month_as_numeric = month(date),
    month_as_date = round_date(date, unit='month'),
    season = as.character(ifelse(month %in% winter_months, "Winter", "Spring-Summer")),
    crab_year = ifelse(
      month_as_numeric >= 11, paste0(year, "_", 1+year), paste0(year-1, "_", year)
    )
  ) %>%
  filter(crab_year %in% crab_years_to_vis) %>%
  dplyr::select(
      Rec_ID, # fish ticket ID
      drvid, # vessel ID
      FINAL_LENGTH,
      date,
      year,
      crab_year,
      year_month,
      month,
      month_as_numeric,
      month_as_date,
      season,
      pacfin_port_code,
      port_group_code,
      TARGET_lbs, # target species by weight (filtered to "DCRB")
      TARGET_rev, # target species by revenue (filtered to "DCRB") # TODO - check do I need to adjust revenue?
      DCRB_lbs, # dungeness crab catch weight
      DCRB_revenue # dungeness crab catch revenue
    )

# start with non-interpolated data

# select distinct ticket IDs from filtered VMS dataframe
dcrb_vms_rec_ids_all <- dcrb_df_all %>%
  dplyr::select(Rec_ID) %>%
  distinct()

# add column to ticket dataframe for whether that ticket was represented in filtered VMS dataframe
dcrb_vms_representation_df_all <- left_join(dcrb_ticket_df, dcrb_vms_rec_ids_all, by = "Rec_ID", keep = TRUE) %>%
   # if ticket ID is NA in dcrb_vms_rec_ids, then the ticket is not represented by VMS data
  mutate(represented_in_vms = !is.na(Rec_ID.y)) %>%
  dplyr::select(-Rec_ID.y) %>%
  rename(Rec_ID = Rec_ID.x) %>%
  # calculate monthly counts
  group_by(month_as_date) %>% 
  mutate(total_unique_vessels = n_distinct(drvid),
         total_landings = sum(DCRB_lbs),
         total_revenue = sum(DCRB_revenue)) %>%
  group_by(month_as_date, represented_in_vms, total_unique_vessels, total_landings, total_revenue) %>%
   # calculate counts represented by VMS
  summarise(unique_vessels = n_distinct(drvid), # if a vessel has a ticket with and without VMS data, it will be double counted
            landings = sum(DCRB_lbs),
            revenue = sum(DCRB_revenue)) %>%
  # calculate percentages represented by VMS
  mutate(percent_unique_vessels = unique_vessels / total_unique_vessels * 100,
         percent_landings = landings / total_landings * 100,
         percent_revenue = revenue / total_revenue * 100)

# visualize count represented vs. not represented for vessels, landings, and revenue
dcrb_vms_representation_df_all %>% 
  pivot_longer(cols = unique_vessels:revenue, names_to = "variable", values_to = "value") %>%
  ggplot(aes(x = month_as_date, y = value, color = represented_in_vms, fill = represented_in_vms)) +
  geom_col(position = "dodge") + 
  facet_grid(variable ~ ., scales = "free_y") +
  expand_limits(y = 0)

# visualize % represented for vessels, landing, and revenue
dcrb_vms_representation_df_all %>%
  filter(represented_in_vms == TRUE) %>%
  dplyr::select(percent_unique_vessels:percent_revenue) %>%
  pivot_longer(cols = percent_unique_vessels:percent_revenue, names_to = "variable", values_to = "value") %>%
  ggplot(aes(x = month_as_date, y = value)) +
  geom_col() + 
  facet_grid(variable ~ ., scales = "free_y") +
  expand_limits(y = 0)

# calculate VMS representation for vessels, landing, and revenue across entire timeframe (not monthly)
# calculate totals across timeframe
timeframe_totals_all <- left_join(dcrb_ticket_df, dcrb_vms_rec_ids_all, by = "Rec_ID", keep = TRUE) %>%
  mutate(represented_in_vms = !is.na(Rec_ID.y)) %>% # if ticket ID is NA in dcrb_vms_rec_ids, then the landings are not represented by VMS data
  dplyr::select(-Rec_ID.y) %>%
  rename(Rec_ID = Rec_ID.x) %>% 
  summarise(total_timeframe_vessels = n_distinct(drvid),
            total_timeframe_landings = sum(DCRB_lbs),
            total_timeframe_revenue = sum(DCRB_revenue))
timeframe_totals_all
# calculate totals for VMS representation across timeframe
timeframe_vms_represented_totals_all <- left_join(dcrb_ticket_df, dcrb_vms_rec_ids_all, by = "Rec_ID", keep = TRUE) %>%
  mutate(represented_in_vms = !is.na(Rec_ID.y)) %>% # if ticket ID is NA in dcrb_vms_rec_ids, then the landings are not represented by VMS data
  dplyr::select(-Rec_ID.y) %>%
  rename(Rec_ID = Rec_ID.x) %>% 
  filter(represented_in_vms == TRUE) %>%
  summarise(total_timeframe_vms_vessels = n_distinct(drvid),
            total_timeframe_vms_landings = sum(DCRB_lbs),
            total_timeframe_vms_revenue = sum(DCRB_revenue))
timeframe_vms_represented_totals_all
# divide totals with VMS representation by totals to get percentage of representation
timeframe_vms_percent_representation_all <- timeframe_vms_represented_totals_all / timeframe_totals_all * 100
timeframe_vms_percent_representation_all

# create representation dataframe grouped by vessel
dcrb_vms_representation_vessel_size_df_all <- left_join(dcrb_ticket_df, dcrb_vms_rec_ids_all, by = "Rec_ID", keep = TRUE) %>%
   # if ticket ID is NA in dcrb_vms_rec_ids, then the ticket is not represented by VMS data
  mutate(represented_in_vms = !is.na(Rec_ID.y)) %>%
  dplyr::select(-Rec_ID.y) %>%
  rename(Rec_ID = Rec_ID.x) %>%
  # select vessel and ticket related columns
  dplyr::select(drvid, FINAL_LENGTH, Rec_ID, represented_in_vms, DCRB_lbs, DCRB_revenue) %>%
  distinct() %>% # now, each row represents one ticket
  # calculate vessel summaries
  group_by(drvid) %>% # now, each row represents one vessel
  summarise(
    vessel_length = max(FINAL_LENGTH), # this just selects one length, since lengths are consistent
    tickets_represented_in_vms = sum(represented_in_vms),
    tickets_not_represented_in_vms = sum(!represented_in_vms),
    total_tickets = n_distinct(Rec_ID),
    percent_tickets_represented_in_vms = tickets_represented_in_vms / total_tickets * 100,
    represented_in_vms_max = max(represented_in_vms),
    represented_in_vms_majority = percent_tickets_represented_in_vms > 50, # true if over 50% of tickets have VMS data
    total_landings = sum(DCRB_lbs),
    total_revenue = sum(DCRB_revenue)
  )

# visualize vessels with VMS by vessel size
# if any ticket has VMS data, consider the vessel as represented in VMS data
dcrb_vms_representation_vessel_size_df_all %>%
  ggplot(aes(x = vessel_length, color = as.factor(represented_in_vms_max), fill = as.factor(represented_in_vms_max))) +
  geom_histogram(position='dodge') +
  theme(legend.position = "bottom")

# visualize vessels with VMS by vessel size
# if >50% of vessel tickets have VMS data, consider the vessel as represented in VMS data
dcrb_vms_representation_vessel_size_df_all %>%
  ggplot(aes(x = vessel_length, color = as.factor(represented_in_vms_majority), fill = as.factor(represented_in_vms_majority))) +
  geom_histogram(position='dodge') +
  theme(legend.position = "bottom")

# repeat for interpolated data (this section could use cleaning up)

# select distinct ticket IDs from filtered VMS dataframe
dcrb_vms_rec_ids_interpolated <- dcrb_df_interpolated %>%
  dplyr::select(Rec_ID) %>%
  distinct()

# add column to ticket dataframe for whether that ticket was represented in filtered VMS dataframe
dcrb_vms_representation_df_interpolated <- left_join(dcrb_ticket_df, dcrb_vms_rec_ids_interpolated, by = "Rec_ID", keep = TRUE) %>%
   # if ticket ID is NA in dcrb_vms_rec_ids, then the ticket is not represented by VMS data
  mutate(represented_in_vms = !is.na(Rec_ID.y)) %>%
  dplyr::select(-Rec_ID.y) %>%
  rename(Rec_ID = Rec_ID.x) %>%
  # calculate monthly counts
  group_by(month_as_date) %>% 
  mutate(total_unique_vessels = n_distinct(drvid),
         total_landings = sum(DCRB_lbs),
         total_revenue = sum(DCRB_revenue)) %>%
  group_by(month_as_date, represented_in_vms, total_unique_vessels, total_landings, total_revenue) %>%
   # calculate counts represented by VMS
  summarise(unique_vessels = n_distinct(drvid), # if a vessel has a ticket with and without VMS data, it will be double counted
            landings = sum(DCRB_lbs),
            revenue = sum(DCRB_revenue)) %>%
  # calculate percentages represented by VMS
  mutate(percent_unique_vessels = unique_vessels / total_unique_vessels * 100,
         percent_landings = landings / total_landings * 100,
         percent_revenue = revenue / total_revenue * 100)

# visualize count represented vs. not represented for vessels, landings, and revenue
dcrb_vms_representation_df_interpolated %>% 
  pivot_longer(cols = unique_vessels:revenue, names_to = "variable", values_to = "value") %>%
  ggplot(aes(x = month_as_date, y = value, color = represented_in_vms, fill = represented_in_vms)) +
  geom_col(position = "dodge") + 
  facet_grid(variable ~ ., scales = "free_y") +
  expand_limits(y = 0)

# visualize % represented for vessels, landing, and revenue
dcrb_vms_representation_df_interpolated %>%
  filter(represented_in_vms == TRUE) %>%
  dplyr::select(percent_unique_vessels:percent_revenue) %>%
  pivot_longer(cols = percent_unique_vessels:percent_revenue, names_to = "variable", values_to = "value") %>%
  ggplot(aes(x = month_as_date, y = value)) +
  geom_col() + 
  facet_grid(variable ~ ., scales = "free_y") +
  expand_limits(y = 0)

# calculate VMS representation for vessels, landing, and revenue across entire timeframe (not monthly)
# calculate totals across timeframe
timeframe_totals_interpolated <- left_join(dcrb_ticket_df, dcrb_vms_rec_ids_interpolated, by = "Rec_ID", keep = TRUE) %>%
  mutate(represented_in_vms = !is.na(Rec_ID.y)) %>% # if ticket ID is NA in dcrb_vms_rec_ids, then the landings are not represented by VMS data
  dplyr::select(-Rec_ID.y) %>%
  rename(Rec_ID = Rec_ID.x) %>% 
  summarise(total_timeframe_vessels = n_distinct(drvid),
            total_timeframe_landings = sum(DCRB_lbs),
            total_timeframe_revenue = sum(DCRB_revenue))
timeframe_totals_interpolated
# calculate totals for VMS representation across timeframe
timeframe_vms_represented_totals_interpolated <- left_join(dcrb_ticket_df, dcrb_vms_rec_ids_interpolated, by = "Rec_ID", keep = TRUE) %>%
  mutate(represented_in_vms = !is.na(Rec_ID.y)) %>% # if ticket ID is NA in dcrb_vms_rec_ids, then the landings are not represented by VMS data
  dplyr::select(-Rec_ID.y) %>%
  rename(Rec_ID = Rec_ID.x) %>% 
  filter(represented_in_vms == TRUE) %>%
  summarise(total_timeframe_vms_vessels = n_distinct(drvid),
            total_timeframe_vms_landings = sum(DCRB_lbs),
            total_timeframe_vms_revenue = sum(DCRB_revenue))
timeframe_vms_represented_totals_interpolated
# divide totals with VMS representation by totals to get percentage of representation
timeframe_vms_percent_representation_interpolated <- timeframe_vms_represented_totals_interpolated / timeframe_totals_interpolated * 100
timeframe_vms_percent_representation_interpolated

# create representation dataframe grouped by vessel
dcrb_vms_representation_vessel_size_df_interpolated <- left_join(dcrb_ticket_df, dcrb_vms_rec_ids_interpolated, by = "Rec_ID", keep = TRUE) %>%
   # if ticket ID is NA in dcrb_vms_rec_ids, then the ticket is not represented by VMS data
  mutate(represented_in_vms = !is.na(Rec_ID.y)) %>%
  dplyr::select(-Rec_ID.y) %>%
  rename(Rec_ID = Rec_ID.x) %>%
  # select vessel and ticket related columns
  dplyr::select(drvid, FINAL_LENGTH, Rec_ID, represented_in_vms, DCRB_lbs, DCRB_revenue) %>%
  distinct() %>% # now, each row represents one ticket
  # calculate vessel summaries
  group_by(drvid) %>% # now, each row represents one vessel
  summarise(
    vessel_length = max(FINAL_LENGTH), # this just selects one length, since lengths are consistent
    tickets_represented_in_vms = sum(represented_in_vms),
    tickets_not_represented_in_vms = sum(!represented_in_vms),
    total_tickets = n_distinct(Rec_ID),
    percent_tickets_represented_in_vms = tickets_represented_in_vms / total_tickets * 100,
    represented_in_vms_max = max(represented_in_vms),
    represented_in_vms_majority = percent_tickets_represented_in_vms > 50, # true if over 50% of tickets have VMS data
    total_landings = sum(DCRB_lbs),
    total_revenue = sum(DCRB_revenue)
  )

# visualize vessels with VMS by vessel size
# if any ticket has VMS data, consider the vessel as represented in VMS data
dcrb_vms_representation_vessel_size_df_interpolated %>%
  ggplot(aes(x = vessel_length, color = as.factor(represented_in_vms_max), fill = as.factor(represented_in_vms_max))) +
  geom_histogram(position='dodge') +
  theme(legend.position = "bottom")

# visualize vessels with VMS by vessel size
# if >50% of vessel tickets have VMS data, consider the vessel as represented in VMS data
dcrb_vms_representation_vessel_size_df_interpolated %>%
  ggplot(aes(x = vessel_length, color = as.factor(represented_in_vms_majority), fill = as.factor(represented_in_vms_majority))) +
  geom_histogram(position='dodge') +
  theme(legend.position = "bottom")
```
