---
title: "Overlay VMS points grid of choice and calcailate value of each 5km x 5km grid cell per year for dungeness crab"
output: 
  html_document:
    toc: true
    toc_float: true
---

## Purpose

Matching the output VMS data to a multi-region grid to use for summarizing fishing effort. Then, filtering data for crab fishing events and assigning revenue and pounds of DCRB to each geolocation.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libs,message=FALSE,warning=FALSE}
library(tidyverse)
library(magrittr)
library(sf)
library(here)
library(knitr)
library(lubridate)
library(glue)
```


## Sample Year

Import files for an example year to match. We use the matched, filtered with FTID VMS data

```{r}
# matched and filteredVMS Pipeline data from 2009 produced by Blake Feist 28 May 2024
vms2009 <- read_rds(here('Confidential', 'processed', 'pipeline output', 'DCRB and CHNK 28May2024', '2009_matched_filtered_withFTID_length.rds'))
# TODO update to recent runs

# 2x2 grid shapefile
grd <- read_sf(here('GIS_layers', 'master_2km_grid_tmer.shp')) 
# TODO update to 5x5km grid
names(grd)

```

## Match Points

```{r}
# convert vms to spatial object (longitude/latitude)
vms_sf <- vms2009 %>%
  st_as_sf(coords=c('LON','LAT'),crs=4326) %>% 
  # then, convert to planar projection to match the grid
  st_transform(st_crs(grd))

# do the join
pt <- proc.time()
vms_grd_match <- vms_sf %>%
  st_join(grd)
proc.time()-pt
```

## Check

Did the join work? How many unique 5km, 25km, and 50km cells are represented in the 2009 data?

```{r checks}
names(vms_grd_match)

length(unique(vms_grd_match$CellID))

```

## All Data

We now match all of the VMS points from all years. We process each year just as one above.

### Import

```{r import all}
# clear wrkspace
rm(list=setdiff(ls(),'grd'))

# load 2009-2023 data. 
# 2009-2023 fish ticket and VMS data
# TODO update data to recent runs

# load rds files
vms_all <- vms2009 %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'pipeline output', 'DCRB and CHNK 28May2024', '2010_matched_filtered_withFTID_length.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'pipeline output', 'DCRB and CHNK 28May2024', '2011_matched_filtered_withFTID_length.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'pipeline output', 'DCRB and CHNK 28May2024', '2012_matched_filtered_withFTID_length.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'pipeline output', 'DCRB and CHNK 28May2024', '2013_matched_filtered_withFTID_length.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'pipeline output', 'DCRB and CHNK 28May2024', '2014_matched_filtered_withFTID_length.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'pipeline output', 'DCRB and CHNK 28May2024', '2015_matched_filtered_withFTID_length.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'pipeline output', 'DCRB and CHNK 28May2024', '2016_matched_filtered_withFTID_length.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'pipeline output', 'DCRB and CHNK 28May2024', '2017_matched_filtered_withFTID_length.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'pipeline output', 'DCRB and CHNK 28May2024', '2018_matched_filtered_withFTID_length.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'pipeline output', 'DCRB and CHNK 28May2024', '2019_matched_filtered_withFTID_length.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'pipeline output', 'DCRB and CHNK 28May2024', '2020_matched_filtered_withFTID_length.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'pipeline output', 'DCRB and CHNK 28May2024', '2021_matched_filtered_withFTID_length.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'pipeline output', 'DCRB and CHNK 28May2024', '2022_matched_filtered_withFTID_length.rds'))) %>%
  bind_rows(read_rds(here('Confidential', 'processed', 'pipeline output', 'DCRB and CHNK 28May2024', '2023_matched_filtered_withFTID_length.rds')))

glimpse(vms_all)

```

### Match

```{r}
# convert VMS to spatial object
pt <- proc.time()
vms_all_sf <- vms_all %>%
  st_as_sf(coords=c('LON','LAT'),crs=4326) %>% 
  # then, convert to planar projection to match the grid
  st_transform(st_crs(grd))
x<-proc.time()-pt
```

Took `r round(x[3]/60,2)` minutes to do the conversion. Now for the join...

```{r join to grid}
# do the join
pt <- proc.time()
vms_all_grd_match <- vms_all_sf %>%
  st_join(grd)
x<-proc.time()-pt
```

The join took`r round(x[3]/60,2)` minutes.

### Save and Write

```{r}
# non-spatial version
glimpse(vms_all_grd_match)
vms_all_grd_match %>% 
  st_set_geometry(NULL) %>% 
  write_rds(here('Confidential', 'processed', 'pipeline output', 'DCRB and CHNK 28May2024', 'vms_all_w_grd_2km.rds'))
# TODO rename output
```

## Prep data for DCRB focused analysis

Ok now that we have the VMS data on the 2km grid, we need to prepare it for analysis as in Samhouri et al. 2021 PRSB, with a focus on DCRB fishing effort.

The code below is based on prep_data_for_scenario_df_function.R

Start by choosing which variables to use to filter to fishing locations only.

```{r fishing data prep 1}

vms_all_grd_match <- read_rds(here('Confidential', 'processed', 'pipeline output', 'DCRB and CHNK 28May2024', 'vms_all_w_grd_2km.rds'))
# TODO update data source
glimpse(vms_all_grd_match)

vms_all_grd_match <- vms_all_grd_match %>% mutate(
  year = lubridate::year(westcoastdate_notime)
)
# TODO add month

# check years included
unique(vms_all_grd_match$year)

#####################################################
# filter the data using the control variables below
# years <- seq(2009,2019,1)
# TODO add iteration over months

# state_agency_code <- "C"
target_rev <- "DCRB"
target_lbs <- "DCRB"
winter_months <- c("November", "December", "January", "February", "March")
removal_types <- c("COMMERCIAL (NON-EFP)", "COMMERCIAL(DIRECT SALES)", "UNKNOWN")
max_speed <- 4.11556 # units of m/s. 4.11556m/s = 8knots (4DCRB) and 1.54333 m/s = 3 knots (4CHNK)
min_speed <- 0 # units = m/s
# placeholder for adding daytime filter appropriate for salmon trolling boats.
# Potentially use suncalc package to use lat, lon, local date/time ('westcoastdate' attribute)
# to filter out points that occured in non-daylight times 

```

Now filter the data based on control variables above: California, DCRB trips only, commercial removal types only, slow vessel speeds. Need to add DEPTH_CATM, but for now ask Blake to filter based on NGDC_M or the depth of each grid cell.

```{r filter fishing data}

# subset the data based on above queries
dcrb_vms_tix_analysis <- vms_all_grd_match %>%
  # filter(agency_code == state_agency_code) %>%
  filter(TARGET_rev == target_rev | TARGET_lbs == target_lbs) %>%
  filter(removal_type_name %in% removal_types) %>%
  #filter(DEPTH_CATM == "0-100m" | DEPTH_CATM == "100-150m") %>%
  #filter(NGDC_M <= 0 & NGDC_M >= -12000) %>% # considered adding 052220. also see "evaluate fishing depths.R". probably unnecessary because NGDC_M refers to centroid depth and we can match depths by 5km grid cell later
  filter(avg_speed_recalc <= max_speed & avg_speed_recalc >= min_speed) %>%
  #filter(is.na(in_port) == TRUE) %>% # only removes ~4000 records
  #filter(port_group_code %in% ports) %>%
  mutate(
         year_month = paste0(lubridate::year(westcoastdate_notime),"_", substr(lubridate::ymd(westcoastdate_notime),6,7)), # substr() ensures month is a 2 digit value
         month = lubridate::month(westcoastdate_notime, label=TRUE, abbr = FALSE),
         month_as_numeric = month(westcoastdate_notime),
         week_of_year = week(westcoastdate_notime),
         day_of_year = yday(westcoastdate_notime),
         season = as.character(ifelse(month %in% winter_months, "Winter", "Spring-Summer")),
         crab_year = ifelse(
           month_as_numeric >= 11, paste0(year,"_",1+year), paste0(year - 1,"_",year)
         )
  ) %>%
  # filter(westcoastdate_notime >= as.Date("2009-11-01") & westcoastdate_notime <= as.Date("2019-08-01")) %>%
  # #filter(month_as_numeric %in% crab_months) %>%
  dplyr::select(
    Rec_ID, VMS_RECNO, drvid,
    westcoastdate_notime, year, crab_year, year_month, month, month_as_numeric, week_of_year, day_of_year, season,
    CellID, pacfin_port_code, port_group_code, NGDC_M,
    TARGET_lbs, TARGET_rev, DCRB_lbs, DCRB_revenue, all_species_lbs, all_species_rev, species_code_all
    ) # DEPTH_CATM, 

glimpse(dcrb_vms_tix_analysis)

```

Distribute rev and lbs to grid cells. 

```{r distribute $ and lbs to cells}

# TODO check - are pings in the files from Blake? I don't remember it in walk through with Blake or with Owen...

### create new columns that apportion pings/lbs/$ from each fish ticket based on proportion of pings related to each ticket in each grid cell

# total records per trip (total number of VMS records associated with each fish ticket)
VMSrecords_per_trip <- dcrb_vms_tix_analysis %>%
  group_by(Rec_ID) %>%
  summarise(trip_VMSrecords = n(),
            .groups = 'drop') #%>%
  #filter(trip_VMSrecords > 1)

# add to fish ticket / vms data, make columns vessels, lbs, and $ per VMS location. note that columns with "...." in them, like ""TARGET...." reflect revenue 
dcrb_vms_tix_analysis_TripInfo <- left_join(VMSrecords_per_trip, dcrb_vms_tix_analysis, by="Rec_ID") %>%
  mutate(
    DCRB_lbs_per_VMSlocation = DCRB_lbs/trip_VMSrecords, # TODO understand trip_VMSrecords
    DCRB_rev_per_VMSlocation = DCRB_revenue/trip_VMSrecords,
    all_species_lbs_per_VMSlocation = all_species_lbs/trip_VMSrecords,
    all_species_rev_per_VMSlocation = all_species_rev/trip_VMSrecords,
    Num_DCRB_Vessels_per_VMSlocation = 1/trip_VMSrecords
  )
glimpse(dcrb_vms_tix_analysis_TripInfo)

# could add in depth filter by excluding all NGDC_M values >-1500

# write out for Blake
write_rds(dcrb_vms_tix_analysis_TripInfo,here::here('Confidential', 'processed', 'pipeline output', 'DCRB and CHNK 28May2024', 'DCRB_vms_tix_daily_2009_2023_lbs_rev_per_geolocation_nodepthfilter.rds'))

#####################################################

#####################################################

### at long last, make the df we want

### all vessels

start.time <- Sys.time()

# TODO I think this needs some values removed from group_by?
con_df_daily_years_2km <- dcrb_vms_tix_analysis_TripInfo %>%
  group_by(year, crab_year, year_month, season, month, month_as_numeric, week_of_year, day_of_year, CellID) %>%
  summarise(
    DCRB_lbs = sum(DCRB_lbs_per_VMSlocation),
    DCRB_rev = sum(DCRB_rev_per_VMSlocation),
    all_species_lbs = sum(all_species_lbs_per_VMSlocation),
    all_species_rev = sum(all_species_rev_per_VMSlocation),
    Num_DCRB_VMS_pings = n(),
    Num_DCRB_Vessels = sum(Num_DCRB_Vessels_per_VMSlocation),
    Num_Unique_DCRB_Vessels = length(unique(as.character(drvid))),
    mean_depth = mean(NGDC_M)
  ) %>%
  ungroup() 
Sys.time() - start.time

glimpse(con_df_daily_years_2km)

### if and when satisfied, write out rds

write_rds(con_df_daily_years_2km,here::here('Confidential', 'processed', 'pipeline output', 'DCRB and CHNK 28May2024', 'DCRB_vms_tix_daily_2009_2023_lbs_rev_per_2kmgrid_nodepthfilter.rds'))
#TODO update output name
```


## Old note from Jameal: Did not yet, but could summarise by yr, yr_month, etc

```{r}
#TODO remove this section?

## code from Kelly for creating dfs for 1) drvid confidentiality (vessels_SPECIES) and 2) pings/grid cell/year with lbs and rev
#unique vessel ids by year
# load raw DCRB pings 2009-2023 (sub CHNK for chinook)
DCRB_vms_tix_daily_2009_2023_lbs_rev_per_geolocation_nodepthfilter <- readRDS("~/Documents/GitHub/VMS-pipeline/Confidential/processed/pipeline output/DCRB and CHNK 28May2024/DCRB_vms_tix_daily_2009_2023_lbs_rev_per_geolocation_nodepthfilter.rds")

vessels_DCRB <- DCRB_vms_tix_daily_2009_2023_lbs_rev_per_geolocation_nodepthfilter %>%
dplyr::select(CellID, year, drvid) %>%
distinct()

# save the df (sub CHNK for chinook_)
write_rds(vessels_DCRB,here::here('Confidential', 'processed', 'pipeline output', 'DCRB and CHNK 28May2024', 'vessels_4_confidentiality_DCRB.rds'))


## figuring out whether unique vessel ids are summed per year?
# load DCRB pings overlaid on 2km grid (sub CHNK for chinook)
DCRB_vms_tix_daily_2009_2023_lbs_rev_per_2kmgrid_nodepthfilter <- readRDS("~/Documents/GitHub/VMS-pipeline/Confidential/processed/pipeline output/DCRB and CHNK 28May2024/DCRB_vms_tix_daily_2009_2023_lbs_rev_per_2kmgrid_nodepthfilter.rds")

test_sum_DCRB <- DCRB_vms_tix_daily_2009_2023_lbs_rev_per_2kmgrid_nodepthfilter 

## create output table with unique vessels, target rev/lba and other rev/lbs, by year (sub CHNK for chinook)
pings_per_gridcell_DCRB <- test_sum_DCRB %>%
group_by(CellID, year) %>%
summarise(DCRB_lbs = sum(DCRB_lbs, na.rm = TRUE),
    DCRB_rev = sum(DCRB_rev, na.rm = TRUE),
    #all_species_lbs = sum(all_species_lbs, na.rm = TRUE),
    #all_species_rev = sum(all_species_rev, na.rm = TRUE),
    Num_DCRB_VMS_pings = n())
    
# save the df (sub CHNK for chinook)
write_rds(pings_per_gridcell_DCRB,here::here('Confidential', 'processed', 'pipeline output', 'DCRB and CHNK 28May2024', 'pings_per_gridcell_DCRB.rds'))
```