---
title: "03_blh_OR-dungeness-crab-permit-data"
format: html
  html:
    embed-resources: true
editor: visual
---

## Purpose

Explore permit tier data.

## Set up

```{r}
# import libraries
library(here)
library(readxl)
library(tidyverse)
library(scales)
```

```{r}
# import data
permit_df <- read_excel(here('Confidential', 'data', 'raw', 'logbook', 'OregonCrabPermitData2007-2020.xlsx'), sheet = "CrabPermitData")
```

## Explorations

How many permits and vessels per year? does permit pot limit vary from year to year for the same vessel? does a single vessel have overlapping permits? are those permits unique or the same? does one permit only ever have one boat? does one vessel only have one permit id?

-   `Number` = Permit ID
-   `Docnum` = Vessel ID

```{r}
# how big is it?
dim(permit_df)

# what's in here?
glimpse(permit_df)

# which years?
sort(unique(permit_df$Year))

# what types of permits?
unique(permit_df$Type)

# how many nulls?
summarise_all(permit_df, ~sum(is.na(.)))

# how many unique permit identifiers?
n_distinct(permit_df$Number)

# how many unique vessel identifiers?
n_distinct(permit_df$Docnum)

# what's the distribution of permit start dates?
permit_df %>% ggplot(aes(x = Begindate)) +
  geom_histogram(bins = 100) + scale_x_datetime(labels = date_format("%Y"), date_breaks = "year")

# what's the distribution of permit end dates?
permit_df %>% ggplot(aes(x = Enddate)) +
  geom_histogram(bins = 100) + scale_x_datetime(labels = date_format("%Y"), date_breaks = "year")

# add columns to explore permit length, begin and end dates
permit_df <- permit_df %>%
  mutate(
    permit_length_days = as.numeric(difftime(Enddate, Begindate, units = "days")),
    begin_year_day = yday(Begindate),
    end_year_day = yday(Enddate)
  )

# what's the distribution of permit length?
permit_df %>%
  ggplot(aes(x = permit_length_days)) + 
  geom_histogram(binwidth = 7)

# are any permits longer than a year?
sum(permit_df$permit_length_days > 365)

# how many records have a begin date after an end date?
sum(permit_df$Begindate > permit_df$Enddate)

# how many records have the same begin and end dates?
sum(permit_df$Begindate == permit_df$Enddate)

# when do permits begin in the year?
range(permit_df$begin_year_day)
permit_df %>% 
  ggplot(aes(x = begin_year_day)) + 
  geom_histogram()

# how many permits begin on the first day of the year?
sum(permit_df$begin_year_day == 1)

# when do permits end in the year?
range(permit_df$end_year_day)
permit_df %>% 
  ggplot(aes(x = end_year_day)) + 
  geom_histogram()

# how many permits end on the last day of the year?
# (not perfect, but including 266 for leap years. the 365 number may be an over estimate, 2nd to last day in leap year.)
sum(permit_df$end_year_day == 365)
sum(permit_df$end_year_day == 366)

# do permits overlap calendar years?
sum(year(permit_df$Begindate) == year(permit_df$Enddate))
```

Takeaways

-   This dataset is pretty small.
-   There is no unique identifier in the dataset, the permit number doesn't necessarily change.
-   Permits are maximum one year in length.
-   The majority of permits begin at the beginning of the calendar year, and end at the end of the calendar year.
-   The majority of permits begin and end in the same calendar year.
-   Some permits have begin date after end date, which is presumably a data entry error.

Cleaning decisions

-   If the permit begin date is after end date, then swap the dates
