---
title: "Hindcasts of California Dungeness crab fishing activity, 2011-2023"
author: "Brooke Hawkins"
date: "`r Sys.Date()`"
output: html_document
knit: |
  (function(input, ...) {
    rmarkdown::render(
      input,
      output_file = paste0(xfun::sans_ext(input), '-', Sys.Date(), '.html')
    )
  })
---

```{r, include=FALSE}
# start timer
start_timer <- proc.time()
```

## Purpose

* Integrate interpolated VMS and fish ticket data to produce time series, rasters, and maps of fishing activity, landed weight, and revenue.
* Deliverable: Monthly maps of fishing activity; tables summarizing monthly fishing activity

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# import libraries
library(tidyverse)
library(here)
library(fredr)
library(sf)
library(rnaturalearth)
library(viridis)
library(magick)

# adjust ggplot theme
theme_replace(axis.text.x=element_text(angle=45, vjust=1, hjust=1),
              axis.ticks.x=element_blank(),
              axis.ticks.y=element_blank())
```

## Load data

1. Load the joined, cleaned, interpolated VMS and fish ticket data for specified years, which will be used to map fishing activity distribution. 
2. Load the cleaned fish ticket data for specified years, which will be used to check how representative the VMS dataset is of all dungeness crab landings. 
3. Load and join the spatial 5km grid, which will also be used to map fishing activity distribution.

Joining the spatial 5km gird is one of the slowest parts of the script.

Pre-requisite: Run the pipeline steps 1-6 (including interpolation) for calendar years 2011-2023 for DCRB.

```{r load-data}
# choose years of data to load
load_years <- 2011:2023

# load VMS and fish ticket data
vms_df <- purrr::map(load_years, function(ly) {
  read_rds(here('Confidential', 'processed_data', 'processed_2025-03-19', 'interpolated', paste0('interpolated_', ly, '.rds')))
}) %>% bind_rows()

# load fish ticket data before it was joined to VMS data, used to check VMS representativeness
ticket_df <- purrr::map(load_years, function(ly) {
  read_rds(here('Confidential', 'processed_data', 'processed_2025-03-19', 'fish_tickets', paste0('fishtix_vlengths_withFTID_', ly, '.rds')))
}) %>% bind_rows()

# load 5km grid shape file
grid_5km <- read_sf(here('spatial_data', 'fivekm_grid_polys_shore_lamb.shp'))

# join 5km grid to VMS and fish ticket data
vms_df <- vms_df %>%
  st_as_sf(coords = c('LON', 'LAT'), crs=4326) %>%
  st_transform(st_crs(grid_5km)) %>%
  st_join(grid_5km) %>%
  st_set_geometry(NULL)
```

```{r create-output-directories}
# make hidncast output directory, if doesn't yet exist
output_dir <- here('Confidential', 'hindcast_output')
if (!dir.exists(output_dir)) dir.create(output_dir)

# create hindcast output subdirectory name based on system date
output_subdir_name <- paste0("hindcast_output_", Sys.Date())

# make hindcast output subdirectory, if doesn't yet exist
output_subdir <- here('Confidential', 'hindcast_output', output_subdir_name)
if (!dir.exists(output_subdir)) dir.create(output_subdir)

# make tables, figures, and maps subdirectories
for (temp_name in c('tables', 'figures', 'maps')) {
  temp_subdir <- here('Confidential', 'hindcast_output', output_subdir_name, temp_name)
  if (!dir.exists(temp_subdir)) dir.create(temp_subdir)
}
rm(temp_name, temp_subdir)
```


4. Load inflation adjustment factors from FRED.

Adjust revenue for inflation using Fred GDP data, adapted from R code by Erin Steiner. This product uses the FRED® API but is not endorsed or certified by the Federal Reserve Bank of St. Louis.

Pre-requisite: Create a FRED account and API key. See https://fred.stlouisfed.org/docs/api/api_key.html.

```{r inflation-adjustment}
# insert your FRED API key
fredr_set_key('')

# download the quarterly inflation adjustments from Fred from 1985 to present
fred_gdpdefl <- fredr(
    series_id = "GDPDEF", 
    observation_start = as.Date(paste0(min(load_years), "-01-01"))
  )

# generate the mean annual deflators based on the quarterly values
gdp_defl <- mutate(fred_gdpdefl, year = year(date)) %>%
  group_by(year) %>%
  summarize(defl = mean(value), .groups = 'drop') %>%
  mutate(inflation_adjustment_factor = defl / defl[year == max(year)]) %>%
  select(year, inflation_adjustment_factor)

# write inflation adjustment factors
write.csv(file = here("Confidential", "hindcast_output", output_subdir_name, "tables", "inflation_adjustment.csv"), x = gdp_defl)

# plot inflation adjustment factors
gdp_defl %>% ggplot(aes(x = as.factor(year), y = inflation_adjustment_factor)) + geom_point()
ggsave(here("Confidential", "hindcast_output", output_subdir_name, "figures", "inflation_adjustment.png"))

# clean up
rm(fred_gdpdefl)
```

The reference year for inflation adjustment is `r max(gdp_defl$year)`.

## Transform data

Transform the joined, cleaned, interpolated VMS and fish ticket data:

1. Filter for dungeness crab related records, California landings, min/max depth and speed filters.
2. Add temporal columns.
3. Adjust revenue for inflation with data from FRED.

Some commonly used acronyms for variable naming in the code include:

* `dcrb` dungeness crab
* `rev` revenue
* `lbs` landings
* `VMS` vessel monitoring system
* `afi` adjusted for inflation

```{r transform-vms-data}
# define filters
target_rev <- "DCRB"         # revenue target
target_lbs <- "DCRB"         # landings target
min_depth <- 0               # minimum depth in meters
max_depth <- -150            # maximum depth in meters
min_speed <- 0               # minimum speed in m/s
max_speed <- 4.11556         # maximum speed in m/s (4.11556 m/s = 8 knots)
crab_year_start <- 11        # month defines start of crab year
winter_months <- c("November", "December", "January", "February", "March") # determine Winter or Spring-Summer season

# transform VMS and fish ticket data
dcrb_df <- vms_df %>%
  # add temporal columns
  mutate(
    year = lubridate::year(westcoastdate_notime),
    month = lubridate::month(westcoastdate_notime, label=TRUE, abbr=FALSE),
    month_numeric = month(westcoastdate_notime),
    year_month = paste0(lubridate::year(westcoastdate_notime),"_", substr(lubridate::ymd(westcoastdate_notime), 6, 7)),
    crab_year = ifelse(month_numeric >= crab_year_start, paste0(year, "_", 1+year), paste0(year-1, "_", year)),
    season = as.character(ifelse(month %in% winter_months, "Winter", "Spring-Summer")),
    week_of_year = week(westcoastdate_notime),
    day_of_year = yday(westcoastdate_notime)
  ) %>%
  # apply filters
  filter(TARGET_rev == target_rev | TARGET_lbs == target_lbs) %>%
  filter(NGDC_M <= min_depth & NGDC_M >= max_depth) %>%
  filter(avg_speed_recalc <= max_speed & avg_speed_recalc >= min_speed) %>% 
  # join inflation adjustment factor and adjust revenue
  left_join(gdp_defl, by = join_by(year)) %>%
  mutate(DCRB_revenue_afi = DCRB_revenue / inflation_adjustment_factor) %>%
  # select columns
  dplyr::select(
    # identifiers
    Rec_ID,           # fish ticket ID
    VMS_RECNO,        # VMS ping ID
    drvid,            # vessel ID
    pacfin_port_code, # port ID
    port_group_code,  # port group ID
    GRID5KM_ID,       # grid cell ID
    # temporal fields
    westcoastdate,
    westcoastdate_notime,
    crab_year, 
    season,
    year,
    year_month,
    month, 
    month_numeric,
    week_of_year,
    day_of_year,
    # vessel length
    FINAL_LENGTH,
    # dungeness crab fields
    DCRB_lbs,
    DCRB_revenue_afi
  )

# count total records, trips and years
n_records <- nrow(dcrb_df)
n_trips   <- n_distinct(dcrb_df$Rec_ID)
n_years   <- n_distinct(lubridate::year(dcrb_df$westcoastdate_notime))

# take a peek at the resulting dataframe
glimpse(dcrb_df)
```

The dungeness crab VMS dataframe has `r n_records` records (interpolated VMS pings), `r n_trips` trips (fish tickets), across `r n_years` years.

```{r}
# clean up
remove(vms_df, n_records, n_trips, n_years)
```

Next, create a monthly gridded summary of fishing activity. 

1. Add trip-level attributes:

* `trip_n_vms_records`: the number of VMS records associated with a fishing trip (fish ticket)

2. Add VMS record attributes, calculated using the trip-level attribute:

* `dcrb_lbs_per_vms_record`: the total landings for a fishing trip, divided by the number of VMS records associated with that trip (e.g. a fish ticket with 1200 lbs landed and 20 VMS records would have 60 lbs per VMS record)
* `dcrb_rev_per_vms_record`: the total revenue for a fishing trip, divided by the number of VMS records associated with that trip (e.g. a fish ticket with $4000 revenue and 20 VMS pings would have \$200 revenue per VMS record)

3. Summarize fishing activity at a monthly gridded level:

* `dcrb_lbs`: the total landings per VMS record in a given grid cell in a given month
* `dcrb_rev`: the total revenue per VMS record in a given grid cell in a given month
* `n_vms_records`: the number of VMS records in a given grid cell in a given month
* `n_unique_vessels`: the number of unique vessels in a given grid cell in a given month
* `confidential_flag`: determines whether the grid cell month has <3 unique vessels; in which case, the data in this grid cell and month needs to be excluded from non-confidential reports

```{r gridded-summary}
# summarize trip-level attributes
trip_df <- dcrb_df %>%
  group_by(Rec_ID) %>%
  summarise(trip_n_vms_records = n())

# join trip-level attributes back to dungeness crab dataframe
dcrb_trip_df <- left_join(dcrb_df, trip_df, by="Rec_ID") %>%
  mutate(
    dcrb_lbs_per_vms_record     = DCRB_lbs / trip_n_vms_records,
    dcrb_rev_per_vms_record     = DCRB_revenue_afi / trip_n_vms_records
  )

# summarize grid cell and month-level attributes
confidential_df <- dcrb_df %>%
  group_by(GRID5KM_ID, year_month) %>%
  summarise(n_unique_vessels = n_distinct(drvid),
            # set confidential_flag TRUE for confidential, FALSE for non-confidential
            confidential_flag = (n_distinct(drvid) < 3), .groups="keep") %>%
  ungroup()

# join grid cell and month-level attributes back to dungeness crab dataframe
dcrb_confidential_df <- dcrb_trip_df %>%
  left_join(confidential_df, by=join_by(GRID5KM_ID, year_month))

# create monthly gridded summary
dcrb_5km_summary_df <- dcrb_confidential_df %>%
  group_by(
    # grid cell ID, makes this a gridded summary
    GRID5KM_ID,
    # temporal columns, makes this a monthly summary
    year,
    crab_year,
    year_month,
    season,
    month,
    month_numeric
  ) %>%
  # sum landings, revenue; count unique vessels and VMS records; determine confidentiality filter
  summarise(
    dcrb_lbs = sum(dcrb_lbs_per_vms_record),
    dcrb_rev = sum(dcrb_rev_per_vms_record),
    n_vms_records = n(),
    n_unique_vessels = n_distinct(drvid),
    # set confidential_flag TRUE for confidential, FALSE for non-confidential
    confidential_flag = max(confidential_flag) == 1,
    .groups = "keep"
  ) %>% 
  ungroup()

# write monthly gridded summary
write.csv(file=here("Confidential", "hindcast_output", output_subdir_name, "tables", "dcrb_5km_summary_df.csv"), x=dcrb_5km_summary_df)

# clean up
remove(trip_df, dcrb_trip_df, confidential_df)
```

Repeat the for fish ticket data, which will be used later to evaluate the representativeness of the VMS data for dungeness crab fishing activity.

```{r transform-ticket-data}
# select distinct fish ticket IDs from filtered VMS dataframe
dcrb_vms_rec_ids <- dcrb_confidential_df %>% 
  distinct(Rec_ID) %>% 
  mutate(vms_represented_flag = TRUE)

# apply same filters as VMS dataframe (except for depth and speed filters, which are from VMS data)
dcrb_ticket_df <- ticket_df %>%
  # add temporal columns
  mutate(
    year = lubridate::year(date),
    month = lubridate::month(date, label=TRUE, abbr=FALSE),
    month_numeric = month(date),
    year_month = paste0(lubridate::year(date),"_", substr(lubridate::ymd(date), 6, 7)),
    crab_year = ifelse(month_numeric >= crab_year_start, paste0(year, "_", 1+year), paste0(year-1, "_", year)),
    season = as.character(ifelse(month %in% winter_months, "Winter", "Spring-Summer")),
    week_of_year = week(date),
    day_of_year = yday(date)
  ) %>%
  # apply filters
  filter(TARGET_rev == target_rev | TARGET_lbs == target_lbs) %>%
  mutate(
    year_month = paste0(lubridate::year(date),"_", substr(lubridate::ymd(date),6,7)),
    month_as_numeric = month(date),
    month_as_date = round_date(date, unit='month'),
    season = as.character(ifelse(month %in% winter_months, "Winter", "Spring-Summer")),
    crab_year = ifelse(
      month_as_numeric >= 11, paste0(year, "_", 1+year), paste0(year-1, "_", year)
    )
  ) %>%
  # join inflation adjustment factor and adjust revenue
  left_join(gdp_defl, by = join_by(year)) %>%
  mutate(DCRB_revenue_afi = DCRB_revenue / inflation_adjustment_factor) %>%
  # select columns
  dplyr::select(
    # identifiers
    Rec_ID,           # fish ticket ID
    drvid,            # vessel ID
    pacfin_port_code, # port ID
    port_group_code,  # port group ID
    # temporal fields
    date,
    crab_year,
    season,
    year,
    year_month,
    month,
    month_numeric,
    week_of_year,
    day_of_year,
    # vessel length
    FINAL_LENGTH,
    # dungeness crab fields
    DCRB_lbs,
    DCRB_revenue_afi
  ) %>%
  # add column to ticket dataframe for whether that ticket was represented by VMS data
  left_join(dcrb_vms_rec_ids, by = join_by(Rec_ID)) %>%
  mutate(vms_represented_flag = if_else(is.na(vms_represented_flag), FALSE, TRUE),
         vms_represented_drvid = if_else(vms_represented_flag, drvid, NA))

# take a peek at the resulting dataframe
glimpse(dcrb_ticket_df)

# clean up
remove(ticket_df)
```

## Create outputs

Output files will be written to `Confidential/hindcast_output`, into three subdirectories: `figures`, `maps`, and `tables`.

Create some overall **time series** and **heatmaps** of fishing activity, revenue and landings. This does not currently use the confidential output, since it is not reporting monthly gridded data, but it can be updated to do so.

```{r fishing-activity-time-series}
# this uses the confidential data, using non-gridded summary to not double count unique vessels
time_series_df <- dcrb_confidential_df %>%
  # add year and month as a date for easier plotting
  mutate(year_month_date = ym(year_month)) %>%
  # group by month and summarize
  group_by(crab_year, year, year_month, year_month_date, month) %>%
  summarise(n_vms_records = n(),
            n_unique_vessels = n_distinct(drvid),
            DCRB_lbs = sum(DCRB_lbs),
            DCRB_revenue_afi = sum(DCRB_revenue_afi),
            .groups = 'drop') %>%
  # group by crab year and calculate cumulative sums
  group_by(crab_year) %>%
  mutate(cumulative_sum_DCRB_lbs = cumsum(DCRB_lbs),
         cumulative_sum_DCRB_revenue_afi = cumsum(DCRB_revenue_afi)) %>%
  ungroup()

# set x-axis label date breaks for time to start x-axis labels in Nov.
set_breaks <- scales::breaks_width("1 year", offset = -61) # -61 days moves start from Jan. 1 to Nov. 1
df_date_breaks <- set_breaks(range(time_series_df$year_month_date))

# set vector of variables to create line plots for
plot_vars <- c('n_vms_records', 'n_unique_vessels', 'DCRB_lbs', 'DCRB_revenue_afi', 'cumulative_sum_DCRB_lbs', 'cumulative_sum_DCRB_revenue_afi')
# create line plots
for (plot_var in plot_vars) {
  time_series_df %>% ggplot() + 
    geom_line(aes(x = year_month_date, y = get(plot_var), color = crab_year)) +
    scale_x_date(breaks = df_date_breaks, date_labels = "%b %Y")
  ggsave(filename = here("Confidential", "hindcast_output", output_subdir_name, "figures", paste0("time_series_", plot_var, ".png")))
}

# plot # VMS pings per month as heatmap
time_series_df %>% ggplot() + geom_tile(aes(x = month, y = as.factor(year), fill = n_vms_records))
ggsave(filename = here("Confidential", "hindcast_output", output_subdir_name, "figures", "heatmap_n_vms_records.png"))

# plot # unique vessels per month as heatmap
time_series_df %>% ggplot() + geom_tile(aes(x = month, y = as.factor(year), fill = n_unique_vessels))
ggsave(filename = here("Confidential", "hindcast_output", output_subdir_name, "figures", "heatmap_n_unique_vessels.png"))
```

Create a table and a bar chart that summarize the **representativeness** of VMS data each year. Metrics reported are:

* *Tickets*: (tickets with matched VMS records) / (tickets)
* *Landings*: (landings from tickets with matched VMS records) / (landings)
* *Revenue*: (revenue from tickets with matched VMS records) / (revenue)
* *Vessels*: (unique vessels on tickets with matched VMS records) / (unique vessels)

```{r representativeness-tables}
# check yearly VMS representativeness for tickets, landings, revenue and vessels
representativeness_yearly_table <- dcrb_ticket_df %>%
  group_by(year) %>%
  summarise(
    # check for tickets
    vms_represented_tickets          = sum(vms_represented_flag),
    total_tickets                    = n(),
    percent_vms_represented_tickets  = trunc(vms_represented_tickets / total_tickets * 10000) / 100,
    # repeat for landings
    vms_represented_landings         = sum(vms_represented_flag * DCRB_lbs),
    total_landings                   = sum(DCRB_lbs),
    percent_vms_represented_landings = trunc(vms_represented_landings / total_landings * 10000) / 100,
    # repeat for revenue
    vms_represented_revenue          = sum(vms_represented_flag * DCRB_revenue_afi),
    total_revenue                    = sum(DCRB_revenue_afi),
    percent_vms_represented_revenue  = trunc(vms_represented_revenue / total_revenue * 10000) / 100,
    # repeat for vessels
    vms_represented_vessels          = n_distinct(vms_represented_drvid, na.rm=TRUE),
    total_vessels                    = n_distinct(drvid),
    percent_vms_represented_vessels  = trunc(vms_represented_vessels / total_vessels * 10000) / 100,
    .groups = 'keep'
  ) %>%
  ungroup() %>%
  arrange(year)

# write yearly result
write.csv(file = here("Confidential", "hindcast_output", output_subdir_name, "tables", "table_vms_representativeness_yearly.csv"),
          x = representativeness_yearly_table)
```

```{r representativeness-time-series}
# plot % represented by VMS for all four metrics, yearly
vms_representativeness_time_series_yearly <- representativeness_yearly_table %>%
  select(year,
         contains("percent_vms_represented")) %>%
  pivot_longer(cols = contains("percent_vms_represented"),
               names_to = "metric",
               names_prefix = "percent_vms_represented_",
               values_to ="percent_vms_represented") %>%
  ggplot(aes(x = as.factor(year), y = percent_vms_represented)) +
  geom_col() +
  xlab("year") +
  facet_grid(metric ~ ., scales = "free_y")
ggsave(filename=here("Confidential", "hindcast_output", output_subdir_name, "figures", "time_series_representativeness_yearly.png"))
vms_representativeness_time_series_yearly
```

Repeat the representativeness table and bar chart, stratified by **vessel size**. Vessel size will be grouped into small vessels (with vessel length less than or equal to `vessel_size_cutoff` feet) and large vessels (greater than `vessel_size_cutoff` feet).

```{r size-stratified-representativeness-tables}
# set a cutoff (in feet) for vessel size stratification in VMS representativeness tables and figures
vessel_size_cutoff <- 40

# check yearly VMS representativeness for tickets, landings, revenue and vessels
stratified_representativeness_yearly_table <- dcrb_ticket_df %>%
  mutate(vessel_size_above_cutoff = FINAL_LENGTH > vessel_size_cutoff) %>%
  group_by(year, vessel_size_above_cutoff) %>%
  summarise(
    # check for tickets
    vms_represented_tickets          = sum(vms_represented_flag),
    total_tickets                    = n(),
    percent_vms_represented_tickets  = trunc(vms_represented_tickets / total_tickets * 10000) / 100,
    # repeat for landings
    vms_represented_landings         = sum(vms_represented_flag * DCRB_lbs),
    total_landings                   = sum(DCRB_lbs),
    percent_vms_represented_landings = trunc(vms_represented_landings / total_landings * 10000) / 100,
    # repeat for revenue
    vms_represented_revenue          = sum(vms_represented_flag * DCRB_revenue_afi),
    total_revenue                    = sum(DCRB_revenue_afi),
    percent_vms_represented_revenue  = trunc(vms_represented_revenue / total_revenue * 10000) / 100,
    # repeat for vessels
    vms_represented_vessels          = n_distinct(vms_represented_drvid, na.rm=TRUE),
    total_vessels                    = n_distinct(drvid),
    percent_vms_represented_vessels  = trunc(vms_represented_vessels / total_vessels * 10000) / 100,
    .groups = 'keep'
  ) %>%
  ungroup() %>%
  arrange(year, vessel_size_above_cutoff)

# write yearly result
write.csv(file = here("Confidential", "hindcast_output", output_subdir_name, "tables", "table_vms_representativeness_yearly_stratified.csv"),
          x = stratified_representativeness_yearly_table)
```

```{r size-stratified-representativeness-time-series}
# plot % represented by VMS for all four metrics stratified by vessel size category, yearly
stratified_vms_representativeness_time_series_yearly <- stratified_representativeness_yearly_table %>%
  select(year,
         vessel_size_above_cutoff,
         contains("percent_vms_represented")) %>%
  pivot_longer(cols = contains("percent_vms_represented"),
               names_to = "metric",
               names_prefix = "percent_vms_represented_",
               values_to ="percent_vms_represented") %>%
  ggplot(aes(x = as.factor(year), y = percent_vms_represented, fill = vessel_size_above_cutoff)) +
  geom_col(position = "dodge") +
  xlab("year") +
  facet_grid(metric ~ ., scales = "free_y")
ggsave(filename=here("Confidential", "hindcast_output", output_subdir_name, "figures", "time_series_representativeness_yearly_stratified.png"))
stratified_vms_representativeness_time_series_yearly
```

For the remaining figures, decide whether your outputs need to be confidential or not, using the `confidential_output_flag`. Set to `FALSE` to create non-confidential output. If a grid cell in a given month has < 3 unique vessels, it cannot be included in non-confidential maps.

```{r decide-on-confidential-output}
# set whether to use the non-confidential or confidential gridded summary for output
confidential_output_flag <- TRUE

if (confidential_output_flag) {
  # if confidential output is ok, then don't remove any VMS records
  dcrb_5km_output_df <- dcrb_5km_summary_df
} else {
  # if non-confidential output is needed, then remove any VMS records where the grid cell and month had < 3 unique vessels, where confidential_flag is FALSE
 dcrb_5km_output_df <- dcrb_5km_summary_df %>% filter(!confidential_flag)
}
```

Regardless of whether you're using confidential or non-confidential output, create a table summarizing how many VMS records are filtered out by the **confidentiality** flag each year.

```{r confidentiality-table}
# calculate the % of VMS records dropped by confidentiality filter, for the entire time period
numerator_drop_confidential <- sum(dcrb_confidential_df$confidential_flag)
denominator_drop_confidential <- dim(dcrb_confidential_df)[1]
percent_drop_confidential <- substr(as.character(numerator_drop_confidential / denominator_drop_confidential * 100), 1, 5)

# repeat for each year
confidential_yearly_table <- dcrb_confidential_df %>% group_by(year) %>%
  summarise(confidential_vms_records = sum(confidential_flag),
            total_vms_records = n(),
            percent_vms_records_confidential = trunc(confidential_vms_records / total_vms_records * 10000) / 100,
            .groups = 'keep') %>%
  ungroup() %>%
  arrange(year)

# write yearly result
write.csv(file = here("Confidential", "hindcast_output", output_subdir_name, "tables", "table_confidentiality_yearly.csv"),
          x = confidential_yearly_table)

# look at table
confidential_yearly_table
```

The non-confidential data has `r numerator_drop_confidential` (`r percent_drop_confidential`%) fewer records (VMS pings), where grid cells had <3 unique vessels in a given month.

```{r}
# clean up
remove(numerator_drop_confidential, denominator_drop_confidential, percent_drop_confidential)
```

Create a **fishing activity table** summarizing the monthly and yearly average # of VMS records and # unique vessels across grid cells and months.

```{r gridded-summary-tables}
# create monthly summary of gridded fishing activity
fishing_activity_monthly_table <- dcrb_5km_output_df %>%
  group_by(year, month_numeric) %>%
  summarise(
    # grid cell count
    n_grid_cells = n(),
    # VMS record count
    min_vms_records     = min(n_vms_records),
    median_vms_records  = median(n_vms_records),
    mean_vms_records    = mean(n_vms_records),
    max_vms_records     = max(n_vms_records),
    iqr_vms_records     = IQR(n_vms_records),
    # unique vessel count
    min_unique_vessels    = min(n_unique_vessels),
    median_unique_vessels = median(n_unique_vessels),
    mean_unique_vessels   = mean(n_unique_vessels),
    max_unique_vessels    = max(n_unique_vessels),
    iqr_unique_vessels    = IQR(n_unique_vessels),
    .groups = 'keep'
  ) %>%
  arrange(year, month_numeric)
# write result
write.csv(file=here("Confidential", "hindcast_output", output_subdir_name, "tables", "table_fishing_activity_monthly.csv"),
          x=fishing_activity_monthly_table)

# repeat for yearly
fishing_activity_yearly_table <- dcrb_5km_output_df %>%
  group_by(year) %>%
  summarise(
    # grid cell month and grid cell count
    n_grid_cell_months = n(),
    n_grid_cells       = n_distinct(GRID5KM_ID),
    # VMS record count
    min_vms_records     = min(n_vms_records),
    median_vms_records  = median(n_vms_records),
    mean_vms_records    = mean(n_vms_records),
    max_vms_records     = max(n_vms_records),
    iqr_vms_records     = IQR(n_vms_records),
    # unique vessel count
    min_unique_vessels    = min(n_unique_vessels),
    median_unique_vessels = median(n_unique_vessels),
    mean_unique_vessels   = mean(n_unique_vessels),
    max_unique_vessels    = max(n_unique_vessels),
    iqr_unique_vessels    = IQR(n_unique_vessels),
    .groups = 'keep'
  ) %>%
  arrange(year)
# write result
write.csv(file=here("Confidential", "hindcast_output", output_subdir_name, "tables", "table_fishing_activity_yearly.csv"),
          x=fishing_activity_yearly_table)

# look at table
fishing_activity_yearly_table
```

Create **fishing activity maps**, based on the gridded summary above. This is one of the slowest parts of the script.

```{r fishing-activity-maps}
# map fishing activity, apply to each month of data
#   data_sf: sf object
#   data_col: vector from data_sf to log-transform and map
#   log_transform: boolean to log-transform data_col on map
#   fill_range: range for the fill of data_col vector
#   bbox: bounding box to use for map
#   title: map title
#   data_col_name: name of data_col vector, used to label map legend and name output PNG
#   png_suffix: optional string, used to name output PNG
#   background_sf: sf object
fishing_activity_map <- function (data_sf, data_col, bbox, log_transform, fill_range, title, data_col_name, png_suffix='', background_sf) {
  # log-transform data, if applicable
  if (log_transform) {
    data_col <- log10(data_col)
    fill_range <- log10(fill_range)
    data_col_name <- paste0("log10_", data_col_name)
  }
  # create map
  map_output <- ggplot() +
    # plot background
    geom_sf(data=background_sf, fill='gray80') +
    # plot log-10 transformed data
    geom_sf(data=data_sf, aes(fill=data_col)) +
    # label plot
    labs(title=title,
         fill=paste(data_col_name, "per 5x5 cell")) +
    # set bounding box
    xlim(bbox[1], bbox[3]) + ylim(bbox[2], bbox[4]) +
    # set fill color and legend limits
    scale_fill_viridis(limits=fill_range) +
    # adjust legend position
    theme(legend.position="bottom")
  # save the output map
  ggsave(filename=here("Confidential", "hindcast_output",  output_subdir_name, "maps", paste0(data_col_name, "_", png_suffix, ".png")))
  # return the output map
  return(map_output)
}

# load west coast states simple features object for background
west_coast_states <- rnaturalearth::ne_states(country = 'United States of America', returnclass = 'sf') %>% 
  filter(name %in% c('California', 'Oregon', 'Washington', 'Nevada')) %>% 
  st_transform(st_crs(grid_5km))

# join geometry from 5km grid to gridded fishing summary
dcrb_5km_output_sf <- dcrb_5km_output_df %>%
  left_join(grid_5km, by=join_by(GRID5KM_ID)) %>%
  st_as_sf()

# set bounding box
bbox <- st_bbox(dcrb_5km_output_sf)

# set limits for min and max fill for # VMS pings
vms_fill_range <- range(dcrb_5km_output_sf$n_vms_records)

# get unique and ordered year, year_month and month from gridded fishing summary
month_df <- dcrb_5km_output_sf %>% 
  st_drop_geometry() %>% 
  select(year, year_month, month) %>% 
  unique() %>% 
  arrange(year_month)

# iterate over months
for (ym in month_df$year_month) {
  # get title as a combination of year and month
  y <- month_df$year[which(month_df$year_month==ym)]
  m <- month_df$month[which(month_df$year_month==ym)]
  title <- paste(y, m)
  # filter data for given month
  ym_data_sf <- dcrb_5km_output_sf %>% filter(year_month == ym)
  # make and save map of # VMS pings
  vms_map <- fishing_activity_map(data_sf = ym_data_sf,
                                  data_col = ym_data_sf$n_vms_records,
                                  bbox = bbox,
                                  log_transform = TRUE,
                                  fill_range = vms_fill_range,
                                  title = title,
                                  data_col_name = 'n_vms_records',
                                  png_suffix = ym,
                                  background_sf = west_coast_states)
}

# clean up
remove(ym, y, m, ym_data_sf, month_df, bbox, title, vms_fill_range)
```

Take static maps created above, and stitch together into an animated gif. Currently, animation gif will skip months with no fishing data, rather than show a blank map for that month - this can be changed. It will stitch together all maps in the `Confidential/maps` output directory for the given metric, even if they weren't generated with this particular run of the script. This is one of the slowest parts of the script.

```{r fishing-activity-gifs}
# create animation of fishing activity maps
#   map_directory: character, directory where map PNG files to animate are stored
#   map_metric_name: character, metric to map, used to search for PNG files in map_directory containing this metric name
fishing_activity_gif <- function (map_directory, map_metric_name) {
  # get list of files in the map directory
  file_list <- list.files(map_directory)
  # look for PNG files with metric name
  map_file_list <- file_list[grepl(map_metric_name, file_list) & grepl('.png', file_list)]
  # read in all images
  image_list <- lapply(paste0(map_directory, "/", map_file_list), image_read)
  # join and animate images
  image_gif <- image_animate(image_join(image_list), fps = 2)
  # write GIF
  image_write(image = image_gif, path = paste0(map_directory, "/animation_", map_metric_name, ".gif"))
  # return GIF
  return(image_gif)
}

# create animation for # VMS pings and # unique vessels
fishing_activity_gif(map_directory = here("Confidential", "hindcast_output", output_subdir_name, "maps"), map_metric_name = "n_vms_records")
```

```{r, include=FALSE}
# end timer
end_timer <- proc.time()
total_timer <- end_timer - start_timer
```

This script took `r round(total_timer[3]/60, 2)` minutes to run.