---
title: "12_blh_example-ranked-persistence-map"
author: "Brooke Hawkins"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# import libraries
library(tidyverse)
library(knitr)

# set seed for reproducibility
set.seed(37)
```

Create a ranked persistence map, which will visualize the top X% of fishing activity based on the number of interpolated pings for a given year in a given grid cell. X will be our cutoff or threshold (I'm using those terms interchangeably).

For simplicity, I use `year` (e.g. 2013) in these examples, but in the actual mapping code, I use use `crab_year` (e.g. 2013_2014 for Nov. 2013 - Oct. 2014).

## Simple - Coast - All time

There are a few versions of this map, let's start with the simple version for the whole coast and the whole time frame.

Consider a simple example, for three grid cells in three years.

```{r make-example-1}
ex_1_df <- tibble(year = c(rep(2013, 3), rep(2014, 3), rep(2015, 3)),
                  grid_cell = c(letters[1:3], sample(letters[1:3]), sample(letters[1:3])),
                  n_pings = c(75, 10, 15, 160, 30, 10, 60, 20, 20))
kable(ex_1_df)
```

For each grid cell, calculate how many years the grid cell had any fishing activity.

```{r simple-persistence-function}
# input: dataframe where each record represents one grid cell in one year
  # grid_cell: column for group by
# output: dataframe where each record represents one grid cell, with calculated column 'years_active'
simple_persistence_output <- function(df) {
  output_df <- df %>%
    group_by(grid_cell) %>%
    summarise(years_active = n_distinct(year), .groups = 'drop')
  return(output_df)
}
```


```{r}
ex_1_map_df <- simple_persistence_output(ex_1_df)
kable(ex_1_map_df)
```

This dataframe would then be joined to the 5km grid by `grid_cell`, and plotted on a map. Each grid cell had activity in every year, so every cell would be included on the map with 3 years of activity.

## Ranked - Coast - All time

Let's make it a little trickier. Now we only want to consider grid cells that contribute to the top X% of fishing activity. We'll still consider the whole coast and the whole time frame.

Let's consider a 75% threshold.

```{r set-threshold}
ex_threshold <- 0.75
```

To decide which grid cells contribute towards the 75% fishing activity, we need to do a couple of things:

1.  Sort each year of data from most pings to fewest pings
2.  Calculate total number of pings per year
3.  Calculate the threshold number of pings per year (if it's a fraction, round up to the nearest whole number)
4.  Calculate a cumulative sum of number of pings for each year
5.  Compare the cumulative number of pings to the threshold number of pings to decide which cells get included in the persistence map

```{r ranked-persistence-function-1}
# input: dataframe where each record represents one grid cell in one year
  # year: column for group by
  # n_pings: column to rank by and compare to threshold within each year
# output: 
  # interim_df: dataframe where each record represents one grid cell in one year, with three calculated columns
    # year_pings: pings for that year
    # threshold_pings: threshold pings for that year
    # cumsum_n_pings: cumulative pings in ranked order for that grid cell
    # include: whether the grid cell meets the threshold pings for that year
 
ranked_persistence_interim <- function(df, threshold) {
  interim_df <- df %>%
    arrange(year, n_pings * -1) %>%
    group_by(year) %>% 
    mutate(year_pings = sum(n_pings),
           threshold_pings = ceiling(threshold * year_pings),
           cumsum_pings = cumsum(n_pings),
           include = (cumsum_pings <= threshold_pings) | 
             (cumsum_pings > threshold_pings & lag(cumsum_pings) < threshold_pings) | 
             (cumsum_pings > threshold_pings & is.na(lag(cumsum_pings) < threshold_pings)))
  return(interim_df)
}
```

```{r}
ex_1_threshold_df <- ranked_persistence_interim(ex_1_df, ex_threshold)
kable(ex_1_threshold_df)
```

This logic for `include` probably seems arbitrarily complicated, but here are a few use cases considered here:

-   What if the cumulative pings is exactly equal to the threshold pings? (2013)
-   What if cumulative pings is not exactly equal to the threshold pings? (2014, 2015)
-   What if the first record reaches the threshold? (2014)

Not *exactly* 75% of the fishing effort is included; it's more accurate to say at least 75% fishing effort is included. For example, in 2014 and in 2015, 80% of the effort is included.

Edge case - I'm still not satisfied with ties at the threshold. For example, in 2015, grid cells b and c both have 20 pings, and grid cell b is included while grid cell c isn't. I haven't figured out a way to handle that exception without iterating over records one by one in a for loop, which seems slow. I did see this happen in the actual VMS data for the 2016-2017 crab year. I doubt it has huge impacts on the maps, just calling out this one remaining caveat.

Anyway, moving on.

When mapping persistence for grid cells that contribute to the top 75% of fishing activity, we now need to find how many years each of the grid cells are included. The steps are:

1.  Filter for grid cells that should be included in the summary
2.  Count the number of distinct years for each grid cell

```{r ranked-persistence-function-2}
# input: dataframe where each record represents one grid cell in one year, with three calculated columns
  # grid_cell: column for group by
  # year_pings: n_pings for that year
  # threshold_pings: threshold * n_pings for that year
  # cumsum_n_pings: cumulative n_pings in ranked order for that grid cell
  # include: whether the grid cell meets the threshold n_pings for that year 
# output: dataframe where each record represents one grid cell, with calculated column 'years_active'
ranked_persistence_output <- function(interim_df) {
  output_df <- interim_df %>%
    filter(include) %>%
    group_by(grid_cell) %>%
    summarize(years_active = n_distinct(year),
              .groups = 'drop')
  return(output_df)
}
```

```{r}
ex_1_threshold_map_df <- ranked_persistence_output(ex_1_threshold_df)
kable(ex_1_threshold_map_df)
```

Only grid cells a and b contributed towards the top 75% fishing effort, and in 2 years each.

Note: the ranked version for a 100% threshold is identical to the simple version. The 100% threshold will just do some sorting and calculating that isn't required, because every value for `include` will be true until 100% of the pings are represented.

## Ranked - State - All time

Now let's do the same exact thing, but instead of considering the whole coast, we'll look at each state separately. We can use a similar dataframe as before, but we need to have state included in the aggregation.

Consider a new example with two years and two states of data.

```{r make-example-2}
ex_2_df <- tibble(year = c(rep(2013, 5), rep(2014, 5)),
                  state = c(rep('C', 3), rep('O', 2), rep('C', 2), rep('O', 3)),
                  grid_cell = c(letters[1:5], sample(letters[1:2]), sample(letters[4:6])),
                  n_pings = c(10, 30, 60, 50, 50, 80, 20, 45, 35, 10))
kable(ex_2_df)
```

There are two versions here:

A. Ranking top X% in each state based on the whole coast's fishing activity. *Universe for analysis = whole coast.* (This just splits the map into separate panels, so the coastal trends are easier to see.)

B. Rank top X% in each state based on that state's fishing activity. *Universe for analysis = each state.* (This is a different analytical task, because the yearly totals need to be grouped by state in addition to by year.)

### Version A

For version A, we do the same process as above, just add the `state` identifier whenever the `grid_cell` identifier is grouped by for aggregations. Since each grid cell has only one state, the results are the same, just with the addition of the the `state` column.

Repeat the checks for whether a grid cell contributes to the top 75% of fishing activity:

```{r}
ex_2_a_threshold_df <- ex_2_df %>%
  arrange(year, state, n_pings * -1) %>% # state is now included here
  group_by(year) %>%
  mutate(year_pings = sum(n_pings),
         threshold_pings = ceiling(ex_threshold * year_pings)) %>%
  ungroup() %>%
  group_by(year) %>%
  mutate(cumulative_pings = cumsum(n_pings),
         include = (cumulative_pings <= threshold_pings) |
           (cumulative_pings > threshold_pings & lag(cumulative_pings) < threshold_pings) |
           (cumulative_pings > threshold_pings & is.na(lag(cumulative_pings) < threshold_pings)))
kable(ex_2_a_threshold_df)
```

Repeat the aggregation for how many years a grid cell contributes to the top 75% of fishing activity:

```{r}
ex_2_a_threshold_map_df <- ex_2_a_threshold_df %>%
  filter(include) %>%
  group_by(grid_cell, state) %>% # state is now included here
  summarize(years_with_activity = n_distinct(year),
            .groups = 'drop')
kable(ex_2_a_threshold_map_df)
```

When considering the whole coast for context, grid cells a and b in California contributed to the top 75% fishing activity in 2 years, grid cell c in California contributed in 1 year, and grid cell d in Oregon contributed in 2 years.

### Version B

For version B, we do the same process as above, but when we calculate the total number of pings and compare a grid cell against a threshold, we compare to the state total, not the whole coast total.

```{r}
ex_2_b_threshold_df <- ex_2_df %>%
  arrange(year, state, n_pings * -1) %>% # state is still included here
  group_by(year, state) %>% # state is now included here too
  mutate(year_pings = sum(n_pings),
         threshold_pings = ceiling(ex_threshold * year_pings)) %>%
  ungroup() %>%
  group_by(year, state) %>% # state is now included here too
  mutate(cumulative_pings = cumsum(n_pings),
         include = (cumulative_pings <= threshold_pings) |
           (cumulative_pings > threshold_pings & lag(cumulative_pings) < threshold_pings) |
           (cumulative_pings > threshold_pings & is.na(lag(cumulative_pings) < threshold_pings)))
kable(ex_2_b_threshold_df)
```

```{r}
ex_2_b_threshold_map_df <- ex_2_b_threshold_df %>%
  filter(include) %>%
  group_by(grid_cell, state) %>% # state is still included here
  summarize(years_with_activity = n_distinct(year),
            .groups = 'drop')
kable(ex_2_b_threshold_map_df)
```

When considering each state separately for context, grid cells a, b, and c in California contributed to the top 75% fishing activity in 1 year each, and grid cells d and e contributed to the top 75% fishing activity in two years each.

```{r}
# make a list of dataframes, one dataframe per state
state_vector <- unique(ex_2_df$state)
# split dataframes into a list by state
state_df_list <- lapply(state_vector, function(s) ex_2_df %>% filter(state == s))
# rank pings in each state separately
state_df_interim_list <- lapply(state_df_list, function(s) ranked_persistence_interim(df = s, threshold = ex_threshold))
# create grid cell summaries in each state separately
state_df_output_list <- lapply(state_df_interim_list, ranked_persistence_output)
# join dataframes across states, including state identifier
state_df_output_labelled_list <- lapply(1:length(state_vector), function(i) cbind(state = state_vector[i], state_df_output_list[[i]]))
# bind dataframes back together
state_df_output <- bind_rows(state_df_output_labelled_list)
# check result
state_df_output
```

This is the same output, but uses the `ranked_persistence_interim` and `ranked_persistence_output` functions from earlier. There is probably a better way to do this, like rewriting the functions to include state and assume the dataframes for input include that column. I'd like a version that's a little more flexible, though, including any regional level for the aggregation grouping.

## Ranked - Coast - Monthly

I'm honestly not sure I like this one yet, I think the interpretation is a bit funky.

Let's go back to thinking about the whole coast, and let's look at how many months across all years had fishing activity. 

The resulting map will show twelve maps, 1 per month. The grid cell color will indicate the number of years the cell contributed to the to 75% fishing activity for that month.

This is analogous to the *Ranked - State - All time - Version B* use case, but we'll swap `month` for `state` when aggregating.

Here's a new example dataframe, including year and month:

```{r make-example-3}
ex_3_df <- tibble(year = rep(2014, 6),
                  month = c(rep(1, 3), rep(2, 3)),
                  grid_cell = c(letters[1:3], sample(letters[1:3])),
                  n_pings = c(70, 20, 10, 90, 70, 40))
kable(ex_3_df)
```

```{r}
ex_3_year_month_threshold_df <- ex_3_df %>%
  arrange(year, month, n_pings * -1) %>%
  group_by(year, month) %>%
  mutate(year_month_pings = sum(n_pings),
         threshold_pings = ceiling(ex_threshold * year_month_pings),
         cumulative_pings = cumsum(n_pings),
         include = (cumulative_pings <= threshold_pings) |
           (cumulative_pings > threshold_pings & lag(cumulative_pings) < threshold_pings) |
           (cumulative_pings > threshold_pings & is.na(lag(cumulative_pings) < threshold_pings)))
kable(ex_3_year_month_threshold_df)
```
```{r}
ex_3_month_threshold_map_df <- ex_3_year_month_threshold_df %>%
  filter(include) %>%
  group_by(grid_cell, month) %>%
  summarize(years_with_activity = n_distinct(year),
            .groups = 'drop')
kable(ex_3_month_threshold_map_df)
```

The January map would show that grid cell a and b both contributed to 1 year of the top 75% fishing activity, across all Januaries.

The February map would show that grid cell and c contributed to 1 year of the top 75% of fishing activity, across all Februaries.

## Ranked - Coast - Yearly

I'm honestly not sure I like this one yet, I think the interpretation is a bit funky.

This is analogous to the *Ranked - Coast - Monthly* use case, only the final aggregation step and interpretation is different.

The resulting map will show thirteen maps, 1 per year. The grid cell color will indicate the number of months the cell contributed to the to 75% fishing activity for that year.

```{r}
ex_3_year_threshold_map_df <- ex_3_year_month_threshold_df %>%
  filter(include) %>%
  group_by(grid_cell, year) %>% # this is now for month
  summarize(months_with_activity = n_distinct(month),
            .groups = 'drop')
kable(ex_3_year_threshold_map_df)
```

The 2014 map would show that grid cell a contributed to the top 75% of fishing activity of a given month for two months in the year, and grid cells b and c contributed to the top 75% of fishing activity of a given month for one month in the year.
