---
title: "Check new download of fish tickets"
format: 
  gfm:
    toc: true
    embed-resources: true
---

## Check 1 - How to load all fish tickets at once?

```{r, include=FALSE}
library(here)
library(tidyverse)
```

Write functions to compare dataframe columns. This might be overkill, but I didn't know a good function that already exists in R to do this.

```{r}
# print whether column count, name, and type match between two dataframes
compare_df_cols <- function(df1, df2) {
  # check column count
  same_col_n <- ncol(df1) == ncol(df2)
  print(paste("same # columns?", same_col_n))

  # check column name
  if (same_col_n) {
    same_col_names <- sum(colnames(df1) == colnames(df2))
    if (same_col_names == ncol(df1)) {
      print(paste('same column names?', TRUE))
    } else {
      print(paste(
        'same column names?',
        FALSE,
        ncol(df1) - same_col_names,
        "differ"
      ))
    }
  }

  # check column type
  if (same_col_n & same_col_names) {
    coltypes1 = sapply(df1, class)
    coltypes2 = sapply(df2, class)
    same_col_types <- sum(coltypes1 == coltypes2)
    if (same_col_types == ncol(df1)) {
      print(paste('same column types?', TRUE))
    } else {
      types_df = tibble(
        col_name = colnames(df1),
        coltypes1,
        coltypes2,
        same = (coltypes1 == coltypes2)
      )
      print(paste('same column types?', FALSE))
      print(types_df %>% filter(same == FALSE))
    }
  }
}
```

Apply functions to pairs of dataframes from 2011-2024.

```{r}
# iterate through years
first_year <- 2011

for (y in 2011:2023) {
  # print progress
  print("---")
  print(paste("comparing year", y, "to", y + 1))

  # load df1
  if (y == first_year) {
    df1 <- read.csv(here(
      'Confidential',
      'raw_data',
      'fish_tickets',
      paste0('fish_tickets_', y, '.csv')
    ))
  } else {
    df1 <- df2
  }

  # load df2
  df2 <- read.csv(here(
    'Confidential',
    'raw_data',
    'fish_tickets',
    paste0('fish_tickets_', y + 1, '.csv')
  ))

  # compare
  compare_df_cols(df1, df2)
}
```

Interpretation:

There's the same number columns and column names year over year, but eight columns are read with different data types by default across years.

Specify the less restrictive data type (underlined) when reading in the CSV, and then `bind_rows()` should work on the dataframes read by CSV.

-   *HILLE_PERMIT* is logical in 2011, [character]{.underline} in 2012-2016, logical in 2017, character in 2018-2024

-   *DEALER_NUM* is integer in 2011-2014, [character]{.underline} in 2015-2024

-   *FTID* is [character]{.underline} in 2011-2016, integer in 2017, character in 2018-2024

-   *NUM_OF_FISH* is [numeric]{.underline} in 2011-2016, integer in 2017, numeric in 2018-2024

-   *EFP_CODE* is logical in 2011-2016, [character]{.underline} in 2017-2024

-   *EFP_NAME* is logical in 2011-2016, [character]{.underline} in 2017-2024

-   *GF_PERMIT_NUMBER* is logical in 2011-2018, [character]{.underline} in 2019-2024

-   *TRIP_SEQ* is [integer]{.underline} 2011-2019, logical 2020-2023, integer 2024

When I ran `bind_rows()` below, at first I got an error about `GEAR_CODE` not matching between integer and double in one set of years, so I also set that to read in as a double.

```{r}
for (y in 2011:2024) {
  # load year of data as df_y
  assign(
    x = paste0("df_", y),
    value = read_csv(
      here(
        'Confidential',
        'raw_data',
        'fish_tickets',
        paste0('fish_tickets_', y, '.csv')
      ),
      col_types = list(
        HILLE_PERMIT = "c",
        DEALER_NUM = "c",
        FTID = "c",
        NUM_OF_FISH = "n",
        EFP_CODE = "c",
        EFP_NAME = "c",
        GF_PERMIT_NUMBER = "c",
        TRIP_SEQ = "i",
        GEAR_CODE = "d"
      )
    )
  )
}

# bind rows
ticket_df <- bind_rows(mget(paste0("df_", 2011:2024)))
```

It looks like `df_2022` leads to some issues when parsing, unclear to me why. I ran problems`(df_2022) %>% View()` and the issue was always with column 105, which is `HSFCA_PERMIT`. It only impacts ~700 records out of 379k records that year, so it's probably ok to not worry about it. None of the other years threw warnings.

## Check 2 - How does new pull of fish ticket data compare to old pull?

I want to see if the # tickets, vessels, landings and revenue are pretty similar between the old and new data.

This is rough but I'm hoping numbers will look pretty similar between pulls.

```{r}
# calculate # tickets, vessels, landings and revenue per year
new_pull_counts <- ticket_df |>
  # apply a few filters
  filter(
    # remove missing vessel ID
    VESSEL_NUM != "MISSING",
    VESSEL_NUM != "UNKNOWN",
    VESSEL_NUM != "",
    # remove tickets landed in AK and transported to WC
    COUNCIL_CODE != "N",
    # remove tribal fishing
    FLEET_CODE != "TI",
    # remove aquaculture
    PARTICIPATION_GROUP_CODE != "A"
  ) |>
  # select a few columns
  select(
    LANDING_YEAR,
    VESSEL_NUM,
    FTID,
    PACFIN_SPECIES_CODE,
    LANDED_WEIGHT_LBS,
    EXVESSEL_REVENUE
  ) |>
  # de-duplicate
  distinct() |>
  # calculate a few summary statistics per year
  group_by(LANDING_YEAR) |>
  summarise(
    n_vessels = n_distinct(VESSEL_NUM),
    n_tickets = n_distinct(FTID),
    sum_landings = sum(LANDED_WEIGHT_LBS, na.rm = TRUE),
    sum_revenue = sum(EXVESSEL_REVENUE, na.rm = TRUE)
  )

# load old data
old_ticket_df <- read_rds(here(
  'Confidential',
  'raw_data',
  'fish_tickets',
  'all_fishtickets_1994_2023.rds'
))

# calculate # tickets, vessels, landings and revenue per year
old_pull_counts <- old_ticket_df |>
  # apply a few filters
  filter(
    # remove missing vessel ID
    VESSEL_NUM != "MISSING",
    VESSEL_NUM != "UNKNOWN",
    VESSEL_NUM != "",
    # remove tickets landed in AK and transported to WC
    COUNCIL_CODE != "N",
    # remove tribal fishing
    FLEET_CODE != "TI",
    # remove aquaculture
    PARTICIPATION_GROUP_CODE != "A"
  ) |>
  # select a few columns
  select(
    LANDING_YEAR,
    VESSEL_NUM,
    FTID,
    PACFIN_SPECIES_CODE,
    LANDED_WEIGHT_LBS,
    EXVESSEL_REVENUE
  ) |>
  # de-duplicate
  distinct() |>
  # calculate a few summary statistics per year
  group_by(LANDING_YEAR) |>
  summarise(
    n_vessels = n_distinct(VESSEL_NUM),
    n_tickets = n_distinct(FTID),
    sum_landings = sum(LANDED_WEIGHT_LBS, na.rm = TRUE),
    sum_revenue = sum(EXVESSEL_REVENUE, na.rm = TRUE)
  )

# set a column indicating which pull the data is from
old_pull_counts$pull = "old"
new_pull_counts$pull = "new"

# calculate difference between old and new pull in similarly shaped dataframe
diff_counts <- old_pull_counts |>
  left_join(new_pull_counts, by = "LANDING_YEAR", suffix = c("_old", "_new")) |>
  mutate(
    n_vessels = n_vessels_new - n_vessels_old,
    n_tickets = n_tickets_new - n_tickets_old,
    sum_landings = sum_landings_new - sum_landings_old,
    sum_revenue = sum_revenue_new - sum_revenue_old,
    pull = "diff",
    pct_vessels = n_vessels / n_vessels_old * 100,
    pct_tickets = n_tickets / n_tickets_old * 100,
    pct_landings = sum_landings / sum_landings_old * 100,
    pct_revenue = sum_revenue / sum_revenue_old * 100
  ) |>
  select(
    LANDING_YEAR,
    n_vessels,
    n_tickets,
    sum_landings,
    sum_revenue,
    pull,
    pct_vessels,
    pct_tickets,
    pct_landings,
    pct_revenue
  )

# view differences
diff_counts

# write csv
write.csv(
  diff_counts,
  here('Confidential', 'sablefish_vms', 'diff_pull_counts.csv'),
  row.names = FALSE
)
```

```{r}
vis_df <- diff_counts |>
  select(LANDING_YEAR, n_vessels, n_tickets, sum_landings, sum_revenue, pull) |>
  rbind(old_pull_counts, new_pull_counts) |>
  pivot_longer(
    cols = c("n_vessels", "n_tickets", "sum_landings", "sum_revenue"),
    names_to = "variable"
  )
vis_df |>
  ggplot(aes(x = LANDING_YEAR, y = value, col = pull)) +
  geom_line() +
  facet_grid(
    variable ~ factor(pull, c("old", "new", "diff")),
    scales = "free_y"
  )
ggsave(
  here('Confidential', 'sablefish_vms', 'diff_pull_counts.png'),
  width = 8,
  height = 5,
  unit = "in"
)
```

I saw a huge gap in 2017, ~10x fewer records than there should have been. I redownloaded the data and reran to see if the new results look better, which they do. The differences now are <1% for the distinct count of vessels and tickets, and the sum of landings and revenue.